---
title: ''
# fontfamily: mathpazo
fontsize: 10pt
geometry: margin=3mm
#linestretch: 0.1
classoption:
- portrait
pagenumbering: TRUE
#whitespace: none
output:
  html_document:
    toc: FALSE
    toc_depth: 3
    toc_float:
      collapsed: yes
      smooth_scroll: yes
  pdf_document:
    toc: FALSE
    number_sections: FALSE
header-includes:
    - \usepackage{multicol}
    - \usepackage{multirow}
    - \usepackage{caption}
    - \setlength\tabcolsep{0.1pt}
    - \setlength\lineskip{0pt}
    - \setlength\parskip{0pt}
---


# 

### 20191002 {#kernel}
[kernel](#kernel)


- Cat and Dog problem

- A simple geometric solution

- A more general solution

Dot product $\vec a\vec b=a_xb_x+a_yb_y=|\vec a||\vec b|\cos(\theta)$

Exercise: 

$g(x)=<C_+-C_-,X-C>=<C_+,X>-<C_-,X>-<C_+,C>+<C_-,C>$;

$<C_+,X>=<\frac1{n_{+}}\sum\limits_{l\in I_+}^nx_i,x>$;

$<C_-,X>=<\frac1{n_{-}}\sum\limits_{l\in I_-}^nx_i,x>$;

$<C_+,C>=<C_+,\frac12C_+>+<C_+,\frac12C_->=\frac1{2n_{+}^2}\sum\limits_{(i,j)\in I_{+}}<x_i,x_j>+\frac12<C_+,C_->$

$<C_-,C>=<C_-,\frac12C_+>+<C_-,\frac12C_->=\frac12<C_+,C_->+\frac1{2n_{-}^2}\sum\limits_{(i,j)\in I_{-}}<x_i,x_j>$


$g(x)=\sum_{l=1}^n\alpha_i<x_i,x>+b$, 

$b=\frac12\left[\frac1{n_{-}^2}\sum\limits_{(i,j)\in I_{-}}<x_i,x_j>-\frac1{n_{+}^2}\sum\limits_{(i,j)\in I_{+}}<x_i,x_j>\right]$

$\alpha_i=\begin{cases}\frac1{n_{+}}&y_i=+1\\-\frac1{n_{-}}&y_i=-1\end{cases}$

\fontsize{10pt}{0pt}
\footnotesize

- Import the iris data

```{r,echo=T,message=F,warning=F,out.width='100%'}
rm(list=ls())
library(datasets)
data(iris)
```

```{r,echo=T,message=F,warning=F,out.width='100%',fig.align='center'}
library(ggplot2)
GGally::ggpairs(iris,mapping=aes(color =Species,shape=Species,alpha=0.3),
        columns=c("Sepal.Length","Sepal.Width","Petal.Length","Petal.Width","Species"))+theme_light()
```

- Define the train and test sets

```{r,echo=T,message=F,warning=F,out.width='100%'}
iris$class <- NA
iris_setosa <- iris[iris$Species=="setosa",]
iris_versicolor <- iris[iris$Species=="versicolor",]
iris_virginica <- iris[iris$Species=="virginica",]

iris_train_se<- iris_setosa[1:40,]
iris_train_ve<- iris_versicolor[1:40,]
iris_train_vi<- iris_virginica[1:40,]

iris_test_se<- iris_setosa[41:50,]
iris_test_ve<- iris_versicolor[41:50,]
iris_test_vi<- iris_virginica[41:50,]

iris_train_se.ve<- rbind(iris_train_se,iris_train_ve)
iris_train_ve.vi<- rbind(iris_train_ve,iris_train_vi)
iris_test_se.ve<- rbind(iris_test_se,iris_test_ve)
iris_test_ve.vi<- rbind(iris_test_ve,iris_test_vi)
```


```{r,eval=F,include=F,message=F,warning=F,out.width='100%'}
# Define a kernel function and classifier
k <- function(x,y){return(x %*% t(y))}
classifier <- function(test,train_po,train_ne){
  
  for(i in 1:nrow(train_po)){
    a_po = 1/nrow(train_po)*sum(k(as.matrix(test),train_po[i,]))
  }
  for(i in 1:nrow(train_ne)){
    a_ne = 1/nrow(train_ne)*sum(k(as.matrix(test),train_ne[i,]))
  }
  
  for(i in 1:(nrow(train_po)-1)){
    for (j in (i+1):nrow(train_po)){
      b_po = -0.5*(1/nrow(train_po)^2)*sum(k(as.matrix(train_po[i,]),train_po[j,]))
    }
  }
  for(i in 1:(nrow(train_ne)-1)){
    for (j in (i+1):nrow(train_ne)){
      b_ne = 0.5*(1/nrow(train_ne)^2)*sum(k(as.matrix(train_ne[i,]),train_ne[j,]))
    }
  }
  
  class = a_po+a_ne+b_po+b_ne
  return(class)
}

for (i in 1:20){
iris_test_se.ve[i,6] <- classifier(iris_test_se.ve[i,1:4],iris_train_se[,1:4],iris_train_ve[,1:4])
}
```

```{r,eval=F,include=F,message=F,warning=F,out.width='100%'}
# define a 3D grid
kd <- with(iris_test_se.ve, MASS::kde2d(iris_train_se.ve$Petal.Length,iris_train_se.ve$Petal.Width, n = 20))
kd <- MASS::kde2d(iris_train_se.ve$Petal.Length,iris_train_se.ve$Petal.Width, n = 20)
```

```{r,eval=F,include=F,message=F,warning=F,out.width='100%'}
# Plot the test and 3D grid
library(plotly)
plot_ly (type='surface',x=kd$x,y=kd$y,z =kd$z, opacity=0.95)%>%add_markers( x = ~iris_test_se.ve$Petal.Length, y = ~iris_test_se.ve$Petal.Width, z = ~iris_test_se.ve$class-2, color = ~iris_test_se.ve$Species,size =10, symbol=iris_test_se.ve$Species, symbols = c('circle-open','cross'))
```

```{r,eval=F,include=F,message=F,warning=F,out.width='100%'}
for (i in 1:100){
iris[i,6] <- classifier(iris[i,1:4],iris_train_se.ve_setosa[,1:4],iris_train_se.ve_versicolor[,1:4])
}
plot_ly (type='surface',x=kd$x,y=kd$y,z =kd$z)%>%add_markers( x = ~iris$Petal.Length, y = ~iris$Petal.Width, z = ~iris$class-2, color = ~iris$Species,size =1 )
```

- Define the kernel function and Computing the classifier

```{r,echo=T,message=F,warning=F,out.width='100%'}
k = function(x,y)  return(sum(x*y)+1)
k.pp=outer(1:40,1:40,Vectorize(function(i,j) k(iris_train_se[i,1:4],iris_train_se[j,1:4])))
k.mm=outer(1:40,1:40,Vectorize(function(i,j) k(iris_train_ve[i,1:4],iris_train_ve[j,1:4])))
b=(sum(k.mm)/(40^2)-sum(k.pp)/(40^2))/2
alpha=ifelse(iris_train_se.ve$Species=="setosa",1/40,-1/40)

k.x=outer(1:80,1:20,Vectorize(function(i,j) k(iris_train_se.ve[i,1:4],iris_test_se.ve[j,1:4])))
iris_test_se.ve[,6]=(t(k.x)%*%alpha+b)

```

```{r,echo=T,message=F,warning=F,out.width='100%'}
k = function(x,y)  return(sum(x*y)+1)
k.pp=outer(1:40,1:40,Vectorize(function(i,j) k(iris_train_vi[i,1:4],iris_train_vi[j,1:4])))
k.mm=outer(1:40,1:40,Vectorize(function(i,j) k(iris_train_ve[i,1:4],iris_train_ve[j,1:4])))
b=(sum(k.mm)/(40^2)-sum(k.pp)/(40^2))/2
alpha=ifelse(iris_train_ve.vi$Species=="virginica",1/40,-1/40)

k.x=outer(1:80,1:20,Vectorize(function(i,j) k(iris_train_ve.vi[i,1:4],iris_test_ve.vi[j,1:4])))
iris_test_ve.vi[,6]=(t(k.x)%*%alpha+b)

```


```{r,echo=F,message=F,warning=F,out.width='100%'}
# Define a grid 
g.n=20
x.min=min(iris_test_se.ve[,3:4])
x.max=max(iris_test_se.ve[,3:4])
z.hat=matrix(NA,nrow=g.n,ncol=g.n)
g=seq(from=x.min,to=x.max,length.out=g.n)
for (i in (1:g.n)){
  for (j in (1:g.n)){
    u=c(0,0,g[i],g[j])
    k.x=outer(1:80,1,Vectorize(function(i,j) k(as.matrix(iris_train_se.ve[i,1:4]),u)))
    z.hat[i,j]=sum(k.x*alpha)+b
  }
}
```



```{r,echo=F,message=F,warning=F,out.width='100%',dev='pdf'}
# Plot the test and grid
z.zero=matrix(0,nrow=20,ncol=20)
library(plotly)
p <- plot_ly (type = 'surface' , x=g,y=g,z=(z.hat), opacity = 0.9 ,showscale =T )%>%
  add_markers( x = iris_test_se.ve$Petal.Length, y = iris_test_se.ve$Petal.Width, z = iris_test_se.ve$class, 
               color =iris_test_se.ve$Species, size=1,
               symbol=iris_test_se.ve$Species, symbols = c('circle-open','cross'),
               opacity = 1)%>%
  layout(scene = list(xaxis = list(title = 'Petal.Length'),
                     yaxis = list(title = 'Petal.Width'),
                     zaxis = list(title = 'Class')))

```

- Evaluate the classifier

```{r,echo=T,message=F,warning=F,out.width='100%'}
iris_test_se.ve$evaluate=ifelse(iris_test_se.ve$class>0,"setosa","versicolor")
1-length(which(iris_test_se.ve$Species==iris_test_se.ve$evaluate))/20

iris_test_ve.vi$evaluate=ifelse(iris_test_ve.vi$class>0,"virginica","versicolor")
1-length(which(iris_test_ve.vi$Species==iris_test_ve.vi$evaluate))/20
```

\begin{center}
    \captionof{table}{Confusion matrix}
    \label{tab:confusion-matrix}
\begin{tabular}{|c|c|c|c||c|c|c|}
\hline
\multirow{2}{*}{}& \multicolumn{6}{c|}{Actural Species} \\
\cline{2-7}
                             & test 1 & Setosa & Versicolor & test 2 & Virginica & Versicolor \\\hline

\multirow{2}{*}{Test Species}& Setosa & 10 & 0 & Virginica & 9 & 0\\
\cline{2-7}
                             & Versicolor & 0 & 10 & Versicolor & 1 & 10\\
\hline
\end{tabular}
  \end{center}

Error rate = 0% and 5% in two tests respectively.

```{r,echo=F,message=F,warning=F,out.width='100%'}
library(pander)
library(dplyr)
pander(bind_cols(iris_test_se.ve[,c(5:7)],iris_test_ve.vi[,c(5:7)]),caption="setosa v.s.versicolor/  /virginica v.s.versicolor")
```

```{r,eval=F,include=F,message=F,warning=F,out.width='100%'}
library(kableExtra)
kable(iris_test_se.ve[,c(5:7)],format="latex",booktabs=F)%>%
     kable_styling(c("striped", "bordered"), full_width = F,font_size = 8)%>%
     add_header_above(c("setosa v.s.versicolor" = 4))
kable(iris_test_ve.vi[,c(5:7)],format="latex",booktabs=F)%>%
     kable_styling(c("striped", "bordered"), full_width = F,font_size = 8)%>%
     row_spec(13,background ="yellow")%>%
     add_header_above(c("virginica v.s.versicolor" = 4))
```

