---
title: "STAT565_Lab"
author: "Shen Qu"
date: 'Mar 11, 2019'
output:
  pdf_document: 
geometry: margin=0.25in
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,cache=FALSE, size="tiny", results="markup", tidy=F)
```

```{r eval=FALSE, include=FALSE}
install.packages("ggplot2")
install.packages("ggpubr")
install.packages("mosaic")
install.packages("agricolae")
install.packages("DescTools")
install.packages("emmeans")
```

# Problem 1 Unbalanced 2-Factor Factorial (Non-significant interaction)

  \textcolor[rgb]{0.5,0.5,0.5}{Source: R.D. White, Jr. (1999). "Are Women More Ethical? Recent Findings on the Effects of Gender Under Moral Development," Journal of Public Administration Research and Theory, Vol. 9, 3, pp.459-471. Description: Ethics scores for a sample of members of the U.S. Coast Guard by Gender (1=Male, 2=Female) and Rank (1=Officer, 2=Enlisted). Data simulated to match cell means and SDs.}

  \textcolor[rgb]{0.5,0.5,0.5}{Variables/Columns:Gender, 8; Rank, 16; Ethics Score,18-24.}

```{r include=FALSE}
library(readxl)
library(ggplot2)
library(ggpubr)
library(mosaic)
library(agricolae)
library(DescTools)
library(dplyr)
library(car)
library(pander)
library(emmeans)
table_ethics <- read.table("Ethics.txt",header = T)
table_ethics$g <-  as.factor(table_ethics$Gender)
table_ethics$r <-  as.factor(table_ethics$Rank)
model_ethics <- aov(Score ~ g*r, data=table_ethics)
```

(a)  \textcolor[rgb]{0.5,0.5,0.5}{Plot the data and report the plot here (A plot with data and means of treatment combinations). Do not report code here. Describe the observed relationship between two factors.}

```{r echo=FALSE, out.width='30%',fig.show='hold',message=F}
str(table_ethics)
ggplot(data = table_ethics, aes(x = r, y = Score, colour=g,shape = g)) +
   stat_summary() + labs(y= "Ethics Score", x="Rank", color="Gender", shape="Gender")
ggline(data = table_ethics, x ="r", y ="Score", add = c("mean", "jitter"),shape="g",  color ="g", linetype ="g", ylab= "Ethics Score", xlab="Rank")
```


(b)  \textcolor[rgb]{0.5,0.5,0.5}{Obtain the numerical summary for each treatment combination and factor levels separately. Report them here in a tabular form}

```{r echo=FALSE}
summary_table_gender<- favstats(Score ~ Gender, data=table_ethics)
summary_table_gender$Gender <- paste("Gender",summary_table_gender$Gender)
summary_table_gender <- rename(summary_table_gender,factorial=Gender)
summary_table_rank<- favstats(Score ~ Rank, data=table_ethics)
summary_table_rank$Rank <- paste("Rank",summary_table_rank$Rank)
summary_table_rank <- rename(summary_table_rank,factorial=Rank)
summary_table_g_r<- favstats(Score ~ Gender+Rank, data=table_ethics)
summary_table_g_r$Gender.Rank <- paste("GR",summary_table_g_r$Gender.Rank)
summary_table_g_r <- rename(summary_table_g_r,factorial=Gender.Rank)
pander(bind_rows(summary_table_gender, summary_table_rank, summary_table_g_r))
```

(c)  \textcolor[rgb]{0.5,0.5,0.5}{Fit the two-factor factorial model and report the complete ANOVA table here. Do not report code here. The complete ANOVA table should have a row for each of the following: main effects of each treatment, two-factor interaction effects, error and total.Report the adjusted (Type III) SS for each effect because of unbalanced design. Note that, adjusted SS and SSE do not add up to SSTotal}

```{r eval=FALSE, include=FALSE}
pander(summary(model_ethics))
sum((table_ethics$Score- mean(table_ethics$Score))^2)

options(contrasts = c("contr.sum","contr.poly"))
drop1(model_ethics, .~., test="F")
pander(Anova(model_ethics, type="III"))
```

 --------------------------------------------------------------
    &nbsp;       Df    Sum Sq      Mean Sq   F value    Pr(>F)   
 --------------- ----- ---------  --------- --------- -----------
      **g**        1    860.9      860.9     6.034     0.01461  

   **r**           1     3048      3048      21.36    5.693e-06 

   **g:r**         1     34.1      34.1      0.239     0.6253   

  **Residuals**   295   42087      142.7      NA         NA     
  
   **Total**      298   46030.1    154.46     NA         NA     
 --------------------------------------------------------------

 ------------------------------------------------------
    Type III       Sum Sq   Df    F value    Pr(>F)   
 ----------------- -------- ----- --------- -----------
 **(Intercept)**   180136    1     1263     1.327e-108 

   **g**           785.2     1     5.503     0.01964   

   **r**            2021     1     14.17     0.000202  

   **g:r**          34.1     1     0.239      0.6253   

  **Residuals**    42087    295     NA          NA     
 
   **Total**       46030.1   298     NA         NA   
 ------------------------------------------------------



(d)  \textcolor[rgb]{0.5,0.5,0.5}{Based on the ANOVA table write your conclusion appropriately. Perform all the necessary tests and report the conclusion along with the p-value. (no need to do pairwise comparison to save the time)}

The line plot shows that two lines are approximately parallel. Difference in score between ranks is same for different genders. There is not obvious interaction effect.

According to ANOVA table, there is not a significant interaction effect from ranks and genders on the score at 5% significance level (P-value=0.6253). That means, effect of ranks and effect of genders on score is independent. Therefore, examine the main effects.

According to type III ANOVA table, there is a significant effect from genders and ranks on the score at 5% significance level (P-value=0.01964, 0.000202, respectively
).

(e)  \textcolor[rgb]{0.5,0.5,0.5}{Report the code here without output.}

```{r, eval =F, size="tiny"}
table_ethics <- read.table("Ethics.txt",header = T)
table_ethics$g <-  as.factor(table_ethics$Gender)
table_ethics$r <-  as.factor(table_ethics$Rank)
model_ethics <- aov(Score ~ g*r, data=table_ethics)
str(table_ethics)
ggplot(data = table_ethics, aes(x = r, y = Score, colour=g,shape = g)) +
   stat_summary() + labs(y= "Ethics Score", x="Rank", color="Gender", shape="Gender")
favstats(Score ~ Gender, data=table_ethics)
favstats(Score ~ Rank, data=table_ethics)
favstats(Score ~ Gender+Rank, data=table_ethics)
pander(summary(model_ethics))
sum((table_ethics$Score- mean(table_ethics$Score))^2)
# To obtain Adjusted SS for factorial design in R when there is an unbalanced design #
# First, it is necessary to set the contrasts option in R.#
options(contrasts = c("contr.sum","contr.poly"))
# Use wither drop1 function or Anova function with type = "III" as below #
#Type III sum of squares give the adjusted (for all other effects) SS and p values of each effect #
drop1(model_ethics, .~., test="F")
Anova(model_ethics, type="III")
```

# Problem 2: 3-Factor Factorial (Non-significant interaction)

  \textcolor[rgb]{0.5,0.5,0.5}{Results from 3-Factor ANOVA to investigate effects of poker skill (average/expert), poker hand (bad/neutral/good), and bet limit (fixed/none) on winnings. There were 25 individuals in each of 12 combinations (each individual in only one treatment).Use a=0.05}

  \textcolor[rgb]{0.5,0.5,0.5}{Variables/Columns:
Skill: 1=Expert, 2=Average; Hand:1=Bad, 2=Neutral, 3=Good; Limit: 1=Fixed, 2=None; Final Cash Balance (euros)}

```{r include=FALSE}
table_poker <- read.csv("Poker.csv")
table_poker$s <- factor(table_poker$skill); levels(table_poker$s) <- c("Exp", "Nov")
table_poker$h <- factor(table_poker$hand); levels(table_poker$h) <- c("Bad","Neu","Goo")
table_poker$l <- factor(table_poker$limit); levels(table_poker$l) <- c("Fix","Non")
model_poker <- aov(cash ~ s*h*l, data=table_poker)
```

(a)  \textcolor[rgb]{0.5,0.5,0.5}{Plot the data and report the plot here (A plot with data and means of treatment combinations). Do not report code here.}

```{r echo=FALSE, out.width='30%',fig.show='hold',message=F}
str(table_poker)
qplot(x=s, y=cash, data=table_poker, facets=l~h, main="Interaction Plot of Cash Balance",xlab="Skill", ylab="Final cash balance in Euro")+theme_bw()
ggplot(data =table_poker, aes(x=s, y=cash, colour=l, shape=h, group=h))+stat_summary() + labs(y= "Mean Cash Balance", x="Skill", shape="Hand")
ggline(data = table_poker, x ="s", y ="cash", add = c("mean", "jitter"),shape="l",  color ="h", linetype ="h", ylab= "Mean Cash Balance", xlab="Skill")
```


(b)  \textcolor[rgb]{0.5,0.5,0.5}{Obtain the numerical summary for each treatment combination and factor levels separately. Report them here in a tabular form}

```{r echo=FALSE, size="tiny"}
pander(favstats(cash ~ s, data=table_poker))
pander(favstats(cash ~ h, data=table_poker))
pander(favstats(cash ~ s+h+l, data=table_poker))
```

(c)  \textcolor[rgb]{0.5,0.5,0.5}{Fit the three-factor factorial model and report the complete ANOVA table here. Do not report code here. The complete ANOVA table should have a row for each of the following: main effects of each treatment, two-factor interaction effects, error and total.}

```{r eval=FALSE, include=FALSE}
pander(summary(model_poker))
sum((table_poker$cash- mean(table_poker$cash))^2)
```

  --------------------------------------------------------------
    &nbsp;         Df    Sum Sq    Mean Sq   F value    Pr(>F)   
  --------------- ----- --------- --------- --------- -----------
      **s**        1     49.17     49.17     2.839     0.09308  

   **h**           2     2647      1323      76.41    2.389e-27 

   **l**           1     31.68     31.68     1.829     0.1773   

   **s:h**         2     219       109.5     6.324    0.002052  

   **s:l**         1     119.1     119.1     6.878    0.009191  

   **h:l**         2     97.29     48.64     2.809     0.06192  

   **s:h:l**       2     42.38     21.19     1.224     0.2957   

  **Residuals**   288    4987      17.32      NA         NA     

   **Total**      299    8192.67   27.40     NA         NA       
  --------------------------------------------------------------


(d)  \textcolor[rgb]{0.5,0.5,0.5}{Based on the ANOVA table write your conclusion appropriately. Perform all the necessary tests and report the conclusion along with the p-value.}

The line plot shows that not all lines are parallel. Difference in cash between skills is not same for different hands. There could be an interaction effect.

According to ANOVA table, there is not a significant interaction effect from skills, hands and limits on the cash at 5% significance level (P-value=0.2957). there is a significant interaction effect from skills versus hands and skills versus limits on the cash at 5% significance level (P-value=0.002052, 0.009191, respectively). That means, effect of skills versus hands and skills versus limits on the cash are not independent. Therefore, examine the simple effects.

(e)  \textcolor[rgb]{0.5,0.5,0.5}{Suppose that we want to compare the average final cash balance between two types of Skill (Expert versus Average) when the type of Hand is 3 (Good) and Limit is 1 (Fixed). This is the simple effect contrast we learned in class. Test this contrast and report conclusion with p value.}

The $H0$ for contrast is $\mu_{131}-\mu_{231}=0$. The $H1$ is they don't equal zero.

The P-value of this contrast test is 0.0019, which is small enough. We can reject the $H0$ and conclude that the average final cash balance between two types of Skill (Expert versus Average) when the type of Hand is 3 (Good) and Limit is 1 (Fixed) is different at 5% significance level.

(f)  \textcolor[rgb]{0.5,0.5,0.5}{Report the code here without output.}

```{r eval=F, include=T}
#If the 3-factor interaction is significant, then test 2-factor interaction for each level of 3rd factor#
#To do all pairwise comparisons at each level of skill(s) factor  #
contrast(lsmeans(model_poker, ~h*l|s), interaction = "pairwise") # You may use and adjustment method as shown in the next contrast #
#To do all pairwise comparisons at each level of hand (h) factor  #
contrast(lsmeans(model_poker, ~s*l|h), interaction = "pairwise", adjust="bonferroni")
#To write contrasts in terms of cell means #
#First, create a 3-factor interaction term of factors s, h and l #
#Then fit the model with that factor #
table_poker$shl<-interaction(table_poker$s,table_poker$h,table_poker$l)
# Save the LSmeans for interaction terms of factors s, h and l #
# Before writing a contrast, check the order of terms in previous lsmf output #
# Then create a vector of contrast in terms of LSmeans of treatment combinations #
# Below is the contrast to test mean for two levels of skill factor (Expert versus Average) #
# when the type of Hand is 3 (Good) and Limit is 1 (Fixed) #
summary(contrast(lsmeans(lm(cash~shl, table_poker), 'shl'), list(c1 = c(rep(0,4), 1, -1, rep(0,6)))))
```
