---
title: "STAT565_Lab"
author: "Shen Qu"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  pdf_document:
#    highlight: pygments
geometry: margin=0.25in
# fontfamily: mathpazo
# fontsize: 12pt
# spacing: double
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,cache=FALSE, size="tiny",cache=FALSE, results="markup", tidy=T)

# theme: default, cerulean, journal, flatly, darkly, readable, spacelab, united, cosmo, lumen, paper, sandstone, simplex, and yeti

# highlight: default, tango, pygments, kate, monochrome, espresso, zenburn, haddock, and textmate

#    includes:
#      in_header: preamble.tex
#      before_body: doc-prefix.tex
#      after_body: doc-suffix.tex

#    df_print: default, kable, tibble

#    citation_package: natbib, biblatex

# size = c("normalsize", "tiny", "scriptsize", "footnotesize", "small", "large", "Large", "LARGE", "huge", "Huge")
```

```{r eval=FALSE, include=FALSE}
install.packages("ggplot2")
install.packages("ggpubr")
install.packages("mosaic")
install.packages("agricolae")
install.packages("DescTools")
```

```{r include=FALSE}
library(readxl)
library(ggplot2)
library(ggpubr)
library(mosaic)
library(agricolae)
library(DescTools)
library(dplyr)
table_protein <- read_excel("Protein.xlsx")
table_protein$Trt1 = as.factor(table_protein$Source)
table_protein$Trt2 = as.factor(table_protein$Amount)
model_protein <- aov(Gain ~ Trt1*Trt2, data=table_protein)
```

(a) Plot the data and report the plot here (A plot with data and means of treatment combinations). Do not report code here. Describe the observed relationship between two factors.

```{r echo=FALSE, out.width='50%'}
glimpse(table_protein)
ggplot(data = table_protein, aes(x = Amount, y = Gain, colour=Source, group=Source))+ 
  geom_point(aes(shape = Source,color = Source),size = 2) + 
  labs(y = "Weight Gained", x="Amount of table_protein", color ="Source of table_protein", shape ="Source of table_protein")
```


(b) Obtain the numerical summary for each treatment combination and factor levels separately. Report them here in a tabular form.

```{r echo=FALSE}
favstats(Gain ~ Source, data=table_protein)
favstats(Gain ~ Amount , data=table_protein)
favstats(Gain ~ Source|Amount, data=table_protein)
```


(c) Fit the two-factor factorial model and report the complete ANOVA table here. Do not report code here. The complete ANOVA table should have a row for each of the following: main effects of each treatment, two-factor interaction effects, error and total.

```{r echo=FALSE}
summary(model_protein)
```

(d) Based on the ANOVA table write your conclusion appropriately. Perform all the necessary tests and report the conclusion along with the p-value.

```{r echo=FALSE}
pairwise.t.test(table_protein$Gain, table_protein$Trt2, p.adj = "none") 

pairwise.t.test(table_protein$Gain, table_protein$Trt2, p.adj = "bonf")

plot(LSD.test(model_protein, trt = "Trt2", alpha = 0.05))

(LSD.test(model_protein, trt = "Trt2", alpha = 0.05)) 

TukeyHSD(model_protein, conf.level = 0.95)

ScheffeTest(model_protein, conf.level = 0.95)
```


(e) Provide the plots of residuals here. Do not report code here.

```{r echo=FALSE, out.width='25%'}
plot(model_protein, pch=16)
```


(f) Based on the residual plots, clearly explain whether assumptions in the model are satisfied or violated.


(g) Report the code here without output.


```{r , pressure, eval =F, size="footnotesize", tidy = T}
table_protein <- read_excel("Protein.xlsx")
glimpse(table_protein)

# Install and load ggplot2 package before using ggplot function #
ggplot(data = table_protein, aes(x = Amount, y = Gain, colour=Source, group=Source))+ 
  geom_point(aes(shape = Source,color = Source),size = 2) + 
  labs(y = "Weight Gained", x="Amount of table_protein", color ="Source of table_protein", shape ="Source of table_protein")

#Plots the Mean and 1SD error bars for each treatment group #
ggplot(data = table_protein, aes(x = Amount, y = Gain, colour=Source,shape = Source, group=Source)) +
  stat_summary() + labs(y= "Weight Gained", x="Amount of table_protein", color="Source", shape="Source")

# Install and load ggpubr package before using ggline function # 
ggline(data = table_protein, x = "Amount", y = "Gain", add = c("mean", "jitter"),shape= "Source",  color = "Source", linetype = "Source", ylab= "Weight Gained", xlab="Amount of table_protein")

#Load mosaic package before using favstats function#
favstats(Gain ~ Source, data=table_protein)
favstats(Gain ~ Amount , data=table_protein)
favstats(Gain ~ Source|Amount, data=table_protein)
#favstats(Gain ~ Source+Amount, data=table_protein)

#Create Categorical variables so that plot of residuals versus each treatment combination can be obtained using plot function with the fitted model later #
table_protein$Trt1 = as.factor(table_protein$Source)
table_protein$Trt2 = as.factor(table_protein$Amount)

model_protein <- aov(Gain ~ Trt1*Trt2, data=table_protein)
summary(model_protein)
plot(model_protein, pch=16)

# Pairwise comparisons using t tests with pooled Standard Deviation #
# The output gives a matrix of p values for each pair of treatments #
pairwise.t.test(table_protein$Gain, table_protein$Trt2, p.adj = "none") 

# Pairwise comparisons using t tests with pooled Standard Deviation and Bonferroni adjustment #
# The output gives a matrix of p values for each pair of treatments #
pairwise.t.test(table_protein$Gain, table_protein$Trt2, p.adj = "bonf")

#Install and load the agricolae package before running the LSD.test function below #
#p.adj option in the LSD.test function can be used to apply different adjustments to control error rates#
plot(LSD.test(model_protein, trt = "Trt2", alpha = 0.05))
#The treatments sharing the same letter on the plot are not different#

(LSD.test(model_protein, trt = "Trt2", alpha = 0.05)) # Using outer parentheses prints the output#

#Tukey's test to get observed difference in means, CI and p value#
TukeyHSD(model_protein, conf.level = 0.95)

# Scheffe's test to get observed difference in means, CI and p value #
# Install and load the DescTools package before using the ScheffeTest function #
ScheffeTest(model_protein, conf.level = 0.95)
```