---
title: 'USP654 Lab 3: Multiple Regression'
author: "Liming Wang"
date: "10/09/2017"
output: 
  html_document:
    toc: true
---

```{r setup, include=F}
knitr::opts_chunk$set(message=FALSE, warning=F, echo=TRUE)
options(width = 2000)
options(repos="https://cran.rstudio.com")
```

## Descriptive statistics among multiple variables

### Correlation Matrix

With base R function `cor`:
```{r}
#install.packages("tidyverse")
library(tidyverse)

mtcars[1:6] %>% cor
```


For fancier outputs, install the `corrr` package:

```{r, echo=TRUE}
install.packages("corrr")
library(corrr)

```

Half correlation matrix:
```{r, echo=TRUE}
mtcars %>% 
  correlate() %>% 
  shave() %>% 
  fashion()
```

Visulize correlation matrix:
```{r, echo=TRUE}
mtcars %>% 
  correlate() %>% 
  shave() %>% 
  rplot()
```

### Scatterplot Matrix

(Looking at the first 6 variables, as the figure becomes illegible when all variables are shown).
```{r, echo=TRUE}
mtcars[1:6] %>% 
  plot
```

Better looking version with `GGally` package
```{r, echo=TRUE}
install.packages("GGally")
```

```{r, echo=TRUE}
library(GGally)
ggpairs(data=mtcars[1:6])
```

## Multiple Regression
In R, run multiple regressions with `lm` (short for **l**inear **m**odel):

```{r, echo=TRUE}
lm(mpg ~ wt + cyl, data=mtcars)
```

## More detailed regression results

1. Pass the results from `lm()` to `summary()` for more detailed information:
```{r, echo=TRUE}
lm(mpg ~ wt + cyl, data=mtcars) %>%
  summary()
```

For better formatting of the results (pretty print), we can use the `texreg` package:
```{r, echo=TRUE}
## Install and load texreg package
install.packages("texreg")
library(texreg)

# Pretty print regression results on screen
screenreg(list(
    lm(mpg ~ wt, data=mtcars),
    lm(mpg ~ wt + cyl, data=mtcars)
))

# Save regression results to a html file
list(lm(mpg ~ wt, data=mtcars),
     lm(mpg ~ wt + cyl, data=mtcars)
     ) %>%
  htmlreg(file="output/lm_mpg-wt+cyl.html")
```

2. Confidence interval of coefficients
```{r, echo=TRUE}
lm(mpg ~ wt + cyl, data=mtcars) %>%
  confint()
```


## Visualize regression results

```{r, echo=TRUE}
ggplot(mtcars, aes(x=wt, y=mpg)) + geom_point() + geom_smooth(method = "lm", se = FALSE)
```

```{r, echo=TRUE}
ggplot(mtcars, aes(x=cyl, y=mpg)) + geom_point() + geom_smooth(method = "lm", se = FALSE)
```

```{r, echo=TRUE}
betas <- lm(mpg ~ wt + cyl, data=mtcars) %>% summary() %>% coef()
ggplot(mtcars, aes(x=wt, y=mpg)) + geom_point() +
  geom_abline(aes(slope=betas["wt", "Estimate"], intercept=betas["(Intercept)", "Estimate"]))

ggplot(mtcars, aes(x=cyl, y=mpg)) + geom_point() +
  geom_abline(aes(slope=betas["cyl", "Estimate"], intercept=betas["(Intercept)", "Estimate"]))
```

## Hypothesis testing of nested models

```{r, echo=TRUE}
## Install and load ggfortify package

lm_mpg_wt <- lm(mpg ~ wt, data=mtcars)
lm_mpg_wt.cyl <- lm(mpg ~ wt + cyl, data=mtcars)

anova(lm_mpg_wt, lm_mpg_wt.cyl)
```

## Tasks

1. ~~Create a new RStudio project and create directory structure following the recommendation in [Project Management With RStudio](http://swcarpentry.github.io/r-novice-gapminder/02-project-intro/)~~;
1. ~~Select a dataset from the [Tools and Datasets] section of the course website or use your own dataset, save the dataset in the data folder of your project~~;
1. ~~Famarlize yourself with your dataset with the data document~~;
1. ~~Create a script that imports your dataset into R and explores it with descriptive statistics and/or visulization~~;
1. Choose a dependent variable and a couple of independent variables and run regression models;
1. Create regression plots (scatter plot superimposed with a regression line) & diagnostic plots.

## Resources
- [Introduction to R on Data Camp](https://campus.datacamp.com/courses/free-introduction-to-r/chapter-1-intro-to-basics-1?ex=1): A self-instruction course covering R basics.
- [R coding basics on Software Carpentry](http://swcarpentry.github.io/r-novice-gapminder/), hands-on tutorial on R basics, lesson 4-11.
