---
title: 'STAT 510: Spatiotemporal Stats'
author: "Prof. Taylor-Rodriguez"
subtitle: Introduction and Exploring SPT Data
output:
  beamer_presentation:
    includes:
      in_header: "preamble.tex"
  ioslides_presentation: default
  slidy_presentation: default
---

```{r setup, echo=FALSE, message=F, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(STRbook)
library(ggplot2)
library(dplyr)
library(tidyr)
library(tidyverse)
library(spacetime)
library(lubridate)
```



# Spatio-Temporal Data Exploration


## Flavors of temporal and spatial data

As I mentionend last time, this field is BIG, and methodology has branched off according to the specific flavor of the data.


### In time

\bigskip
\bit
\item Time intervals type: regular or irregular? 

\item Is time discrete or continuous? 

\item Is the random event the time at which an event occurs?
\eit

## Flavors of temporal and spatial data

### In space:
Spatial data corresponds to either a fixed time-point or a temporal aggregation over multiple timepoints and can be:

\bigskip
\bit
\item \textcolor{orange}{Geostatistical:} point-level/coordinate observations, space is treated as a continuous surface over a given spatial domain


\item \textcolor{orange}{Areal:} space is discretized into a grid, polygons or small areas


\item \textcolor{orange}{Point-process:} the locations where events occur are themselves random, and each location can be complemented with attributes (called marks)
\eit


## Spatio-temporal data

Obviously spatio-temporal data will be defined by the combination of features in time and space the data have.

\bigskip
\textcolor{blue}{For this course, we'll focus on data that is discrete in time and \emph{geostatistical} or \emph{areal} in space}


<!-- ## Representation of spatio-temporal data -->

<!-- * Most often, spatio-temporal (ST) data is formated as \textsf{.csv} -->

<!--  -->
<!-- * ST data are represented throughout the book using the class definitions found in the \textsf{spacetime} \textsf{R} package -->

<!--  -->
<!-- \bdes -->
<!-- \item[time-wide] Columns in data are time points -->
<!-- \item[space-wide] Columns in data are spatial features -->
<!-- \item[long-format] Each row corresponds to a specific location and time-point combination (inefficient but easier to manipulate) -->
<!-- \edes -->

<!-- ## Representation of spatio-temporal data -->


# Exploratory Analysis of ST Data

## Empirical Spatial and Temporal Statistics


Denote by $\lrb{Z(s_i,t_j)}_{i=1,j=1}$ \textcolor{red}{the set of observations} at

\bigskip

* locations $\lrb{s_i:i=1,\ldots,m}$, and

* times  $\lrb{t_j:j=1,\ldots,T}$


\bigskip
\begin{block}{The empirical spatial means vector is}
$$\hat{\bmu}_{z,s} = \left[\bmat \hat{\mu}_{z,s}(s_1)\\ \vdots \\ \hat{\mu}_{z,s}(s_m) \emat\right]= \left[\bmat \frac{1}{T}\sum_{j=1}^T Z(s_1,t_j)\\ \vdots \\ \frac{1}{T}\sum_{j=1}^T Z(s_m,t_j) \emat\right]=\frac{1}{T} \sum_{j=1}^T \bZ_{t_j}$$
\end{block}


## Empirical Spatial and Temporal Statistics

```{r loadNOAA, echo=F, eval=T}
locs <- read.table(system.file("extdata",
                               "Stationinfo.dat", 
                               package = "STRbook"),
col.names = c("id", "lat", "lon"))
Times <- read.table(system.file("extdata",
                                "Times_1990.dat",
                                package = "STRbook"),
col.names = c("julian", "year", "month", "day"))
Tmax <- read.table(system.file("extdata",
                               "Tmax_1990.dat", 
                               package = "STRbook"))
names(Tmax) <- locs$id
Tmax <- bind_cols(Times,Tmax)
Tmax_long <- gather(Tmax, id, z, -julian, -year, -month,
                    -day) %>% 
  filter(!(z<=-9998))
```

### Empirical Spatial Means (NOAA Data)

\bigskip
```{r spMeans, echo=F, eval=T, message=F, fig.height=2.5, fig.width=3, fig.align='center'}
Tmax_long %>% 
  group_by(id) %>% #,year) %>% 
  summarise(z=mean(z)) %>% 
  mutate(id=as.numeric(id)) %>% 
  left_join(locs,by="id") %>% 
  # filter(year<1993) %>% 
  ggplot() + 
  geom_point(aes(x=lon,y=lat,colour=z),size=2) +
  col_scale(name = "degF") + 
  xlab("Longitude (deg)") + 
  ylab("Latitude (deg)") + 
  geom_path(data = map_data("state"),# plot points
            aes(x = long, y = lat, group = group)) +
  # facet_grid(.~year) + 
  coord_fixed(xlim = c(-105, -75),
              ylim= c(25,50)) +
  theme_bw()
```

\smallskip
\scriptsize{In these data, stations have different number of measurements, so we use instead $\hat{\mu}_{z,s}(s_i)=\frac{1}{T_i}\sum_{j=1}^{T_i} Z(s_i,t_j)$}

## Empirical Spatial and Temporal Statistics

Conversely, one could average over space to get the *empirical temporal mean*, given by $$\hat{\mu}_{z,t}(t_j)=\frac{1}{m}\sum_{i=1}^{m} Z(s_i,t_j)$$

```{r tsMeans, echo=F, eval=T, message=F, fig.height=2, fig.width=4, fig.align='center'}
Tmax_tsmu <- Tmax_long %>% 
  group_by(year,month,day) %>% #,year) %>% 
  summarise(z=mean(z)) %>% 
  mutate(year=as.character(year),
         month=as.character(month),
         day=as.character(day)) %>% 
  unite("date",year,month,day) %>% 
  mutate(date=ymd(date)) %>% 
  filter(date<ymd("19920101"))

rndsample <- sample(Tmax_long$id,100,replace=F)
Tmax_long %>% 
  filter(id%in%rndsample) %>% 
  group_by(id,year,month,day) %>% #,year) %>% 
  mutate(year=as.character(year),
         month=as.character(month),
         day=as.character(day)) %>% 
  unite("date",year,month,day) %>% 
  mutate(date=ymd(date)) %>% 
  filter(date<ymd("19920101")) %>% 
  ggplot() + 
  geom_line(aes(x=date,y=z,group=id),
            color="cornflowerblue",
            alpha=0.2,size=.1) +
  geom_line(data=Tmax_tsmu,
            aes(x=date,y=z),size=0.2) +
  ylab("Max temp (degF)") + 
  xlab("date") + 
  theme_bw()
```

## Empirical Spatial and Temporal Statistics

\begin{block}{For 2 locations, the lag-$\tau$ empirical spatial covariance is}

\scriptsize{$$\hat{C}_z^{(\tau)}(s_i,s_k)=\frac{1}{T-\tau}\sum_{j=\tau+1}^{T} (Z(s_i,t_j)-\hat{\mu}_{z,s}(s_i))(Z(s_k,t_j-\tau)-\hat{\mu}_{z,s}(s_k))$$}
\normalsize
for $\tau=0,1,\ldots,T-1.$
\end{block}


\begin{block}{For all locations, the lag-$\tau$ empirical spatial covariance matrix is}
\scriptsize{$$\hat{\bC}_z^{(\tau)}=\frac{1}{T-\tau}\sum_{j=\tau+1}^{T} (\bZ_{t_j}-\hat{\bmu}_{z,s})(\bZ_{t_j-\tau}-\hat{\bmu}_{z,s})'$$}
\normalsize
where $\bZ_{t_j}=(Z(s_1,t_j),Z(s_2,t_j),\ldots,Z(s_m,t_j))'$, and $\hat{\bmu}_{z,s}$ is defined analogously.
\end{block}


## Empirical Spatial and Temporal Statistics

\begin{block}{Between two different datasets, the lag-$\tau$ empirical cross-covariance matrix is}
\scriptsize{$$\hat{\bC}_{z,x}^{(\tau)}=\frac{1}{T-\tau}\sum_{j=\tau+1}^{T} (\bZ_{t_j}-\hat{\bmu}_{z,s})(\bX_{t_j-\tau}-\hat{\bmu}_{x,s})'$$}
\normalsize
Can be used to characterize ST dependence in two variables
\end{block}


## Measures of Spatio-Temporal dependence

\begin{block}{Empirical ST Covariogram}

We want to describe covariability in the ST data as a function of temporal lags and space. Assume that


\bigskip
\bit
\item the first moment depends on space but not on time


\bigskip

\item the second moment depends on lag differences in space and time
\eit
\end{block}

Then the ST covariogram for spatial and temporal lags $\bh$ and $\tau$, respectively, is


\scriptsize{$$\hat{\bC}_{z}(\bh,\tau)=\frac{1}{|N_s(\bh)|}\frac{1}{|N_t(\tau)|}\sum_{s_i,s_k\in N_s(\bh)} \sum_{t_j,t_\ell\in N_t(\tau)}(Z(s_i,t_j)-\hat{\mu}_{z,s}(s_i))(Z(s_k,t_\ell)-\hat{\mu}_{z,s}(s_k))$$}


## Measures of Spatio-Temporal dependence

\begin{block}{Empirical ST Semivariogram}
$$\gamma_z(s_i,s_k;t_j,t_\ell)=\frac{1}{2}\text{var}(Z(s_i;t_j)-Z(s_k;t_\ell)).$$


If the covariance is only in terms of displacement in space and differences in time, then
\begin{eqnarray*}
\gamma_z(s_i,s_k;t_j,t_\ell)&=&\text{var}(Z(s+\bh;t+\tau)-Z(s;t))\\
&=&C_z(\0,0)-C_z(\bh,\tau).
\end{eqnarray*}
with $\bh=s_i-s_k$ and $\tau=t_j-t_\ell$.
\end{block}
\scriptsize{\textcolor{red}{Note:} This last equation doesn't hold even if $\exists \gamma_z(\bh;\tau)$ if there is no $C_z(\bh;\tau)$ that is stationary.}

## In class-problem

Show, under the conditions previously assumed, that

\begin{eqnarray*}
\text{var}(Z(s+\bh;t+\tau)-Z(s;t))&=&C_z(\0,0)-C_z(\bh,\tau).
\end{eqnarray*}

## Measures of Spatio-Temporal dependence

\begin{block}{ST Semivariogram}
If $C_z(\bh;\tau)$ is well-defined, the the semivariogram depends on

\bigskip
\bdes
\item[Nugget effect] $$\underset{\stackrel{\bh\rightarrow \0}{ \tau\rightarrow 0}}{\lim}\gamma_z(\bh;\tau)$$ 


\item[Sill:] $$\underset{\stackrel{\bh\rightarrow \infty}{ \tau\rightarrow\infty}}{\lim}\gamma_z(\bh;\tau)$$ 


\item[Partial sill:] Sill minus Nugget
\edes
\end{block}

\scriptsize
\textcolor{red}{Note:} Before fitting a semivariogram, non-stationary behavior in $C_z(\bh;\tau)$ may be avoided by fitting (and removing) linear and quadratic trends in ST coordinates.

## Measures of Spatio-Temporal dependence

\begin{block}{Empirical ST Semivariogram}

\bigskip
Assuming that $\mu_{z,s}$ is constant across all locations, then
\scriptsize
\begin{eqnarray*}
\hat{\gamma}_z(\bh;\tau)&=&\hat{C}_z(\0;0)-\hat{C}_z(\bh;\tau)\\
&=& \frac{1}{|N_s(\bh)|}\frac{1}{|N_t(\tau)|}\sum_{s_i,s_k\in N_s(\bh)} \sum_{t_j,t_\ell\in N_t(\tau)}(Z(s+\bh,t+\tau)-Z(s,t))^2
\end{eqnarray*}
\end{block}

```{r esemivar, echo=F, eval=T, message=F, warning=F,fig.height=2.2, fig.width=3, fig.align='center'}
data("STObj3", package = "STRbook")
STObj4 <- STObj3[, "1993-07-01::1993-07-31"]

vv <- gstat::variogram(object = z~1 + lat,
                       data = STObj4,
                       width = 80,
                       cutoff = 1000,
                       tlags = 0.01:6.01)
plot(vv)
```


## Empirical Orthogonal Functions

\begin{block}{}

\bigskip
\bit
\item Help identify spatial patterns in ST data

\bigskip
\item Can be used for dimension reduction

\bigskip
\item Are ST manifestation of Principal Components
\eit
\end{block}

## Interlude: A Review of PCA

* Assume iid random $\bx_i=(x_{i1}, x_{i2},\ldots, x_{ip})'$ with $p\times p$ covariance matrix $\text{var}(\bx_i)=\bC_x$.


* Suppose we want to build variables $$a_{ik}=w_{k 1} x_{i1}+ w_{k 2}x_{i2} + \cdots + w_{kp}x_{ip}$$ for $k=1, \ldots, p$, that are *orthogonal* to each other.


* How should the $w_{kj}$'s be chosen?


## Interlude: A Review of PCA

Using spectral decomposition: $$\bC_x = W \Lambda_x W',$$ with $W W' = W'W=\text{I}_p$ and $\Lambda=\text{diag}(\lambda_1,\lambda_2,\cdots, \lambda_p)$, with $$\lambda_1\geq \lambda_2\geq \cdots \geq \lambda_p>0.$$ 

 
Dropping the $i$
$$\ba = W' \bx = \left[\bmat \bw_{1}' \bx\\ \vdots \\ \bw_{p}'\bx \emat\right] = \left[\bmat a_1\\ \vdots \\ a_{p} \emat\right],$$

Note that $\text{var}(\ba)= W'\text{var}(\bx) W = W'(W \Lambda_x W') W = \Lambda_x.$


## Interlude: A Review of PCA

\bit
\item Since $\Lambda_x$ is diagonal then the $a_k$'s are independent (orthogonal to each other) 


\bigskip
\item Furthermore, note that
\eit

\vspace{-1cm}
\begin{eqnarray*}
\text{var}(x_1)+\cdots+\text{var}(x_p) &=&\text{trace}(C_x) \\
&=& \lambda_1+\cdots+\lambda_p \\
&=&\text{var}(a_1)+\cdots+\text{var}(a_p)
\end{eqnarray*}


\bit
\item Lastly, since $\lambda_1\geq \ldots \geq \lambda_p.$

$a_1$ captures the largest amount of variability in $\bx$, $a_2$ the second largest, etc.
\eit

\scriptsize
\textcolor{red}{Note:}In practice $\bC_x$ is unknown, so we use the estimator $\hat{\bC}_x$ instead.


## Empirical Orthogonal Functions

\bit
\item Replace $\bZ_{t_j}$ for $\bx_i$, and use the lag-0 empirical spatial covariance matrix $\hat{\bC}_z^{(0)}$


\bigskip
\item Spectrally decompose $\hat{\bC}_z^{(0)}$ into $\Psi \Lambda \Psi,$
where 
\eit

\vspace{-0.7cm}
\begin{eqnarray*}
\Psi_{m\times m}&=&(\bpsi_1, \ldots, \bpsi_m)'\;\text{with}\\
\bpsi_k &=& (\psi_k(s_1), \ldots, \psi_k(s_m))'.
\end{eqnarray*}


\bit
\item The $k$th principal component times series is $$\ba_k=\left[\bmat a_k(t_1)\\ a_k(t_2)\\ \vdots\\ a_k(t_T) \emat\right]=\left[\bmat \bpsi_k' \bZ_{t_1}\\ \bpsi_k' \bZ_{t_2}\\ \vdots\\ \bpsi_k' \bZ_{t_T} \emat\right].$$
\eit


## Empirical Orthogonal Functions

The $\bpsi_k$'s ($k=1,\ldots,m$) are known as the Empirical Orthogonal Functions (EOFs).  Main uses are:

\bigskip

\benum
\item Identify salient spatial patterns in ST data

\bigskip

\item Reduce the dimensionality of spatial or ST random effects (more on this later)
\eenum


## Empirical Orthogonal Functions

To do dimension reduction, we need to decide the number of EOFs to use. How?

\bigskip
Here are some possibilities


\benum
\item Use the $\lambda_k$'s to determine the proportion of overall variance to retain


\bigskip
\item Use scree plot and find after including how many EOFs the variance contribution flattens out


\bigskip
\item Identify the *significant* EOFs by obtaining EOFs for several spatially permuted versions of the data.  See where the tue scree plot intersects with those of the permuted datasets
\eenum


## Empirical Orthogonal Functions

\begin{block}{A more efficient method to obtain EOF's}
Note that letting

$$\tilde{\bZ}=\frac{1}{\sqrt{T-1}}(\bZ-\1_T \hat{\bmu}_{z,s}')$$ 
Implies that


$$\Longrightarrow \hat{\bC}_z^{(0)}=\tilde{\bZ}'\tilde{\bZ}  =\bPsi \bLambda \bPsi'$$


Also, considering the SVD $\tilde{\bZ}=\bU \bD \bV'$, we have $$\hat{\bC}_z^{(0)}=(\bU \bD \bV')'(\bU \bD \bV')=\underbrace{\bV}_{\Psi}\underbrace{\bD'\bD}_{\Lambda}\underbrace{\bV'}_{\Psi'}$$
This equality implies that SVD can be used (and is recommended) to obtain the EOF's.
\end{block}

## ST Canonical Correlation Analysis (CCA)


\bit
\item \textcolor{red}{\bf Goal:} create 2 new sets of variables, which arise as linear combinations of two multivariate (mvt) datasets and that are maximally correlated

\bigskip
\item  The first few pairs of canonical variables are very similar (since maximaly correlated)

\bigskip
\item  The spatial patterns in the weights show the areas in space that are most responsible for the high correlations
\eit



\bigskip

In the ST case, each \textcolor{blue}{\emph{spatial location}} corresponds to a \textcolor{blue}{\emph{variable}}, and each \textcolor{orange}{\emph{time-point}} corresponds to a \textcolor{orange}{\emph{sample}}


## ST Canonical Correlation Analysis (CCA)

This, from two datasets denoted by
\begin{eqnarray*}
\lrb{\bZ_{t_t}=(Z(s_1;t_j),\ldots,Z(s_m;t_j))':j=1,\ldots,T}&&\\
\lrb{\bX_{t_t}=(X(r_1;t_j),\ldots,X(r_n;t_j))':j=1,\ldots,T}
\end{eqnarray*}
 
ST-CCA, for $k=1,2,\ldots,\min\lrb{m,n}$, seeks to generate variables

 \begin{eqnarray*}
a_k(t_j)&=&\sum_{i=1}^m \xi_{ik} Z(s_i;t_j)\;=\;\mbs{\xi}_k' \bZ_{t_j}\\
b_k(t_j)&=&\sum_{\ell=1}^n \psi_{\ell k} X(r_{\ell};t_j)\;=\;\mbs{\psi}_k' \bX_{t_j}
\end{eqnarray*}


with weights $\mbs{\xi}_k$ and $\bpsi_k$ a.k.a. \emph{canonical variables}
 

## In-class Problem

The $k$th canonical correlation $r_k$ is given by $$r_k=\text{corr}(a_k,b_k)=\frac{\mbs{\xi}_k' \bC_{z,x}^{(0)} \bpsi_k}{(\mbs{\xi}_k' \bC_{z}^{(0)} \mbs{\xi}_k)^{1/2} (\mbs{\psi}_k' \bC_{x}^{(0)} \mbs{\psi}_k)^{1/2}}$$
show this result.

## ST Canonical Correlation Analysis (CCA)

### Calculating ST-CCA in practice

\scriptsize
\begin{eqnarray*}
r_k=\text{corr}(a_k,b_k) &=& \frac{\mbs{\xi}_k' \bC_{z,x}^{(0)} \bpsi_k}{(\mbs{\xi}_k' \bC_{z}^{(0)} \mbs{\xi}_k)^{1/2} (\mbs{\psi}_k' \bC_{x}^{(0)} \mbs{\psi}_k)^{1/2}}\\
&=&\frac{\overbrace{\mbs{\xi}_k'(\hat{\bC}_z^{(0)})^{1/2}}^{\equiv\tilde{\mbs{\xi}}_k'}(\hat{\bC}_z^{(0)})^{-1/2} \bC_{z,x}^{(0)} (\hat{\bC}_x^{(0)})^{-1/2}\overbrace{(\hat{\bC}_x^{(0)})^{1/2}\bpsi_k}^{^{\equiv\tilde{\mbs{\psi}}_k}} }{(\mbs{\xi}_k' (\hat{\bC}_z^{(0)})^{1/2}(\hat{\bC}_z^{(0)})^{1/2} \mbs{\xi}_k)^{1/2} (\mbs{\psi}_k' (\hat{\bC}_x^{(0)})^{1/2}(\hat{\bC}_x^{(0)})^{1/2} \mbs{\psi}_k)^{1/2}} \\
&=&\frac{\tilde{\mbs{\xi}}_k'(\hat{\bC}_z^{(0)})^{-1/2} \bC_{z,x}^{(0)} (\hat{\bC}_x^{(0)})^{-1/2}\tilde{\bpsi}_k }{(\tilde{\mbs{\xi}}_k'\tilde{\mbs{\xi}}_k)^{1/2} (\tilde{\bpsi}_k'\tilde{\bpsi}_k)^{1/2}}
\end{eqnarray*}


\normalsize
It is well-known that from the SVD of $$(\hat{\bC}_z^{(0)})^{-1/2} \bC_{z,x}^{(0)} (\hat{\bC}_x^{(0)})^{-1/2},$$  $r_1^2$ corresponds to the largest singular value, and $\tilde{\mbs{\xi}}_1$ and $\tilde{\bpsi}_1$ are the first left and right singular vectors, respectively.

## ST Canonical Correlation Analysis (CCA)

### Calculating ST-CCA in practice

\scriptsize
```{r echo=T}
data("SSTlandmask", package = "STRbook")
data("SSTlonlat", package = "STRbook")
data("SSTdata", package = "STRbook")

delete_rows <- which(SSTlandmask == 1)
SSTdata <- SSTdata[-delete_rows, 1:396]
 
## Put data into space-wide form
Z <- t(SSTdata)
```

\vfill

## ST Canonical Correlation Analysis (CCA)

### Calculating ST-CCA in practice

\scriptsize
```{r echo=T}
## First find the matrix we need to subtract:
spat_mean <- apply(SSTdata, 1, mean)
nT <- ncol(SSTdata)

## Then subtract and standardize:
Zspat_detrend <- Z - outer(rep(1, nT), spat_mean)
Zt <- 1/sqrt(nT - 1)*Zspat_detrend

## get the SVD
E <- svd(Zt)
```
\vfill


## ST Canonical Correlation Analysis (CCA)

### Calculating ST-CCA in practice

\scriptsize
```{r echo=T,eval=F, fig.height=2.3, fig.width=3.5, fig.align="center"}
nEOF <- 10
EOFset1 <- E$u[1:(nT-7), 1:nEOF] * sqrt(nT - 1)
EOFset2 <- E$u[8:nT, 1:nEOF] * sqrt(nT - 1)

cc <- cancor(EOFset1, EOFset2) # compute CCA 

options(digits = 3) 
CCA_df <- data.frame(t = 1:(nT - 7),
                     CCAvar1 = (EOFset1 %*% cc$xcoef[,1])[,1],
                     CCAvar2 = (EOFset2 %*% cc$ycoef[,1])[,1])
t_breaks <- seq(1, nT, by = 60) # breaks for x-labels 
year_breaks <- seq(1970,2002,by=5) # labels for x-axis 
ggplot(CCA_df) +
  geom_line(aes(t, CCAvar1), col = "dark blue") + 
  geom_line(aes(t, CCAvar2), col = "dark red") + 
  scale_x_continuous(breaks = t_breaks, labels = year_breaks) + 
  ylab("CCA variables") + xlab("Year") + theme_bw()
```
\vfill


## ST Canonical Correlation Analysis (CCA)

### Calculating ST-CCA in practice

\scriptsize
```{r echo=F,eval=T, fig.height=2.3, fig.width=3.5, fig.align="left"}
nEOF <- 10
EOFset1 <- E$u[1:(nT-7), 1:nEOF] * sqrt(nT - 1)
EOFset2 <- E$u[8:nT, 1:nEOF] * sqrt(nT - 1)

cc <- cancor(EOFset1, EOFset2) # compute CCA 

options(digits = 3) 
CCA_df <- data.frame(t = 1:(nT - 7),
                     CCAvar1 = (EOFset1 %*% cc$xcoef[,1])[,1],
                     CCAvar2 = (EOFset2 %*% cc$ycoef[,1])[,1])
t_breaks <- seq(1, nT, by = 60) # breaks for x-labels 
year_breaks <- seq(1970,2002,by=5) # labels for x-axis 
ggplot(CCA_df) +
  geom_line(aes(t, CCAvar1), col = "dark blue") + 
  geom_line(aes(t, CCAvar2), col = "dark red") + 
  scale_x_continuous(breaks = t_breaks, labels = year_breaks) + 
  ylab("CCA variables") + xlab("Year") + theme_bw()
```

\vfill

<!-- ## Visualizing spatio-temporal data -->

<!-- ### Time-Series Plots -->


<!-- ## Visualizing spatio-temporal data -->

<!-- ### Hovlmoller Plots -->


<!-- ## Visualizing spatio-temporal data -->

<!-- ### Other Visualization Options -->

