---
title: "nimble_practise"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set( eval =T,echo = T,cache=T,collapse=T, out.width='50%',fig.show='hold',message=F,warning=F)
# options(scipen=10)
# options(digits=4)
if (!require(pacman)) install.packages("pacman")
pacman::p_load(nimble) 
# readxl,ggplot2,GGally,tidyverse, MASS,car,olsrr,lmtest,GAD,FrF2,emmeans
library(nimble, warn.conflicts = FALSE)
```

```{r,eval=F,echo=F}
writeLines('PATH="${RTOOLS40_HOME}\\usr\\bin;${PATH}"', con = "~/.Renviron")
Sys.which("make")
pkgbuild::find_rtools(debug = TRUE)
```

#  {.tabset .tabset-fade .tabset-pills}

## Var-Comp

Variance components model

Building a model from BUGS code in R using NIMBLE

See <https://r-nimble.org/examples>.


```{r}
dyesCode <- nimbleCode({
# Model
   for (i in 1:BATCHES) {
      for (j in 1:SAMPLES) {
         y[i,j] ~ dnorm(mu[i], sd = sigma.within);
      }
      mu[i] ~ dnorm(theta, sd = sigma.between);
   }
   
# Priors
   theta ~ dnorm(0.0, 1.0E-10);
   sigma.within ~ dunif(0, 100)
   sigma.between ~ dunif(0, 100)
})
```

```{r}
classicDyesModel <- readBUGSmodel('dyes', dir = getBUGSexampleDir('dyes'))
```

```{r}
dyesModel <- nimbleModel(dyesCode, constants = list(BATCHES = 6, SAMPLES = 5))
```

```{r}
data <- matrix(c(1545, 1540, 1595, 1445, 1595, 1520, 1440, 1555, 1550, 
1440, 1630, 1455, 1440, 1490, 1605, 1595, 1515, 1450, 1520, 1560, 
1510, 1465, 1635, 1480, 1580, 1495, 1560, 1545, 1625, 1445), nrow = 6)

dyesModel$setData(list(y = data))
dyesModel$y
```

```{r}
dyesModel$theta <- 1500
dyesModel$mu <- rnorm(6, 1500, 50)
dyesModel$sigma.within <- 20
dyesModel$sigma.between <- 20
dyesModel$y[1,]
dyesModel$theta
```

- Calculate log probability densities for part or all of the model

```{r}
## arbitrary example
dyesModel$calculate(c('theta', 'mu[1:6]', 'y[,2]'))
```

- Simulate part or all of the model

```{r}
dyesModel$mu
dyesModel$simulate(c('mu[1:3]'))
dyesModel$mu
```

- Query the model’s relationships

```{r}
## arbitrary example
dyesModel$getDependencies(c('theta', 'mu[3]'))
```

- Plot the model graph

```{r}
library(igraph)
plot(dyesModel$getGraph())
```


- Compile the model

```{r}
compiled_dyesModel <- compileNimble(dyesModel)
```

```{r}
compiled_dyesModel$theta <- 1450
compiled_dyesModel$calculate() ## all nodes by default
```


## Linear Reg

```{r}
set.seed(1)
p <- 15    # number of explanatory variables
n <- 100   # number of observations
X <- matrix(rnorm(p*n), nrow = n, ncol = p) # explanatory variables
true_betas <- c(c(0.1, 0.2, 0.3, 0.4, 0.5), rep(0, p-5)) # coefficients
sigma <- 1
y <- rnorm(n, X %*% true_betas, sigma)
```


### Manual specification

```{r}
code <- nimbleCode({
  beta0 ~ dnorm(0, sd = 100)
  beta1 ~ dnorm(0, sd = 100)
  beta2 ~ dnorm(0, sd = 100)
  sigma ~ dunif(0, 100)        # prior for variance components based on Gelman (2006)
  for(i in 1:n) {
    y[i] ~ dnorm(beta0 + beta1*x1[i] + beta2*x2[i], sd = sigma) # manual entry of linear predictors
  }
})

## extract data for two predictors and center for better MCMC performance
x1 <- X[,1] - mean(X[,1])
x2 <- X[,2] - mean(X[,2])

constants <- list(n = n, x1 = x1, x2 = x2)
data <- list(y = y)
inits <- list(beta0 = mean(y), beta1 = 0, beta2 = 0, sigma = 1)
model <- nimbleModel(code, constants = constants, data = data, inits = inits) # build model
```



```{r}
mcmcConf <- configureMCMC(model) # assign default samplers to nodes
```

```{r}
mcmcConf$printSamplers() # look at default sampler assignments
```

```{r}
mcmc.out <- nimbleMCMC(code = code, constants = constants,
                       data = data, inits = inits,
                       nchains = 2, niter = 1000,
                       summary = TRUE, WAIC = T,
                       monitors = c('beta0','beta1','beta2','sigma')) # 

names(mcmc.out)
```

```{r}
mcmc.out$summary
mcmc.out$WAIC 
```



### Using `inprod`

```{r}
code <- nimbleCode({
  beta0 ~ dnorm(0, sd = 100)
  for(k in 1:p)
    beta[k] ~ dnorm(0, sd = 100)
  sigma ~ dunif(0, 100)  # prior for variance components based on Gelman (2006)
  for(i in 1:n) {
    y[i] ~ dnorm(beta0 + inprod(beta[1:p], x[i, 1:p]), sd = sigma)
  }
})

X <- sweep(X, 2, colMeans(X))  # center for better MCMC performance

constants <- list(n = n, p = p, x = X)
data <- list(y = y)
inits <- list(beta0 = mean(y), beta = rep(0, p), sigma = 0.5)
model <- nimbleModel(code, constants = constants, data = data, inits = inits)
```


```{r}
mcmcConf <- configureMCMC(model)
mcmcConf$printSamplers() ## Look at sampler assignments in this case.
```

```{r}
mcmc.out <- nimbleMCMC(code = code, constants = constants,
                       data = data, inits = inits,
                       nchains = 2, niter = 1000,
                       summary = TRUE, WAIC = T,
                       monitors = c('beta0','beta','sigma')) # 

names(mcmc.out)
```

```{r}
mcmc.out$summary
mcmc.out$WAIC 
```

### Using matrix algebra

```{r}
code <- nimbleCode({
  beta0 ~ dnorm(0, sd = 100)
  beta[1:p] ~ dmnorm(zeros[1:p], omega[1:p, 1:p])
  sigma ~ dunif(0, 100)  # prior for variance components based on Gelman (2006)
  for(i in 1:n) {
    y[i] ~ dnorm(beta0 + (beta[1:p] %*% x[i, 1:p])[1,1], sd = sigma)
  }
})

constants <- list(n = n, p = p, x = X, zeros = rep(0, p), omega = 0.0001 * diag(p))
data <- list(y = y)
inits <- list(beta0 = mean(y), beta = rep(0, p), sigma = 0.5)
model <- nimbleModel(code, constants = constants, data = data, inits = inits)
```


```{r}
mcmcConf <- configureMCMC(model)
mcmcConf$printSamplers()
```

```{r}
mcmc.out <- nimbleMCMC(code = code, constants = constants,
                       data = data, inits = inits,
                       nchains = 2, niter = 1000,
                       summary = TRUE, WAIC = T,
                       monitors = c('beta0','beta','sigma')) # 

names(mcmc.out)
```

```{r}
mcmc.out$summary
mcmc.out$WAIC 
```

- Compute all n values of the linear predictor in a single vactorized step

```{r}
code <- nimbleCode({
     beta0 ~ dnorm(0, sd = 100)
     beta[1:p] ~ dmnorm(zeros[1:p], omega[1:p, 1:p])
     sigma ~ dunif(0, 100)  # prior for variance components based on Gelman (2006)
     linpred[1:n] <- (x[1:n, 1:p] %*% beta[1:p])[1:n,1]
     for(i in 1:n) {
           y[i] ~ dnorm(beta0 + linpred[i], sd = sigma)
       }
})

constants <- list(n = n, p = p, x = X, zeros = rep(0, p), omega = 0.0001 * diag(p))
data <- list(y = y)
inits <- list(beta0 = mean(y), beta = rep(0, p), sigma = 0.5)
model <- nimbleModel(code, constants = constants, data = data, inits = inits)
```

```{r}
mcmcConf <- configureMCMC(model)
mcmcConf$printSamplers()
```

```{r}
mcmc.out <- nimbleMCMC(code = code, constants = constants,
                       data = data, inits = inits,
                       nchains = 2, niter = 1000,
                       summary = TRUE, WAIC = T,
                       monitors = c('beta0','beta','sigma')) # 

names(mcmc.out)
```

```{r}
mcmc.out$summary
mcmc.out$WAIC 
```

- An equivalent model simply calculates the probability density of all y[i] elements together

```{r}
dnorm_vec <- nimbleFunction( ## Define the distribution
    run = function(x = double(1), mean = double(1), sd = double(0), log = integer(0, default = 0)) {
        returnType(double(0))
        logProb <- sum(dnorm(x, mean, sd, log = TRUE))
        if(log) return(logProb)
        else return(exp(logProb)) 
    })

rnorm_vec <- nimbleFunction( ## Define a simulation function, optionally.
    run = function(n = integer(0), mean = double(1), sd = double(0)) {
        returnType(double(1))
        if(n != 1) print("rnorm_vec only allows n = 1; using n = 1.")
        smp <- rnorm(n, mean, sd)
        return(smp)
    })

code <- nimbleCode({
     beta0 ~ dnorm(0, sd = 100)
     beta[1:p] ~ dmnorm(zeros[1:p], omega[1:p, 1:p])
     sigma ~ dunif(0, 100)  # prior for variance components based on Gelman (2006)
     linpred[1:n] <- beta0 + x[1:n, 1:p] %*% beta[1:p]
     y[1:n] ~ dnorm_vec(linpred[1:n], sigma)
})

constants <- list(n = n, p = p, x = X, zeros = rep(0, p), omega = 0.0001 * diag(p))
data <- list(y = y)
inits <- list(beta0 = mean(y), beta = rep(0, p), sigma = 0.5)
model <- nimbleModel(code, constants = constants, data = data, inits = inits)
```

```{r}
mcmcConf <- configureMCMC(model)
mcmcConf$printSamplers() ## Look at sampler assignments
model$getNodeNames()     ## Look at node names in the re-written model
```

```{r}
mcmc.out <- nimbleMCMC(code = code, constants = constants,
                       data = data, inits = inits,
                       nchains = 2, niter = 1000,
                       summary = TRUE, WAIC = T,
                       monitors = c('beta0','beta','sigma')) # 

names(mcmc.out)
```

```{r}
mcmc.out$summary
mcmc.out$WAIC 
```




## Logistic Reg

MCMC for logistic regression with random effects

```{r}
## define the model
code <- nimbleCode({
    beta0 ~ dnorm(0, sd = 10000)
    beta1 ~ dnorm(0, sd = 10000)
    sigma_RE ~ dunif(0, 1000)
    for (i in 1:N) {
        beta2[i] ~ dnorm(0, sd = sigma_RE)
        logit(p[i]) <- beta0 + beta1 * x[i] + beta2[i]
        r[i] ~ dbin(p[i], n[i])
    }
})

## constants, data, and initial values
constants <- list(N = 10)

data <- list(
    r = c(10, 23, 23, 26, 17, 5, 53, 55, 32, 46),
    n = c(39, 62, 81, 51, 39, 6, 74, 72, 51, 79),
    x = c(0,  0,  0,  0,  0,  1, 1,  1,  1,  1)
)

inits <- list(beta0 = 0, beta1 = 0, sigma_RE = 1)

## create the model object
glmmModel <- nimbleModel(code = code, constants = constants, data = data, 
                         inits = inits, check = FALSE)
```




- Default MCMC Algorithm

```{r}
glmmMCMC <- buildMCMC(glmmModel)
```




- Compile the model and MCMC algorithm

```{r}
CglmmModel <- compileNimble(glmmModel)
```


```{r}
mcmc.out <- nimbleMCMC(code = code, constants = constants,
                       data = data, inits = inits,
                       nchains = 2, niter = 1000,
                       summary = TRUE, WAIC = F,
                       monitors = c('beta0','beta1','beta2')) # 

names(mcmc.out)
```

```{r}
mcmc.out$summary
mcmc.out$WAIC 
```


```{r,eval=F}
CglmmMCMC <- compileNimble(glmmMCMC, project = glmmModel)
```


- Execute MCMC algorithm and extract samples

```{r,eval=F}
samples <- runMCMC(CglmmMCMC, niter = 1000)
```

- Customize an MCMC Algorithm

```{r}
glmmModel <- nimbleModel(code = code, constants = constants, 
                         data = data, inits = inits)
```

```{r}
spec <- configureMCMC(glmmModel, nodes = NULL)
spec$addSampler(type = 'slice', target = 'beta0')
spec$addSampler(type = 'slice', target = 'beta1')
spec$addSampler(type = 'RW', target = 'sigma_RE')
spec$addSampler(type = 'RW_block', target = 'beta2[1:10]')
customGlmmMCMC <- buildMCMC(spec)
```

```{r}
CglmmModel <- compileNimble(glmmModel)
CglmmMCMC <- compileNimble(customGlmmMCMC, project = glmmModel)
samples <- runMCMC(CglmmMCMC, niter = 1000)
```


## Chap2-intro

Chapter 2 Lightning introduction



See <https://r-nimble.org/html_manual/cha-welcome-nimble.html>.

### 2.2 Creating a model

```{r}
pumpCode <- nimbleCode({ 
  for (i in 1:N){
      theta[i] ~ dgamma(alpha,beta)
      lambda[i] <- theta[i]*t[i]
      x[i] ~ dpois(lambda[i])
  }
  alpha ~ dexp(1.0) # 
  beta ~ dgamma(0.1,1.0)
})

pumpConsts <- list(N = 10,
                   t = c(94.3, 15.7, 62.9, 126, 5.24,31.4, 1.05, 1.05, 2.1, 10.5))

pumpData <- list(x = c(5, 1, 5, 14, 3, 19, 1, 1, 4, 22))

pumpInits <- list(alpha = 1 ,beta = 1, # alpha =1.802
                  theta =rep(0.1, pumpConsts$N) ) #x/t
```

```{r}
pump <- nimbleModel(code = pumpCode, name = "pump", constants = pumpConsts,
                    data = pumpData, inits = pumpInits)

pump$getNodeNames()
```

```{r}
pump$x
pump$logProb_x
pump$alpha
pump$theta
pump$lambda
```

```{r}
pump$plot()
pump$plotGraph()
```

```{r}
# Show all dependencies of alpha and beta terminating in stochastic nodes
pump$getDependencies(c("alpha", "beta"))
# Now show only the deterministic dependencies
pump$getDependencies(c("alpha", "beta"), determOnly = TRUE)
# Check that the lifted node was initialized. 
pump[["lifted_d1_over_beta"]] # It was.
# Now let's simulate new theta values
set.seed(1) # This makes the simulations here reproducible
pump$simulate("theta")
pump$theta   # the new theta values
# lambda and logProb_x haven't been re-calculated yet
pump$lambda # these are the same values as above
pump$logProb_x
pump$getLogProb("x") # The sum of logProb_x
pump$calculate(pump$getDependencies(c("theta")))
pump$lambda  # Now they have.
pump$logProb_x
```

### 2.3 Compiling the model

```{r}
Cpump <- compileNimble(pump)
Cpump$theta
```

### 2.4 One-line invocation of MCMC

```{r}
mcmc.out <- nimbleMCMC(code = pumpCode, constants = pumpConsts,
                       data = pumpData, inits = pumpInits,
                       nchains = 2, niter = 1000,
                       summary = TRUE, WAIC = TRUE,
                       monitors = c('alpha','beta','theta')) # 

names(mcmc.out)
```

```{r}
mcmc.out$summary
mcmc.out$WAIC 
```

### 2.5 Creating, compiling and running a basic MCMC configuration

```{r}
pumpConf <- configureMCMC(pump, print = TRUE)
```

```{r}
pumpConf$addMonitors(c("alpha", "beta", "theta"))
```
```{r}
pumpMCMC <- buildMCMC(pumpConf)
CpumpMCMC <- compileNimble(pumpMCMC, project = pump)

niter <- 1000
set.seed(1)
samples <- runMCMC(CpumpMCMC, niter = niter)

par(mfrow = c(2, 2), mai = c(.8, .8, .1, .2))
plot(samples[ , "alpha"], type = "l", xlab = "iteration",
     ylab = expression(alpha),col="cornflowerblue")
plot(samples[ , "beta"], type = "l", xlab = "iteration",
     ylab = expression(beta),col="cornflowerblue")
plot(samples[ , "alpha"], samples[ , "beta"], xlab = expression(alpha),
     ylab = expression(beta),col="cornflowerblue")
plot(samples[ , "theta[1]"], type = "l", xlab = "iteration",
     ylab = expression(theta[1]),col="cornflowerblue")
acf(samples[, "alpha"]) # plot autocorrelation of alpha sample
acf(samples[, "beta"])  # plot autocorrelation of beta  sample
```

### 2.6 Customizing the MCMC

```{r}
pumpConf$addSampler(target = c("alpha", "beta"), type = "RW_block",
                    control = list(adaptInterval = 100))
                                     
pumpMCMC2 <- buildMCMC(pumpConf)

# need to reset the nimbleFunctions in order to add the new MCMC
CpumpNewMCMC <- compileNimble(pumpMCMC2, project  = pump,
                              resetFunctions = TRUE)

set.seed(1)
CpumpNewMCMC$run(niter)
```


```{r}
samplesNew <- as.matrix(CpumpNewMCMC$mvSamples)

par(mfrow = c(2, 2), mai = c(.8, .8, .1, .2))
plot(samplesNew[ , "alpha"], type = "l", xlab = "iteration",
     ylab = expression(alpha),col="cornflowerblue")
plot(samplesNew[ , "beta"], type = "l", xlab = "iteration",
     ylab = expression(beta),col="cornflowerblue")
plot(samplesNew[ , "alpha"], samplesNew[ , "beta"], xlab = expression(alpha),
     ylab = expression(beta),col="cornflowerblue")
plot(samplesNew[ , "theta[1]"], type = "l", xlab = "iteration",
     ylab = expression(theta[1]),col="cornflowerblue")
acf(samplesNew[, "alpha"]) # plot autocorrelation of alpha sample
acf(samplesNew[, "beta"])  # plot autocorrelation of beta  sample
```

### 2.7 Running MCEM

```{r}
pump2 <- pump$newModel()

box = list( list(c("alpha","beta"), c(0, Inf)))

pumpMCEM <- buildMCEM(model = pump2, latentNodes = "theta[1:10]",
                      boxConstraints = box)
pumpMLE <- pumpMCEM$run()
pumpMLE
```

### 2.8 Creating your own functions

```{r}
simNodesMany <- nimbleFunction(
    setup = function(model, nodes) {
        mv <- modelValues(model)
        deps <- model$getDependencies(nodes)
        allNodes <- model$getNodeNames()
    },
    run = function(n = integer()) {
        resize(mv, n)
        for(i in 1:n) {
            model$simulate(nodes)
            model$calculate(deps)
            copy(from = model, nodes = allNodes,
                 to = mv, rowTo = i, logProb = TRUE)
        }
    })

simNodesTheta1to5 <- simNodesMany(pump, "theta[1:5]")
simNodesTheta6to10 <- simNodesMany(pump, "theta[6:10]")
```


```{r}
set.seed(1)  # make the calculation repeatable
pump$alpha <- pumpMLE[1]
pump$beta <- pumpMLE[2]
# make sure to update deterministic dependencies of the altered nodes
pump$calculate(pump$getDependencies(c("alpha","beta"), determOnly = TRUE))

```

```{r}
saveTheta <- pump$theta
simNodesTheta1to5$run(10)
simNodesTheta1to5$mv[["theta"]][1:2]
```

```{r}
simNodesTheta1to5$mv[["logProb_x"]][1:2]
```

```{r}
CsimNodesTheta1to5 <- compileNimble(simNodesTheta1to5,
                                    project  = pump, resetFunctions = TRUE)
Cpump$alpha <- pumpMLE[1]
Cpump$beta <- pumpMLE[2]
Cpump$calculate(Cpump$getDependencies(c("alpha","beta"), determOnly = TRUE))
```

```{r}
Cpump$theta <- saveTheta

set.seed(1)
CsimNodesTheta1to5$run(10)
```

```{r}
CsimNodesTheta1to5$mv[["theta"]][1:2]
CsimNodesTheta1to5$mv[["logProb_x"]][1:2]
```

