---
title: 'STAT561 Fall 2017 HW1'
author: "Shen Qu"
date: "9/29/2017"
output: 
  html_document:
   toc: true
   toc_float: true

---


1.13 If $P(A)=1/3$ and $P(B^c)=1/4$, can A and B be disjoint? Explain.  

> According to properties of the probability function

$$P(B)=1-P(B^c)=1-\frac14=\frac34$$

> According to the Bonferroni's Inequality:

$$P(A_1\cap ...\cap A_n)\ge\sum_{i=1}^\infty P(A_i)-(n-1)$$

$$\implies\quad P(A\cap B)\ge P(A)+P(B)-(2-1)=\frac13+\frac34-1=\frac1{12}$$

> According to properties of the probability function, $A\cap B\ne\emptyset$. Therefore, A and B don't satisfy the definition of pairwise disjoint.

1.19 If a multivariate function has continuous partial derivatives, the order in which the derivatives are calculated does not matter. Thus, for example, the function f(x, y) of two variables has equal third partials  

$$\frac{\partial^3}{\partial x^2\partial y}f(x,y)=\frac{\partial^3}{\partial y\partial x^2}f(x,y)$$

   (a) How many fourth partial derivatives does a function of three variables have? 
    
We can write the function as:
$$\frac{\partial^4}{\partial x^a\partial y^b\partial z^c}f(x,y,z),\quad a+b+c=4$$

```{r}
C <- function(n,k){choose(n,k)} # combination
P <- function(n,k){choose(n,k) * factorial(k)} # permutation
C(6,4)

```

   (b) Prove that a function of n variables has $\binom{n+r-1}{r})$ rth partial derivatives.  

 > A function of n variables has rth partial derivatives. The count of partial deribatives means a combination selected from n variable with r time. 

1.20 My telephone rings 12 times each week, the calls being randomly distributed among the 7 days. What is the probability that I get at least one call each day?  
(Answer: .2285)
(https://jcnts.wordpress.com/2011/05/15/at-least-one-call-each-day/)
We assume each call is unique.

```{r}
# X1…n ∼Binomial(size,p)
# Generate 1000000 occurrences of 12 calls, each with 7/12 probability
# Finding density with simulation
mean(rbinom(1000000,12,7/12)==7)
# Calculating exact probability density
dbinom(7, 12, 7/12)
```

If each call is randomly assigned to a day, then this is equivalent to asking for the probability of finding each of 7 coupons on 12 draws in the Coupon collector's problem. It's also equivalent to asking the probability that a random 12-letter word written with a 7-letter alphabet contains all 7 letters.
There are a total of 7^12 words, and the number containing all 7 letters is, by inclusion/exclusion:

$$7^{12}-\binom766^{12}+\binom755^{12}-\binom744^{12}+\binom733^{12}-\binom722^{12}+\binom711^{12}$$
$$\frac{3162075840}{7^{12}}\approx 22.846\%$$

1.27 Verify the following identities for $n\ge2$.  

(a) $\sum_{k=0}^n(-1)^k(_k^n)=0$  

$$\sum_{k=0}^n(-1)^k(_k^n)=\begin{cases}-\binom n1,-\binom n3,...,-\binom n{n-2},-\binom nn &when k=odd\\ 0,\binom n2,\binom n4,...\binom n{n-2},\binom nn &when k=even\end{cases}=0$$

$$\sum_{k=0}^n(-1)^k(_k^n)=0-\binom n1+\binom n2-\binom n3+\binom n4...+\binom n{n-3}-\binom n{n-2}+\binom n{n-1}+\binom nn =0$$

(b) $\sum_{k=1}^nk(_k^n)=n2^{n-1}$  

(c) $\sum_{k=1}^n(-1)^{k+1}k(_k^n)=0$  


---

1.38 Prove each of the following statements. ( Assume that any conditioning event has positive probability.)  

(a) If P(B) = 1 , then P(A|B) = P(A) for any A.  

(b) If A $\subset$ B, then P(B|A) = 1 and P(A|B) = P(A)/P(B).  

(c) If A and B are mutually exclusive, then  

$$P(A|A\cup B) = \frac{P(A)}{P(A) + P(B)}$$

(d) $P(A)\cap B\cap C) = P(A|B\cap C)P(B|C)P(C)$ .  