---
title: "STAT561 Fall 2018 HW7"
author: 'Shen Qu'
date: "11/22/2018"
output: html_document
---

**3.28** Show that each of the following families is an exponential family.

`(a) normal family` with either parameter $\mu$ or $\sigma$ known

When $\mu$ is known, $f(x|\sigma^2)=\frac{1}{\sqrt{2\pi\sigma^2}}e^{-\frac1{2\sigma^2}(x-\mu)^2}={h(x)}{c(\sigma^2)}e^{w_1(\sigma^2)t_1(x)}$

$h(x)$ | $c(\sigma^2)$ | $w_1(\sigma^2)$ | $t_1(x)$
--- | --- | --- | ---
 $1$ | $\frac{1}{\sqrt{2\pi\sigma^2}}I_{(0,\infty)}(\sigma^2)$ | $-\frac1{2\sigma^2}$ | $(x-\mu)^2$
<span></span> | <span></span> | <span></span> | <span></span>


When $\sigma$ is known, $f(x|\mu)=\frac{1}{\sqrt{2\pi\sigma^2}}e^{-\frac1{2\sigma^2}(x-\mu)^2}=\frac{1}{\sqrt{2\pi\sigma^2}}e^{-\frac{x^2}{2\sigma^2}+\frac{x\mu}{\sigma^2}-\frac{\mu^2}{2\sigma^2}}=e^{-\frac{x^2}{2\sigma^2}}\frac{e^{-\frac{\mu^2}{2\sigma^2}}}{\sqrt{2\pi\sigma^2}}e^{\frac{x}{\sigma^2}\mu}$

$h(x)$ | $c(\mu)$ | $w_1(\mu)$ | $t_1(x)$
------ | -------- | --------- | ------
 $e^{-\frac{x^2}{2\sigma^2}}$ | $\frac{1}{\sqrt{2\pi\sigma^2}}e^{-\frac{\mu^2}{2\sigma^2}}$ | $\mu$ | $\frac{x}{\sigma^2}$

****
`(b) gamma family` with either parameter $\alpha$ or $\beta$ known or both unknown

When $\alpha$ is known, $f(x|\beta)=\frac{x^{\alpha-1}}{\Gamma(\alpha)\beta^{\alpha}}e^{-\frac{x}{\beta}}={h(x)}{c(\beta)}e^{w_1(\beta)t_1(x)}$

$h(x)$ | $c(\beta)$ | $w_1(\beta)$ | $t_1(x)$
------ | -------- | --------- | ------
 $\frac{x^{\alpha-1}}{\Gamma(\alpha)},x>0$ | $\frac{1}{\beta^{\alpha}}$ | $\frac1{\beta}$ | $-x$

When $\beta$ is known, $f(x|\alpha)=e^{-\frac{x}{\beta}}\frac{1}{\Gamma(\alpha)\beta^{\alpha}}e^{(\alpha-1)\ln{x}}={h(x)}{c(\alpha)}e^{w_1(\alpha)t_1(x)}$

$h(x)$ | $c(\alpha)$ | $w_1(\alpha)$ | $t_1(x)$
------ | -------- | --------- | ------
 $e^{-\frac{x}{\beta}}$ | $\frac{1}{\Gamma(\alpha)\beta^{\alpha}}$ | $\alpha-1$ | $\ln x$

When $\alpha$ and $\beta$ are unknown, $f(x|\alpha,\beta)=\frac{1}{\Gamma(\alpha)\beta^{\alpha}}e^{(\alpha-1)\ln{x}-\frac{x}{\beta}}={h(x)}{c(\beta)}e^{w_1(\beta)t_1(x)+w_2(\alpha,\beta)t_2(x)}$

$h(x)$ | $c(\alpha,\beta)$ | $w_1(\alpha)$ | $t_1(x)$ | $w_2(\alpha,\beta)$ | $t_2(x)$
------ | -------- | --------- | ------
 $I_{(0,\infty)}(x)$ | $\frac{1}{\Gamma(\alpha)\beta^{\alpha}}$ | $\alpha-1$ | $\ln x$ | $-\frac1\beta$ | $x$

****
`(c) beta family` with either parameter $\alpha$ or $\beta$ known or both unknown

When $\alpha$ is known, $f(x|\beta)=\frac{x^{\alpha-1}}{B(\alpha,\beta)}(1-x)^{\beta-1}=\frac{x^{\alpha-1}}{B(\alpha,\beta)}e^{(\beta-1)\ln(1-x)}={h(x)}{c(\beta)}e^{w_1(\beta)t_1(x)}$

$h(x)$ | $c(\beta)$ | $w_1(\beta)$ | $t_1(x)$
------ | -------- | --------- | ------
 $x^{\alpha-1}I_{[0,1]}(x)$ | $\frac{1}{B(\alpha,\beta)}$ | $\beta-1$ | $\ln(1-x)$

When $\beta$ is known, $f(x|\alpha)=\frac{(1-x)^{\beta-1}}{\Gamma(\alpha)\beta^{\alpha}}e^{(\alpha-1)\ln{x}}={h(x)}{c(\alpha)}e^{w_1(\alpha)t_1(x)}$

$h(x)$ | $c(\alpha)$ | $w_1(\alpha)$ | $t_1(x)$
------ | -------- | --------- | ------
 $(1-x)^{\beta-1}I_{[0,1]}(x)$ | $\frac{1}{B(\alpha,\beta)}$ | $\alpha-1$ | $\ln x$

When $\alpha$ and $\beta$ are unknown, $f(x|\alpha,\beta)=\frac{1}{B(\alpha,\beta)}x^{\alpha-1}(1-x)^{\beta-1}=\frac{1}{B(\alpha,\beta)}e^{(\alpha-1)\ln x}e^{(\beta-1)\ln(1-x)}={h(x)}{c(\beta)}e^{w_1(\alpha)t_1(x)+w_2(\beta)t_2(x)}$

$h(x)$ | $c(\alpha,\beta)$ | $w_1(\alpha)$ | $t_1(x)$ | $w_2(\beta)$ | $t_2(x)$
------ | ---- | ---- | ---- | ---- | -----
 $I_{[0,1]}(x)$ | $\frac{1}{B(\alpha,\beta)}$ | $\alpha-1$ | $\ln x$ | $\beta-1$ | $\ln(1-x)$

****
`(d) Poisson family`

When $\lambda$ is known, $f(x|\lambda)=\frac{\lambda^x}{x!}e^{-\lambda}=\frac{1}{x!}e^{-\lambda}e^{\ln(\lambda)x}={h(x)}{c(\lambda)}e^{w_1(\lambda)t_1(x)}$

$h(x)$ | $c(\lambda)$ | $w_1(\lambda)$ | $t_1(x)$
------ | -------- | --------- | ------
$\frac{1}{x!}I_{0,1,2..}(x)$ | $e^{-\lambda}$ | $\ln(\lambda)$ | $x$

****
`(e) negative binomial family` with r known, 0<p<1

When $p$ is known, $f(x|p)=\binom{x-1}{r-1}p^r(1-p)^{x-r}=\binom{x-1}{r-1}(\frac{p}{1-p})^r(1-p)^{x}=\binom{x-1}{r-1}(\frac{p}{1-p})^re^{\ln(1-p) x}={h(x)}{c(p)}e^{w_1(p)t_1(x)}$

$h(x)$ | $c(p)$ | $w_1(p)$ | $t_1(x)$
------ | -------- | --------- | ------
$\binom{x-1}{r-1}I_{r,r+1..}(x)$ | $(\frac{p}{1-p})^r$ | $\ln(1-p)$ | $x$


**3.33** For each of the following families:

(i) Verify that it is an exponential family.

(ii) Describe the curve on which the $\theta$ parameter vector lies.

(iii) Sketch a graph of the curved parameter space.

`(a)` $n(\theta, \theta)$

$f(x|\theta,\theta)=\frac{1}{\sqrt{2\pi\theta}}e^{-\frac1{2\theta}(x-\theta)^2}=\frac{1}{\sqrt{2\pi\theta}}e^{-\frac{x^2}{2\theta}+x-\frac{\theta}{2}}=e^x\frac{e^{-\frac{\theta}{2}}}{\sqrt{2\pi\theta}}e^{-\frac{x^2}{2\theta}}={h(x)}{c(\theta)}e^{w_1(\theta)t_1(x)}$

$h(x)$ | $c(\theta)$ | $w_1(\theta)$ | $t_1(x)$
------ | ------------- | -------------- | ------
 $e^xI_{(-\infty,\infty)}(x)$ | $\frac{e^{-\frac{\theta}{2}}}{\sqrt{2\pi\theta}},\theta>0$ | $-\frac1{2\theta}$ | $x^2$

Therefore, this function is an exponential family. The $\theta$ parameter vector lies on a nonnegative real line.

```{r, eval=FALSE, include=FALSE}
curve(dnorm(10, mean = x, sd = x, log = FALSE), -100, 100)

curve(pnorm(10, mean = x, sd = x, lower.tail = TRUE, log.p = FALSE), -100, 100)

curve(qnorm(0.5, mean = x, sd = 2*x, lower.tail = TRUE, log.p = FALSE), -100, 100)

curve(rnorm(x, mean = 0, sd = 1), -1000, 1000)



curve(exp(-((x-theta)^2))/(2*theta)/(sqrt(2*pi*theta)), 0, 10)

```


****
`(b)` $n(\theta, a\theta^2)$, a known

$f(x|\theta,a\theta^2)=\frac{1}{\sqrt{2\pi{a}\theta^2}}e^{-\frac1{2a\theta^2}(x-\theta)^2}=\frac{1}{\sqrt{2\pi{a}\theta^2}}e^{-\frac{x^2}{2a\theta^2}+\frac{x}{a\theta}-\frac{1}{2a}}=\frac{e^{-\frac{1}{2a}}}{\sqrt{2\pi{a}\theta^2}}e^{-\frac{x^2}{2a\theta^2}}e^{\frac{x}{a\theta}}={h(x)}{c(\theta)}e^{w_1(\theta)t_1(x)}$

$h(x)$ | $c(\theta)$ | $w_1(\theta)$ | $t_1(x)$ | $w_2(\theta)$ | $t_2(x)$
------ | ---- | ---- | ---- | ---- | -----
 $I_{(-\infty,\infty)}(x)$ | $\frac{e^{-\frac{1}{2a}}}{\sqrt{2\pi{a}\theta^2}},-\infty<\theta<\infty,a>0$ | $-\frac1{2a\theta^2}$ | $x^2$ | $\frac1{a\theta}$ | $x$

Therefore, this function is an exponential family. The $\theta$ parameter vector lies on a parabola.

****
`(c)` gamma($\alpha, 1/\alpha$)

$f(x|\alpha,\frac1{\alpha})=\frac{\alpha^{\alpha}}{\Gamma(\alpha)}e^{(\alpha-1)\ln{x}-\alpha{x}}={h(x)}{c(\alpha)}e^{w_1(\alpha)t_1(x)+w_2(\alpha)t_2(x)}$

$h(x)$ | $c(\alpha,\beta)$ | $w_1(\alpha)$ | $t_1(x)$ | $w_2(\alpha,\beta)$ | $t_2(x)$
------ | ---- | ---- | ---- | ---- | -----
 $I_{(0,\infty)}(x)$ | $\frac{\alpha^{\alpha}}{\Gamma(\alpha)}$ | $\alpha-1$ | $\ln x$ | $-\alpha$ | $x$

Therefore, this function is an exponential family. The $\alpha$ parameter vector lies on a line.

****
`(d)` $f(x|\theta) = C exp (-(x-\theta)^4)$ , C a normalizing constant

$f(x|\theta)=Ce^{-(x-\theta)^4}=Ce^{-x^4}e^{-\theta^4}e^{4x^3\theta}e^{-6x^2\theta^2}e^{4x\theta^3}={h(x)}{c(\theta)}e^{w_1(\theta)t_1(x)+w_2(\theta)t_2(x)+w_3(\theta)t_3(x)}$

$h(x)$ | $c(\theta)$ | $w_1(\theta)$ | $t_1(x)$ | $w_2(\theta)$ | $t_2(x)$ | $w_3(\theta)$ | $t_3(x)$
------ | ---- | ---- | ---- | ---- | ----- | ---- | -----
 $Ce^{-x^4}I_{(-\infty,\infty)}(x)$ | $e^{-\theta^4},-\infty<\theta<\infty$ | $4\theta$ | $x^3$ | $-6\theta^2$ | $x^2$ | $4\theta^3$ | $x$

Therefore, this function is an exponential family. The $\theta$ parameter vector lies on a 3D line.

```{r, eval=FALSE, include=FALSE}
library(plot_ly)
library(reshape2)

#pp <- function (n,r=4) {
    x <- seq(-r*pi, r*pi, len=n)
    df <- expand.grid(x=x, y=x)
    df$r <- sqrt(df$x^2 + df$y^2)
    df$z <- cos(df$r^2)*exp(-df$r/6)
    df
#}
#data_xyz <- pp(100)
#data_z <- acast(data_xyz, x~y, value.var = "z")
#plot_ly(z = data_z,  type = "surface")

#x <- seq(-pi, pi, len = 20)
#y <- seq(-pi, pi, len = 20)
#g <- expand.grid(x = x, y = y)
#g$z <- sin(sqrt(g$x^2 + g$y^2))
#wireframe(z ~ x * y, g, drape = TRUE,
#aspect = c(3,1), colorkey = TRUE)

```

```{r, eval=FALSE, include=FALSE}
library(plotly)

count <- 3000

x <- c()
y <- c()
z <- c()
c <- c()

for (i in 1:count) {
  r <- i * (count - i)
  x <- c(x, r * cos(i / 30))
  y <- c(y, r * sin(i / 30))
  z <- c(z, i)
  c <- c(c, i)
}

data <- data.frame(x, y, z, c)

p <- plot_ly(data, x = ~x, y = ~y, z = ~z, type = 'scatter3d', mode = 'lines',
        line = list(width = 4, color = ~c, colorscale = list(c(0,'#BA52ED'), c(1,'#FCB040'))))
```

****

**3.38** Let Z be a random variable with pdf f(z). Define $z_{\alpha}$ to be a number that satisfies this relationship:
$$\alpha= P ( Z > z_{\alpha}) =\int_{z_{\alpha}}^{\infty}f(z)dz.$$
Show that if X is a random variable with pdf ($l/\sigma)f((x-\mu)/\sigma$) and $x_{\alpha}=\sigma z_{\alpha}+\mu$, then $P(X>x_{\alpha})=\alpha$. (Thus if a table of $z_{\alpha}$ values were available, then values of $x_{\alpha}$ could be easily computed for any member of the location-scale family.)

According to Theorem 3.5.6, f(x) is pdf, $\mu,\sigma\in\mathbf{R}$, $\sigma>0$. X is a random variable with pdf ($1/\sigma)f((x-\mu)/\sigma$) when and only when Z is a random variable with pdf f(z) and $X=\sigma Z+\mu$

$$\therefore P(X>x_{\alpha})=P(\sigma Z+\mu>\sigma z_{\alpha}+\mu)= P(Z>z_{\alpha})=\alpha $$