---
title: 'linear Regression'
author: "Shen Qu"
date: "10/17/2018"
output: 
  html_document:
    toc: true
---

```{r setup, include=F}
knitr::opts_chunk$set(message=FALSE, warning=F, echo=TRUE)
options(width = 2000)
options(repos="https://cran.rstudio.com")
```

# {.tabset .tabset-fade .tabset-pills}

## Regression Analysis

http://rich-iannone.github.io/DiagrammeR/index.html

```{r, echo=TRUE}

library(DiagrammeR)

grViz("
digraph nicegraph {

  # graph, node, and edge definitions
  graph [compound = true, nodesep = .5, ranksep = .25,
         color = crimson]

  node [fontname = Helvetica, fontcolor = darkslategray,
        shape = rectangle, fixedsize = true, width = 1,
        color = darkslategray]

  edge [color = grey, arrowhead = none, arrowtail = none]

  # subgraph for R information
  subgraph cluster0 {
    node [fixedsize = true, width = 3]
    '@@1-1' -> '@@1-2' -> '@@1-3' -> '@@1-4'
    '@@1-4' -> '@@1-5' -> '@@1-6' -> '@@1-7'
  }

  # subgraph for RStudio information
  subgraph cluster1 {
    node [fixedsize = true, width = 3]
    '@@2' -> '@@3'
  }

  Information             [width = 1.5]
  Information -> R
  Information -> RStudio
  R -> '@@1-1'            [lhead = cluster0]
  RStudio -> '@@2'        [lhead = cluster1]

}

[1]: 1:7
[2]: 38
[3]: 'Current \\n xxx'

")

```

``````{r, echo=TRUE}

DiagrammeR::mermaid("
graph LR
A(Rounded)-->B[Rectangular]
B-->C{A Rhombus}
C-->D[Rectangle One]
C-->E[Rectangle Two]
")
```


```{r, echo=TRUE}
###
# Create two graphs and combine them into one
###

library(DiagrammeR)

# Create the first graph
nodes_1 <-
  create_node_df(10, nodes = 1:10,
    label = FALSE,
    type = "lower",
    style = "filled",
    color = "aqua",
    shape = "square",
    data = c(3.5, 2.6, 9.4, 2.7))

edges_1 <-
  create_edge_df(
    from = 1:9,
    to = 2:10,
    rel = "requires",
    color = "green",
    data = c(2.7, 8.9, 2.6, 0.6))

graph_1 <-
  create_graph(
    nodes_df = nodes_1,
    edges_df = edges_1,
    directed = TRUE)

# Create the second graph (note that node ID values
# are different from those of the first graph)
nodes_2 <-
  create_node_df(10, nodes = 1:10,
    label = TRUE,
    type = "upper",
    style = "filled",
    color = "red",
    shape = "rectangle",
    data = c("0.5", "3.9", "3.7", "8.2"))

edges_2 <-
  create_edge_df(
    from = 1:9,
    to = 2:10,
    rel = "receives",
    arrowhead = "vee",
    color = "red")

graph_2 <-
  create_graph(
    nodes_df = nodes_2,
    edges_df = edges_2,
    directed = F)

# Combine the two graphs, the
# global graph attribute
# `graph_attrs = "rankdir = LR"`
# will be retained since it is
# part of the graph supplied as `x`
combined_graph <-
  combine_graphs(x = graph_1, y = graph_2)

# Display the combined graph
render_graph(combined_graph)


```
## Multiple Regression Models