---
title: 'USP537 Problem Sets'
author: "Shen Qu"
date: "Winter 2019"
output: html_document
---

#  {.tabset .tabset-fade .tabset-pills}

## Problem Set 1

 1. Portland’s transit agency TriMet wants to increase transit usage by 15%. They expect no changes in the demand for transit and will lower the fare to attract the additional riders. The estimated price elasticity of demand is 0.3(actual elasticity used by TriMet).

 a. Calculate the percentage change necessary in fares to achieve the increased transit usage and the change in fare box revenue (use simple percentages rather than the midpoint elasticities). 

When no changes in the demand for transit, the value of usage moves along the MD curve, then 

$$\text{Elasticity }\eta=\frac{\Delta Q/Q}{\Delta P/P}\implies\Delta P/P=\frac{\Delta Q/Q}{\eta}=\frac{0.15}{0.3}=50\%$$
$$TR=50\%P(1+15\%)Q=57.5\%PQ$$

50% decrease in fares could achive the increase transit usage by 15%. The new revenue will be 57.5% of previous revenue.

 ---
 
 b. Would you expect the long-run price elasticity of demand to be larger or smaller? Explain why?

I expect the long-run price elasticity of demand to be larger. Demand tends to be more elastic in the long rung rather than in the short run, because when prices change consumers often need more time to respond and change their shopping habits such as househould or firms' location, vehicle ownership, further change travel behaviors. (Mankiw, 2004).

 * Reference

_Mankiw (2004) “Principals of Macroeconomics”,South-WesternCollegePub,UK_

 ---
 
 c. Do you expect that costs will rise? Consider total cost, fixed cost, and variable cost. Explain why it would or would not.

If the extra riderships don't exceed current transit capacity, fixed cost would not change. The variable cost will rise for the addtional operation cost such as energy consumption and labor cost. Thus, the total cost would increase. If current transit is underload, the change would be tiny.

 ---
 
 2. In “The Urban Transportation Problem: A Reexamination and Update”, by John F. Kain([GTW] Chap11), the author concludes the paper with a recommendation: “There remains a strong, if not compelling, case to begin charging road users the full social cost of their road use, for reversing the public ownership and operation of transit services, and for stopping wasteful expenditures on ineffective rail systems.” Please write a short analysis (approximately1 page, single-spaced, typed) highlighting your arguments either in support of or against this recommendation.

There are three points in Kain’s statement: charging road users the full social cost, reversing public ownership and operation of transit services, and stopping ineffective rail systems (Kain, 2001 p.396). The three concepts are related. In general, I agree with him and try to make a further discussion.

 - Charging road users the full social cost
 
The principle of road pricing is to balance the cost and externality. Because the road using costs do not reflect true costs to society and environment, users are not paying the full costs of their purchases and even unaware of it. They will tend to use too much that is subsidized by society. The markets are distorted and inefficient.

Road user charges have two functions. The first one is used for revenue generation, which was used to road infrastructure financing. The second function is as a transportation demand management tool, which can reduce peak hour travel, release traffic congestion, and encourage transit development. Kain (2001) believes the later: “In the short run, motorists would make fewer trips, change their time of travel, and alter their destinations…In the long run, higher road user charges would eventually affect the dispersed patterns of workplaces and residences that were in part induced by congestion” (P.393).

The problem is how to get “the right price”? Prices should reflect all the costs and externalities including both negative and positive. Many economists try to measure the lost such as congestion, air pollution, GHG emissions, visual intrusion, noise, and road accidents.  But “economists disagree over how to set tolls, how to cover common costs, what to do with any excess revenues, whether and how "losers" from tolling previously free roads should be compensated, and whether to privatize highways.  It is hard to measure the benefit of road using such as economic growth, the communication of information and technology, the recreation activities, and the emotional need for social lives. Kain’s main idea is charging low-occupancy vehicles and subsidizing transit (p.393). It is also a huge challenge at the operating level, which leads to the second topic: privatization.

The application of congestion charges is currently limited to a small number of cities and urban roads.  The technology of electronic road pricing through ITS is making detailed billing based on actual use feasible and affordable.

 - Reversing public ownership and operation of transit services
 
Transit is a quasi-public good, Off-peak travelers’ consumption does not reduce any other's consumption of that good.  In some time or location, transit is scarce goods. Both excludability and rivalry are possible. Hart (1995) proposes the party with the better investment technology should be the owner , while Besley and Ghatak (2001) argue that the party who has a larger valuation for the public good should be the owner.  Klein, Moore and Reja (1997) point out public transit authorities have tended to over-provide their service. (P.12). It seems that privatization can help to improve efficiency and achieve benefit maximum inside the transportation system, promote the resources redistribution by the hand of market. But Market cannot solute the problem of externalities by itself. The free travel and dispersed patterns might become the wealthy’s privilege. For transit, the scarce time and segments will turn to “club goods,” which means high excludability and low rivalry in consumption. In social justice or equality, the public department still needs to provide alternative goods or subsidies. Especially, mass transit, what is commonly known as natural monopolies, need huge initial investment and have essentially zero marginal costs. It beyond the issues of ownership or operation.

 - The ineffective rail systems
 
Kain (2001) believes busway can close match between demand and capacity with the lower operating costs, (P.379). “Research on mode choice shows that travelers value the following characteristics: short trip times, avoidance of transfers and waiting time, door-to-door service, reliability, comfort, seat availability, storage space, security, and flexibility” (Klein, Moore and Reja, 1997. P.9). In this way, rail system is the most expensive and least flexible, should be stopped.

Other reasons Kain gave are disappointing ridership than projected (p.362), the dispersed employment patterns and low population densities (P.366). It is true that the low density in the U.S. make transit operators difficult to provide competitive services. Meaning while, in some cities with mid-high income and high density such as New York and Hong Kong, a powerful rail system support the cities operation and play an irreplaceable role (Pushkarev and Zupan,1997; Cervero,1998). 

Review the history, “the steady rise in auto ownership and declines in transit use since 1895 were largely efficient responses to changes in technology and rising incomes” (Kain, 2001. P.359). “The transit market is contracting because of natural forces” (Klein, Moore and Reja, 1997. P.16). Therefore, we should be careful when planning a rail in a med-low density area and have to pay the heavy cost of rail system in some cities. Beside advocating BRT and TOD in middle cities or growing area, next generation of technology such as automated driving system could turn the trend in future and lead to an unpredictable land use pattern.

 * Reference

  _[1] Kain, John F. (2001). Essays in transportation economics and policy : a handbook in honor of john r. meyer, Chapter 11. Brookings Institution Press._

  _[2] Paul Johnson, Andrew Leicester & George Stoye. (2012). Fuel for Thought – The what, why and how of motoring taxation. Institute for Fiscal Studies and Royal Automobile Club Foundation for Motoring. Executive Summary, pp. v._
  
  _[3] Lindsey, Robin. (2006). Do Economists Reach a Conclusion on Road Pricing? The Intellectual History of an Idea. Econ Journal Watch. 3 (2): 292–379._
  
  _[4] Small, Kenneth A.; Verhoef, Erik T. (2007). The Economics of Urban Transportation. Routledge, New York. pp. 148–153._
  
  _[5] Demsetz, Harold. (1970). Full Access The Private Production of Public Goods. Journal of Law and Economics. 13 (2): 293–306._
  
  _[6] Hart, Oliver. (1995). Firms, Contracts, and Financial Structure. Oxford University Press._
  
  _[7] Besley, Timothy; Ghatak, Maitreesh. (2001). Government Versus Private Ownership of Public Goods. The Quarterly Journal of Economics. 116 (4): 1343–72._

  _[8] Klein, D. B., Moore, A. T., & Reja, B. (1997). Curb rights: a foundation for free enterprise in urban transit. Brookings Institution Press._


 ---
 
 3. Rail City operates a train service on weekdays only. The fixed cost is $120,000 per day. The operating cost is $100 per train-hour, with service adjusted to provide 200 rides per train-hour. This yields a marginal cost of $0.50 per rider. Use this where information on marginal cost is needed. The following information is available regarding cost and demand on a daily basis.

|        |Cost|Revenue|Demand
|---     | ---|--- | ---
|Fixed   |$FC=\$120000$| |
|Marginal|$MC=0.5$|$MR=10-0.0002Q$|$P=\$10-0.0001Q$|
|Total   |$TC=FC+VC=\$120000+0.5Q$|$TR=P\times Q=10Q-0.0001Q^2$|

 a. What number of riders will use the system if the price is set equal to marginal cost(This is essentially equivalent to assuming that a benevolent social planner sets the price so that maximum efficiency is reached)? How much service will be provided (Q*)? What are the subsidies or profits at this level of cost and price? What is the consumer surplus?

Let $P^*=MC$, then $10-0.0001Q^*=0.5\implies Q^*=95000$ riders. The service is $95000/200=475$ train-hour.

$TR-TC=(P^*Q^*)–(120000+P^*Q^*)=-\$120000$

$CS^*= \frac12(10-P^*)Q^*=\frac12\times(10-0.5)\times95000=\$451250$

When the price equals marginal cost, there will be 95000 riders and 475 train-hour service will be provided.It needs $120000 subsidies for operation. The consumer surplus reach the maximum \$451250.

 ---
 
 b. If the price is raised to $2.00, how many people will use the train, and what happens to total revenue? What subsidy is required? Calculate any efficiency loss or gain associated with this change. Compare the change in subsidy with the change in consumer surplus.

When $P_2=\$2.00$, there will be $Q_2=\frac{10-2}{0.0001}=80000$ riders.

$TR=P_2Q_2=2\times80000=\$160000$.

$TR-TC=160000–(120000+0.5\times8000)=\$0$, $\Delta(TR-TC)=0-(-120000)=\$120000$

$CS=\frac12(10-P_2)Q_2=\frac12(10-2)80000=320000$, $\Delta{CS}=320000-451250=-\$131250$

$PS=(P_2-P^*)Q_2=(2-0.5)80000=120000$

$TS=CS+PS=320000+120000=440000$

$DWL=\frac12\Delta P\Delta Q=\frac12(2.0-0.5)(80000-95000)=-\$11250$

When the price equals to \$2, there are 80000 riders. The total revenue is \$160000 and don't need subsidies. But the consumer surplus loss \$131250, which is greater than the decrease in subsidies \$120000. It means we should not raise price for reducing subsidies.

 ---
 
 c. Suppose that a monopolist is allowed to run the system and set price. What quantity ($Q_m$) and price ($P_m$) will be set and what profits or losses will be generated if the monopolist wants to maximize profits? Calculate any efficiency loss associated with this level of service.

When $MR=MC$, the monopolist maximizes profits. $10–0.0002Q_m=0.5\implies Q_m=47500$ riders,  $P_m=10–0.0001Q=10–4.75=\$5.25$

$TR-TC=(P_m-P^*)Q_m-FC=(5.25-0.5)\times47500–120000=\$105625$

$CS=\frac12(10-P_m)Q_m=\frac12(10-5.25)47500=112812.5$

$PS=(P_m-P^*)Q_m=(5.25-0.5)47500=225625$

$TS=CS+PS=112812.5+225625=338437.5$

$DWL=\frac12\Delta P\Delta Q=\frac12(5.25-0.5)(47500-95000)=-\$112812.5$

When the marginal revenue equals marginal cost, the price is \$5.25 and riders are 47500. The monopolist get the maximum profit of \$105625. However, the total efficientcy lost are \$112812.5. We should not sacrific the publis interest for maximizing the monopolist's profit.

 ---
 
 d. Could the monopolist make a profit if fixed costs are $250,000 per day?  Does this mean that it would be a mistake to build the train based on economic benefits and costs for this level of fixed cost? Explain why you think it would or would not be from an economic efficiency perspective and include appropriate calculations.

Let $FC=\$250000$, the monopolist's maximum profit is $TR-VC-FC=(P_m-P^*)Q_m-250000=-\$24375$.

The monopolist won't make a profit. To examine the efficiency, we should compare subsidies and cunsumer surplus. 

If the price is set to $P^*=0.5$, the social benefit is $CS+TR-TC=451250+(0.5-0.5)95000-250000=\$201250$

If the price is set to $P_m=5.25$, the social benefit is $CS+TR-TC=112812.5-24375=\$88437.5$

Let $TS+TR-TC=\frac12(10+P-2P^*)Q-250000=-0.00005Q^2+9.5Q–250000=0$, $Q_{1,2}=\frac{-b\pm \sqrt{b^2-4ac}}{2a}=\frac{9.5\pm\sqrt{40.25}}{0.0001}$, $Q_1=31558,Q_2=158442$. For $Q\le95000$, so $Q=31557.11,P=6.844289$.

Then $CS+TR-TC=49792.56-49792.57=\$0$. Once the price is lower than \$6.844289, the riders are more than 31558, the cunsumer surplus will be larger than subsidies.

The reults prove that the train is worth to be bulit because social benefit are positive rather than considering the profit of monopolist. Only if the riders are less than 31558, building this train is a mistake for a negative social benefit.


|P(\$)|Rider|Fixed Cost|Variable Cost|Total Cost||Revenue|subsidies/profit||Consumer Surplus|Producer Surplus|Total Surplus        | DWL                     |
|---  | --- | ---      |---          | ---      |---| ---| ---       | --- |---             | ---            | ---                 |---                      |
|$P$  |$Q$  |$FC$      |$P^*Q$       |$FC+VC$   ||$PQ$   |$(P-P^*)Q-FC$   ||$\frac12(10-P)Q$|$(P-P^*)Q$      |$\frac12(10+P-2P^*)Q$|$\frac12\Delta P\Delta Q$|
|0.50 |95000|120000    |47500        |167500    ||47500  | -120000        ||451250          | 0              |451250               |0                        |
|2.00 |80000|120000    |40000        |160000    ||160000 | 0              ||320000          | 120000         |440000               |-11250                   |
|5.25 |47500|120000    |23750        |143750    ||249375 | 105625         ||112812.5        | 225625         |338437.5             |-112812.5                |
|0.50 |95000|250000    |47500        |297500    ||47500  | -250000        ||451250          | 0              |451250               |0                        |
|5.25 |47500|250000    |23750        |273750    ||249375 | -24375         ||112812.5        | 225625         |338437.5             |-112812.5                |
|6.84 |31558|250000    |15779        |265779    ||215986 | -49792.57      ||49792.56        | 200207.4       |250000               |-201250                  |

```{r echo=FALSE}
curve(10-0.0001*x, 0, 100000,col = "red",xlab ="Q", ylab ="P");text(40000,7, "MD")
curve(10-0.0002*x, add = TRUE, col = "violet");text(40000,3, "MR")
abline(h = 0.5, lty = 2);abline(h = 2, lty = 4)
abline(v =95000, lty=2);text(95000,0.8, "(95000,0.5)")
abline(v =47500, lty=2);text(50000,5.5, "(47500,5.25)");text(47500,0.8, "(47500,0.5)")
abline(v =80000, lty=4);text(80000,2.5, "(80000,2)")
polygon(x=c(47500,95000,47500), y=c(0.5,0.5,5.25), density=10, angle=45,lty =3);text(60000,2.5, "DWL")
abline(v =31558, lty=4);text(35000,8, "(31558,6.8442)")
polygon(x=c(0,31558,0), y=c(6.8442,6.8442,10), density=6, angle=90,lty =3);text(15000,8, "CS")
polygon(x=c(0,31558,31558,0), y=c(0.5,0.5,6.8442,6.8442), density=6, angle=0,lty =3);text(15000,4, "PS")
```

```{r eval=FALSE, include=FALSE}
uniroot(function(x)(-0.00005*x^2+9.5*x-250000), lower=0, upper=100000)
```



 
 
## Problem Set 2

 1. Elasticities and Consumer Surplus 

In April of 1973, the Massachusetts Bay Transportation Authority (MBTA) initiated an experiment called “Dime Time.”(See attached article) On weekdays between the hours of 10am and 3pm, the fare for travel on the MBTA’s rail lines was reduced from 25 cents to 10 cents, but was 25 cents during all other hours. During the “Dime Time” experiment, which lasted until August1975, fares for bus travel remained at 20 cents per ride during all hours. The “Dime Time” experiment provided an unusual opportunity to evaluate the sensitivity of midday rail transit demand to a reduction in fare, a subject of interest to many transit operators and planners. 

Table 1: Average weekday MBTA rail transit passengers between 10am and 3pm with “Dime Time”, by previous mode of travel and income class

|		| income class	
|Previous travel mode| Low	| Moderate | high | total
|----|---|---|---|---
| MBTA rail transit	|  
| 7am - 10 am	   |370	 | 215  |175  |	760
|10 am - 3 pm    | 9500| 8250	|5100	|22850
|3 pm - 6 pm	   | 255 | 165	|135	|555
|MBTA bus transit|1080 | 680	|340	|2100
|Private automobile	|	 		
|driver	         |895	 | 525	|400	|1820
|passenger       |220  | 75	  |80	  |375
|walk            |850  | 380	|90	  |1320
|no previous trip|1850 | 570	|90	  |2510
|Toatal          |15020| 10860|6410	|32290

Note on interpretation of this table: This table shows the average number of people who traveled via rail between 10am and 3pm during the “Dime Time” experiment. The different categories tell us what mode these people used before switching to rail.
So, for example, we know that the total number of people riding the rail system during off-peak hours under “Dime Time” was 32,290. Of these 32,290, 22,850 passengers rode during these hours before the lower fare was introduced. 2,100 of the 32,290 previously rode the bus, but switched over to rapid transit with the new lower fare.


  a. Using the data provided in Table 1, estimate the fare elasticity of total weekday demand for MBTA rail transit service during midday (10am¬3pm).

$$\text{Elasticity }\eta=\frac{\Delta Q/Q}{\Delta P/P}=\frac{\frac{32290-22850}{22850}}{\frac{0.1-0.25}{0.25}}=-0.6885485$$


```{r}
((32290-22850)/22850)/((0.1-0.25)/0.25)
((15020-9500)/9500)/((0.1-0.25)/0.25)
((10860-8250)/8250)/((0.1-0.25)/0.25)
((6410-5100)/5100)/((0.1-0.25)/0.25)

```

The fare elasticity of total weekday demand for MBTA rail transit service during midday (10am¬3pm) is -0.6885485. 

 ---
 
  b. According to this elasticity estimate, is the demand for midday rail transit service elastic or inelastic? Comment on your answer.

For $-0.6885485\in(-1,1)$, the demand for midday rail transit service is inelastic. The result shows that the ridershipeach will decrease 0.6885% when fare has 1% increase, which indicates that the consumers are not sensitive with fare change.

  c. How does the elasticity of demand for midday rail transit service vary among the three income groups (low, moderate, and high) for whichridership data are reported inTable 1? Comment onthese variations.

$$\eta_{Low}=\frac{\Delta Q/Q}{\Delta P/P}=\frac{\frac{15020-9500}{9500}}{\frac{0.1-0.25}{0.25}}=-0.9684211$$

$$\eta_{Mid}=\frac{\Delta Q/Q}{\Delta P/P}=\frac{\frac{10860-8250}{8250}}{\frac{0.1-0.25}{0.25}}=-0.5272727$$

$$\eta_{Mid}=\frac{\Delta Q/Q}{\Delta P/P}=\frac{\frac{6410-5100}{5100}}{\frac{0.1-0.25}{0.25}}=-0.4281046$$


Results indicates that the fare elasticity of demand for midday rail transit service varies among income groups. Low income group has highest fare elasticity. The fare elasticity of moderate and high income groups are similar, --0.5272 and -0.4281 respectively.  If there is 1% decrease in fare, there will be 0.9684 ridership increase for low income group, 0.5272 ridership increase for moderate income group and 0.4281 increase for high income group.  Low income group is most sensitive to the fare change. 

  d. Table 1 also reports the prior behavior of MBTA rail transit passengers traveling during “Dime Time” hours. Boston’s regional transportation planning agency estimated that 233,760 auto trips (including both driver and passenger trips) were made in the MBTA service area on a typical weekday before the “Dime Time” program began. From these data, estimate the cross elasticity of demand for auto travel with respect to fare for midday rail transit trips. How do you interpret the arithmetic sign of this elasticity?

Cross-Elasticity = %∆Demand/ % ∆Fare = ((-(1820+375))/233760)/((0.10-0.25)/0.25) = 0.0157
The cross-elasticity of demand for auto travel with respect to fare for midday rail transit trips is 0.0157. The positive arithmetic sigh indicates that auto trips and rail transit trips are substitutes. If there is 1% decrease in transit fare, there will be 0.157% decrease in auto trips. 


  e. The MBTA estimated that prior to the “Dime Time” experiment, average weekday travel on the rail service during peak periods (7¬10am and 3¬6pm) was 203,670 passengers. Estimate the cross¬-elasticity of demand for peak-¬period rail travel with respect to the off¬-peak rail fare.

Cross-Elasticity = %∆Demand/ % ∆Fare = ((-(760+555))/203670)/((0.10-0.25)/0.25) = 0.0107
The cross-elasticity of demand for peak-period rail travel with respect to the off-peak rail fare is 0.0107. The positive arithmetic sigh indicates that peak-period rail travel and rail off-peak rail transit trips are substitutes. If there is 1% decrease in transit fare, there will be 0.0107% decrease in auto trips. 

  f. Please draw a demand curve andanswer the following questions using the diagram:

   (i) Label the area of total daily benefits to riders, as measured by the increase in consumer surplus, of the “Dime Time” experiment?
  
  In above figure, point B represents the price (0.25) and ridership (22850) before Dime Time. Point A represents price (0.10) and ridership (32290). The area marked with oblique lines indicates the increase in consumer surplus because of Dime Time.   
  
   (ii) Label how much of the increase in consumer surplus represented fare savings to passengers who already traveled on the rail system during the midday at the initial fare?
  
  In above figure, point B represents the price (0.25) and ridership (22850) before Dime Time. Point A represents price (0.10) and ridership (32290). The area marked with oblique lines indicates increase in consumer surplus represented fare savings to passengers who already traveled on the rail system during the midday at the initial fare.   

  (iii) Describe verbally how were the benefits in(ii) divided among riders in the three income groups?

The increase in consumer surplus to passengers who already traveled on the rail transit can be divided among riders in the three income groups. The number of low income people is largest, 9500, 42%. The next is moderate income people, 8250, 36%. The number of high income group is 5100, which accounts for smallest share 22%. 

   (iv) Label on the diagram how much of the total benefit was received by new riders (who were initially not using rail) who began to use the rail system when the midday fare was reduced to 10 cents?

In above figure, point B represents the price (0.25) and ridership (22850) before Dime Time. Point A represents price (0.10) and ridership (32290). The area marked with oblique lines indicates increase in consumer surplus represented fare savings to new riders (who were initially not using rail) who began to use the rail system when the midday fare was reduced to 10 cents  

   (v) Describe how were the benefits in(iv) divided among riders inthe three income groups? Why is it important to distinguishbetweenconsumer benefits to passengers who had beenusing the rail system at the higher fare and those received by new riders who beganto use the system inresponse to the fare change?

Among the riders, there are 5220 (58%) low income people, 2610 (28%) moderate income people and 1310 (14%) high income people. The percentage of low income people is largest. 

It is very important to distinguish between consumer benefits for existing passengers and new riders. Existing passengers were already willing to ride midday transit for the old fare, so their increase in consumer surplus from the program results in money saved but no behavioral change. New riders attracted by the new lower fare received some consumer surplus resulting from their travel mode change. The additional consumer surplus for the existing rail riders essentially translate into benefits for the consumer without any benefits for the transit agency. Besides, how policy attracts new riders can provide useful implication for policy-makers. 

 2. Using an aggregate demand model

Lee’s 1980 study of gasoline demand in California yielded the following estimation results: (dependent variable is 1000s gallons of gasoline $R^2=0.92$). 

|Explanatory Variable|Constant|Real Gasoline Price(\$)|Real Income (billion \$)|Population (millinons)|DEC73|JAN74|FEB74|MAR74|APR74|
| ---                | ---:   | ---:                  |  ---:                  |   ---:               |---: |---: |---: |---: |---: |
|Coefficient Estimate|-25193.1|-18552.8               |277.3                   |1567.9  |-1801.5  |1629.7  |-2313.1  |2524.1  |162.7|


  a. Two significant by-products of automobile use are air pollution and traffic congestion. In order to reduce the extent of each, suppose that the federal government raises the gasoline tax by $0.25. According to the above estimation results, what effect will this have on gasoline consumption in California? 

∆GasDemand = β1* ∆RealGasolinePrice = -18552.8*0.25 = - 4638.2

Based on the aggregate demand model, if gasoline price increases $0.25, there gasoline demand will decrease 4638200 gallons of gas in California. 

  b. Use budget lines and indifference curves to illustrate the effect of the tax increase on gasoline consumption and consumer welfare (utility). In addition to the total effect, please label the substitution and income effects.
 
 The above figure illustrates the effect of the tax increase on gasoline consumption and consumer welfare (utility). Gasoline consumption is the X-axis. 
 
Due to the tax increase, the budget line has pivoted from B1 to B2 and the consumption point has moved from A to C. The decrease in the quantity demanded can be divided into two effects. 
In the above figure, from point A to point B is substitution effect. The substitution effect is when the consumer switches consumption patterns due to increase of gas price alone but remains on the same indifference curve. To identify the substitution effect a new budget line needs to be constructed. The budget line B1* is added, this budget line is parallel with the budget line B2 and tangential to I1. Therefore, the movement from Q1 to Q2 is purely due to the substitution effect. 

In the above figure, from point B to point C is income effect. The income effect highlights how consumption will change due to the consumer having a change in purchasing power as a result of the price change. The higher price means the budget line is B2, hence the optimum consumption point is Q2. This point is on a lower indifference curve (I2). 

  c. Wanting to increase the price of gasoline but not reduce economic welfare, suppose that, in addition to the \$0.25 cent tax on gasoline, the government also provides consumers with an income tax rebate amounting to \$16.7 billion dollars (total amount).

   (i) According to the estimation results above, what would be the effect of the gasoline tax and income tax policies on gasoline demand?


∆GasDemand = β1* ∆RealGasolinePrice + β2* ∆RealIncome = -18552.8*0.25 + 227.3*16.7 
= -4638.2 + 4630.91 = -7.29

The effect of the gasoline tax and income tax policies will decrease gasoline demand 7290 gallons. 

   (ii) According to economic theory, would the gasoline tax rise combined with the welfare-offsetting income tax rebate produce any reduction in the demand for gasoline?

A gas tax increase would result in both substitution effect (reduction in gasoline demand because of higher gas price) and income effect (reduction in gasoline demand because of lower purchase power). The income tax rebate would simply result in an income effect (increase in gas demand because of higher income). The welfare-offsetting income tax rebate should offset the income effects of the gas tax, and should not offset the substitution effect (reduction in gasoline demand because of higher gas price).

  (iii) Given your analysis in parts (i) and (ii), is it possible to implement a policy that is consistent with the goals of environmentalists and yet not reduce the economic welfare of consumers?

Yes. It is possible to implement a policy that is consistent with the goals of environmentalists (gasoline tax decrease gasoline demand) and yet not reduce the economic welfare of consumers (tax rebate keeps the same consumer welfare). 

 * Assuming that the dummy variables for the gasoline crises (DEC73 to MAR74) represent significant queuing during those months, we can estimate the estimated gasoline crisis queuing cost premiumper gallon for DEC73 as follows:
 
 ∆QueuingCostJan74/∆Crisis=  τ_2/β_2 = -1629.7/(-18522.8)=0.0878

∆QueuingCostFeb74/∆Crisis=  τ_3/β_2 = -(-2313.1)/(-18552.8)=0.1247

∆QueuingCostMar74/∆Crisis=  τ_4/β_2 = -(-2524.1629.7)/(-18522.8)=0.1360

The queuing cost premium for each gallon of gasoline for JAN74, FEB74, MAR74 is 0.0878, 0.1247, 0.1360 respectively. 

  d. Calculate the income elasticity of demand, given that the sample mean for Income was $0.0011 billion dollars and gas demand was 0.35 thousand gallons.

Elasticitylow = %∆Demand/ % ∆Price = ∆G/∆P*  P/G= β_2*  P/G=177.3*0.0011/0.35=0.8715 
The income elasticity of demand for gasoline was 0.872. If there is a 1% increase in income, there will be a 0.87% increase in gasoline demand. Because 0.08715 < 1, the elasticity of demand is inelastic. 

 3. Chicken or the egg

In class, we discussed the supply led and demand led views of transportation and economic development. There are nine major transportation projects listed below. While it is difficult to clearly define such projects as either demand led or supply led, some will display more characteristics of one or the other.

  A. CrossRail – this project is to build two new railway connections under central London. CrossRail 1, approved in 2007, will run east-west and is due to open in 2018. It will complement Thameslink services, which commenced north-south rail services through the re-opened Snow Hill tunnel in 1989.

  B. Channel Tunnel – opened in 1995 and links Britain to France.

  C. M6 Toll motorway – The opening of this tollway around Birmingham, UK, in 2003 effectively provides a Birmingham by-pass and considerably reduced through journey times.

  D. Great Belt Link – opened in 1995 between the island of Zeeland and (in effect) the rest of Denmark, providing a road and rail link.

  E. Golden Gate Bridge – this bridge opened in 1937 across the San Francisco Bay, providing the first fixed link northwards out of San Francisco.

  F. Skye Bridge – this linked mainland Scotland to the island of Skye and was opened in 1995, originally with a highly controversial toll that was eventually abolished at the end of 2004.

  G. Mundra container terminal – this is the first container terminal at the port of Mundra on the Gujarat coast in northwest India, and started its operations in 2005.

  H. Jubilee line extension – this extension opened in 1999 and connected the London Underground system to then developing Docklands area in London

  I. WES Commuter Rail – this diesel-powered commuter rail service began operating in 2009, and extends rail service from Portland to Tigard, Tualatin and Wilsonville for the first time.

   a. Categorize each of these projects into supply or demand led projects, and explain the reasons why you came to that particular view. Note that in many cases you do not need to know the specifics regarding the particular project to come to an educated guess, although some basic Googling of the locations that you are not familiar with might help with your guess.

   b. From the example projects listed, what overall conclusions can you draw regarding supply led and demand led transportation initiatives, particularly the balance between the two?


```{r echo=FALSE}

plot(x=NULL, y = NULL,xlim=c(0,10), ylim=c(0,5),log = "", main = NULL, sub = NULL, xlab ="X1", ylab ="X2",asp =1)
curve((x)/(x-1), 1, 8,col = "red",add = TRUE,);text(5,1, "U1");
curve((x)/(x-2), 2, 8,col = "red",add = TRUE,);text(5,1.5, "U2");
curve((x)/(x-3), 3, 8,col = "violet",add = TRUE,);text(5,2, "U3")

abline(h = 0.5, lty = 2);abline(h = 2, lty = 4)
abline(v =95000, lty=2);text(95000,0.8, "(95000,0.5)")
abline(v =47500, lty=2);text(50000,5.5, "(47500,5.25)");text(47500,0.8, "(47500,0.5)")
abline(v =80000, lty=4);text(80000,2.5, "(80000,2)")
polygon(x=c(47500,95000,47500), y=c(0.5,0.5,5.25), density=10, angle=45,lty =3);text(60000,2.5, "DWL")
abline(v =31558, lty=4);text(35000,8, "(31558,6.8442)")
polygon(x=c(0,31558,0), y=c(6.8442,6.8442,10), density=6, angle=90,lty =3);text(15000,8, "CS")
polygon(x=c(0,31558,31558,0), y=c(0.5,0.5,6.8442,6.8442), density=6, angle=0,lty =3);text(15000,4, "PS")
```