---
title: ""
author: ''
date: ""
output: html_document
---

#  {.tabset .tabset-fade .tabset-pills}


## HW1

### Problem 1: 
In a home economic experiment, two temperatures (factor A) and two lengths of washing time (factor B) gave the following treatment means. The experiment was conducted as a completely randomized design with n=5 for teach treatment combination. The variable of interest was the amount of impurities (in grams) remaining in the material washed. Assume fixed effects for both factors.

```{r echo=FALSE}
library(kableExtra)
Factor <- c("(1)","a","b","ab")
Mean <- c(6.2,14.1,5.4,7.8)
kable(t(data.frame(Factor,Mean)))%>%kable_styling("condensed",full_width = F, position = "center")
```

<!--|     |(1) |a	  |b	 |ab
| --- | -- | -- | -- | --
|Mean	|6.2 |14.1|5.4 |7.8-->

(i) Compute the numerical values of the following effects and interpret in terms of question context:

a. main effect of A

$A=\frac12(-\bar y_{(1)}+\bar y_a-\bar y_b+\bar y_{ab})=\frac12(-6.2+14.1-5.4+7.8)=5.15$

The average amount of impurities remaining in the material washed increase by 5.15 grams as the temperatures changes from low to high. 

b. main effect of B

$B=\frac12(-\bar y_{(1)}-\bar y_a+\bar y_b+\bar y_{ab})=\frac12(-6.2-14.1+5.4+7.8)=-3.55$

The average amount of impurities remaining in the material washed decrease by 3.55 grams as the lengths of washing time changes from short to long. 

c. interaction effect of A and B
	
$AB=\frac12(\bar y_{(1)}-\bar y_a-\bar y_b+\bar y_{ab})=\frac12(6.2-14.1-5.4+7.8)=-2.75$

The average amount of impurities remaining in the material washed decrease by 2.75 grams as the temperatures changes from low to high and the lengths of washing time changes from short to long. 

(ii) Explain which effect is highest from the effects in part (i).

The main effect of A is highest and main effect of B is smaller than that while it is higher than interaction effect.

(iii) Compute the sum of square for each effect in part (i).

$SS_A=\frac{n}4C_A^2=\frac{n}4(-\bar y_{(1)}+\bar y_a-\bar y_b+\bar y_{ab})^2=\frac54(-6.2+14.1-5.4+7.8)^2=132.6125$

$SS_B=\frac{n}4C_B^2=\frac{n}4(-\bar y_{(1)}-\bar y_a+\bar y_b+\bar y_{ab})^2=\frac54(-6.2-14.1+5.4+7.8)^2=63.0125$

$SS_{AB}=\frac{n}4C_{AB}^2=\frac{n}4(\bar y_{(1)}-\bar y_a-\bar y_b+\bar y_{ab})^2=\frac54(6.2-14.1-5.4+7.8)^2=37.8125$

```{r eval=FALSE, include=FALSE}
(14.1+7.8-6.2-5.4)^2*5/4
(5.4+7.8-6.2-14.1)^2*5/4
(7.8+6.2-14.1-5.4)^2*5/4
29.8*16+132.6125+63.0125+37.8125

132.6125/29.8
63.0125/29.8
37.8125/29.8

df(132.6125/29.8, 1, 16,log= FALSE)
df(63.0125/29.8, 1, 16,log= FALSE)
df(37.8125/29.8, 1, 16,log= FALSE)

pf(132.6125/29.8, 1, 16, lower.tail =F, log.p = FALSE)
pf(63.0125/29.8, 1, 16, lower.tail =F, log.p = FALSE)
pf(37.8125/29.8, 1, 16, lower.tail =F, log.p = FALSE)

1-pf(132.6125/29.8, 1, 16, lower.tail = TRUE, log.p = FALSE)
1-pf(63.0125/29.8, 1, 16, lower.tail = TRUE, log.p = FALSE)
1-pf(37.8125/29.8, 1, 16, lower.tail = TRUE, log.p = FALSE)
```

(iv) If the MSE = 29.8, then compute all the terms of the ANOVA table for this study.

|Source of Variation|   SS          |      df      | MS     | F      |P-value   |
|         ---       |   ---         |   ---        | ---    | ---    | ---      |
|Factor A           |132.6125       |   $2-1=1$    |132.6125|$\frac{132.6125}{29.8}=4.450084$|0.0510038 |
|Factor B           | 63.0125       |   $2-1=1$    | 63.0125|$\frac{63.0125}{29.8}=2.114513$|0.1652436 |
|Interaction AB     | 37.8125       |$(2-1)(2-1)=1$| 37.8125|$\frac{37.8125}{29.8}=1.268876$|0.2765897 |
|Error              |$29.8*16=476.8$|$4*(5-1)=16$  | 29.8   |
|     |$132.6125+63.0125+37.8125$|
|Total|$+476.8=710.2375$|$2^2*5-1=19$  |

(v) Based on your ANOVA table, explain which of effects in part (i) are statistically significant at 5% significance level.

Based on the ANOVA table, all the P-value are larger than 0.05. There is no factor has significant effects at 5% significance level.

### Problem 2: 

(i) Create a table of plus and minus for a $2^5$ factorial design. Make sure to use standard order of letter assignments for treatment combinations.

<!--|     | A	| B	| C | D | E | AB | AC | AD | AE | BC | BD | BE | CD | CE | DE | ABC | ABD | ABE | ACD | ACE | ADE | BCD | BCE | BDE | CDE | ABCD | ABCE | ABDE | ACDE | BCDE | ABCDE |
| --- | --| --| --| --| --| -- | -- | -- | -- | -- | -- | -- | -- | -- | -- | --  | --  | --  | --  | --  | --  | --  | --  | --  | --  |  --  |  --  |  --  |  --  |  --  |  --   |
|(1)  | - | - | - | - | - | +  | +  | +  | +  | +  | +  | +  | +  | +  | +  | -   | -   | -   | -   | -   | -   | -   | -   | -   | -   |  +   |  +   |  +   |  +   |  +   |  -    |
|a    | + | - | - | - | - | -  | -  | -  | -  | +  | +  | +  | +  | +  | +  | +   | +   | +   | +   | +   | +   | -   | -   | -   | -   |  -   |  -   |  -   |  -   |  +   |  +    |
|b    | - | + | - | - | - | -  | +  | +  | +  | -  | -  | -  | +  | +  | +  | +   | +   | +   | -   | -   | -   | +   | +   | +   | -   |  -   |  -   |  -   |  +   |  -   |  +    |
|ab   | + | + | - | - | - | +  | -  | -  | -  | -  | -  | -  | +  | +  | +  | -   | -   | -   | +   | +   | +   | +   | +   | +   | -   |  +   |  +   |  +   |  -   |  -   |  -    |
|c    | - | - | + | - | - | +  | -  | +  | +  | -  | +  | +  | -  | -  | +  | +   | -   | -   | +   | +   | -   | +   | +   | -   | +   |  -   |  -   |  +   |  -   |  -   |  +    |
|ac   | + | - | + | - | - | -  | +  | -  | -  | -  | +  | +  | -  | -  | +  | -   | +   | +   | -   | -   | +   | +   | +   | -   | +   |  +   |  +   |  -   |  +   |  -   |  -    |
|bc   | - | + | + | - | - | -  | -  | +  | +  | +  | -  | -  | -  | -  | +  | -   | +   | +   | +   | +   | -   | -   | -   | +   | +   |  +   |  +   |  -   |  -   |  +   |  -    |
|abc  | + | + | + | - | - | +  | +  | -  | -  | +  | -  | -  | -  | -  | +  | +   | -   | -   | -   | -   | +   | -   | -   | +   | +   |  -   |  -   |  +   |  +   |  +   |  +    |
|d    | - | - | - | + | - | +  | +  | -  | +  | +  | -  | +  | -  | +  | -  | -   | +   | -   | +   | -   | +   | +   | -   | +   | +   |  -   |  +   |  -   |  -   |  -   |  +    |
|ad   | + | - | - | + | - | -  | -  | +  | -  | +  | -  | +  | -  | +  | -  | +   | -   | +   | -   | +   | -   | +   | -   | +   | +   |  +   |  -   |  +   |  +   |  -   |  -    |
|bd   | - | + | - | + | - | -  | +  | -  | +  | -  | +  | -  | -  | +  | -  | +   | -   | +   | +   | -   | +   | -   | +   | -   | +   |  +   |  -   |  +   |  -   |  +   |  -    |
|cd   | - | - | + | + | - | +  | -  | -  | +  | -  | -  | +  | +  | -  | -  | +   | +   | -   | -   | +   | +   | -   | +   | +   | -   |  +   |  -   |  -   |  +   |  +   |  -    |
|abd  | + | + | - | + | - | +  | -  | +  | -  | -  | +  | -  | -  | +  | -  | -   | +   | -   | -   | +   | -   | -   | +   | -   | +   |  -   |  +   |  -   |  +   |  +   |  +    |
|acd  | + | - | + | + | - | -  | +  | +  | -  | -  | -  | +  | +  | -  | -  | -   | -   | +   | +   | -   | -   | -   | +   | +   | -   |  -   |  +   |  +   |  -   |  +   |  +    |
|bcd  | - | + | + | + | - | -  | -  | -  | +  | +  | +  | -  | +  | -  | -  | -   | -   | +   | -   | +   | +   | +   | -   | -   | -   |  -   |  +   |  +   |  +   |  -   |  +    |
|abcd | + | + | + | + | - | +  | +  | +  | -  | +  | +  | -  | +  | -  | -  | +   | +   | -   | +   | -   | -   | +   | -   | -   | -   |  +   |  -   |  -   |  -   |  -   |  -    |
|e    | - | - | - | - | + | +  | +  | +  | -  | +  | +  | -  | +  | -  | -  | -   | -   | +   | -   | +   | +   | -   | +   | +   | +   |  +   |  -   |  -   |  -   |  -   |  +    |
|ae   | + | - | - | - | + | -  | -  | -  | +  | +  | +  | -  | +  | -  | -  | +   | +   | -   | +   | -   | -   | -   | +   | +   | +   |  -   |  +   |  +   |  +   |  -   |  -    |
|be   | - | + | - | - | + | -  | +  | +  | -  | -  | -  | +  | +  | -  | -  | +   | +   | -   | -   | +   | +   | +   | -   | -   | +   |  -   |  +   |  +   |  -   |  +   |  -    |
|ce   | - | - | + | - | + | +  | -  | +  | -  | -  | +  | -  | -  | +  | -  | +   | -   | +   | +   | -   | +   | +   | -   | +   | -   |  -   |  +   |  -   |  +   |  +   |  -    |
|de   | - | - | - | + | + | +  | +  | -  | -  | +  | -  | -  | -  | -  | +  | -   | +   | +   | +   | +   | -   | +   | +   | -   | -   |  -   |  -   |  +   |  +   |  +   |  -    |
|abe  | + | + | - | - | + | +  | -  | -  | +  | -  | -  | +  | +  | -  | -  | -   | -   | +   | +   | -   | -   | +   | -   | -   | +   |  +   |  -   |  -   |  +   |  +   |  +    |
|ace  | + | - | + | - | + | -  | +  | -  | +  | -  | +  | -  | -  | +  | -  | -   | +   | -   | -   | +   | -   | +   | -   | +   | -   |  +   |  -   |  +   |  -   |  +   |  +    |
|ade  | + | - | - | + | + | -  | -  | +  | +  | +  | -  | -  | -  | -  | +  | +   | -   | -   | -   | -   | +   | +   | +   | -   | -   |  +   |  +   |  -   |  -   |  +   |  +    |
|bce  | - | + | + | - | + | -  | -  | +  | -  | +  | -  | +  | -  | +  | -  | -   | +   | -   | +   | -   | +   | -   | +   | -   | -   |  +   |  -   |  +   |  +   |  -   |  +    |
|bde  | - | + | - | + | + | -  | +  | -  | -  | -  | +  | +  | -  | -  | +  | +   | -   | -   | +   | +   | -   | -   | -   | +   | -   |  +   |  +   |  -   |  +   |  -   |  +    |
|cde  | - | - | + | + | + | +  | -  | -  | -  | -  | -  | -  | +  | +  | +  | +   | +   | +   | -   | -   | -   | -   | -   | -   | +   |  +   |  +   |  +   |  -   |  -   |  +    |
|abce | + | + | + | - | + | +  | +  | -  | +  | +  | -  | +  | -  | +  | -  | +   | -   | +   | -   | +   | -   | -   | +   | -   | -   |  -   |  +   |  -   |  -   |  -   |  -    |
|abde | + | + | - | + | + | +  | -  | +  | +  | -  | +  | +  | -  | -  | +  | -   | +   | +   | -   | -   | +   | -   | -   | +   | -   |  -   |  -   |  +   |  -   |  -   |  -    |
|acde | + | - | + | + | + | -  | +  | +  | +  | -  | -  | -  | +  | +  | +  | -   | -   | -   | +   | +   | +   | -   | -   | -   | +   |  -   |  -   |  -   |  +   |  -   |  -    |
|bcde | - | + | + | + | + | -  | -  | -  | -  | +  | +  | +  | +  | +  | +  | -   | -   | -   | -   | -   | -   | +   | +   | +   | +   |  -   |  -   |  -   |  -   |  +   |  -    |
|abcde| + | + | + | + | + | +  | +  | +  | +  | +  | +  | +  | +  | +  | +  | +   | +   | +   | +   | +   | +   | +   | +   | +   | +   |  +   |  +   |  +   |  +   |  +   |  +    | 
-->

```{r echo=FALSE, message=FALSE,}
library(dplyr)
five_factor <- tribble(
~effect, ~A, ~B, ~C, ~D, ~E, ~AB, ~AC, ~AD, ~AE, ~BC, ~BD, ~BE, ~CD, ~CE, ~DE, ~ABC, ~ABD, ~ABE, ~ACD, ~ACE, ~ADE, ~BCD, ~BCE, ~BDE, ~CDE, ~ABCD, ~ABCE, ~ABDE, ~ACDE, ~BCDE, ~ABCDE,
"1" , "[-]", "[-]", "[-]", "[-]", "[-]", "[+]", "[+]", "[+]", "[+]", "[+]", "[+]", "[+]", "[+]", "[+]", "[+]", "[-]", "[-]", "[-]", "[-]", "[-]", "[-]", "[-]", "[-]", "[-]", "[-]","[+]","[+]","[+]","[+]","[+]","[-]" ,
"a"   , "[+]", "[-]", "[-]", "[-]", "[-]", "[-]", "[-]", "[-]", "[-]", "[+]", "[+]", "[+]", "[+]", "[+]", "[+]", "[+]", "[+]", "[+]", "[+]", "[+]", "[+]", "[-]", "[-]", "[-]", "[-]","[-]","[-]","[-]","[-]","[+]","[+]" ,
"b"   , "[-]", "[+]", "[-]", "[-]", "[-]", "[-]", "[+]", "[+]", "[+]", "[-]", "[-]", "[-]", "[+]", "[+]", "[+]", "[+]", "[+]", "[+]", "[-]", "[-]", "[-]", "[+]", "[+]", "[+]", "[-]","[-]","[-]","[-]","[+]","[-]","[+]" ,
"ab"  , "[+]", "[+]", "[-]", "[-]", "[-]", "[+]", "[-]", "[-]", "[-]", "[-]", "[-]", "[-]", "[+]", "[+]", "[+]", "[-]", "[-]", "[-]", "[+]", "[+]", "[+]", "[+]", "[+]", "[+]", "[-]","[+]","[+]","[+]","[-]","[-]","[-]" ,
"c"   , "[-]", "[-]", "[+]", "[-]", "[-]", "[+]", "[-]", "[+]", "[+]", "[-]", "[+]", "[+]", "[-]", "[-]", "[+]", "[+]", "[-]", "[-]", "[+]", "[+]", "[-]", "[+]", "[+]", "[-]", "[+]","[-]","[-]","[+]","[-]","[-]","[+]" ,
"ac"  , "[+]", "[-]", "[+]", "[-]", "[-]", "[-]", "[+]", "[-]", "[-]", "[-]", "[+]", "[+]", "[-]", "[-]", "[+]", "[-]", "[+]", "[+]", "[-]", "[-]", "[+]", "[+]", "[+]", "[-]", "[+]","[+]","[+]","[-]","[+]","[-]","[-]" ,
"bc", "[-]", "[+]", "[+]", "[-]", "[-]", "[-]", "[-]", "[+]", "[+]", "[+]", "[-]", "[-]", "[-]", "[-]", "[+]", "[-]", "[+]", "[+]", "[+]", "[+]", "[-]", "[-]", "[-]", "[+]", "[+]","[+]","[+]","[-]","[-]","[+]","[-]" ,
"abc", "[+]", "[+]", "[+]", "[-]", "[-]", "[+]", "[+]", "[-]", "[-]", "[+]", "[-]", "[-]", "[-]", "[-]", "[+]", "[+]", "[-]", "[-]", "[-]", "[-]", "[+]", "[-]", "[-]", "[+]", "[+]","[-]","[-]","[+]","[+]","[+]","[+]" ,
"d" , "[-]", "[-]", "[-]", "[+]", "[-]", "[+]", "[+]", "[-]", "[+]", "[+]", "[-]", "[+]", "[-]", "[+]", "[-]", "[-]", "[+]", "[-]", "[+]", "[-]", "[+]", "[+]", "[-]", "[+]", "[+]","[-]","[+]","[-]","[-]","[-]","[+]" ,
"ad", "[+]", "[-]", "[-]", "[+]", "[-]", "[-]", "[-]", "[+]", "[-]", "[+]", "[-]", "[+]", "[-]", "[+]", "[-]", "[+]", "[-]", "[+]", "[-]", "[+]", "[-]", "[+]", "[-]", "[+]", "[+]","[+]","[-]","[+]","[+]","[-]","[-]" ,
"bd", "[-]", "[+]", "[-]", "[+]", "[-]", "[-]", "[+]", "[-]", "[+]", "[-]", "[+]", "[-]", "[-]", "[+]", "[-]", "[+]", "[-]", "[+]", "[+]", "[-]", "[+]", "[-]", "[+]", "[-]", "[+]","[+]","[-]","[+]","[-]","[+]","[-]" ,
"cd", "[-]", "[-]", "[+]", "[+]", "[-]", "[+]", "[-]", "[-]", "[+]", "[-]", "[-]", "[+]", "[+]", "[-]", "[-]", "[+]", "[+]", "[-]", "[-]", "[+]", "[+]", "[-]", "[+]", "[+]", "[-]","[+]","[-]","[-]","[+]","[+]","[-]" ,
"abd", "[+]", "[+]", "[-]", "[+]", "[-]", "[+]", "[-]", "[+]", "[-]", "[-]", "[+]", "[-]", "[-]", "[+]", "[-]", "[-]", "[+]", "[-]", "[-]", "[+]", "[-]", "[-]", "[+]", "[-]", "[+]","[-]","[+]","[-]","[+]","[+]","[+]" ,
"acd", "[+]", "[-]", "[+]", "[+]", "[-]", "[-]", "[+]", "[+]", "[-]", "[-]", "[-]", "[+]", "[+]", "[-]", "[-]", "[-]", "[-]", "[+]", "[+]", "[-]", "[-]", "[-]", "[+]", "[+]", "[-]","[-]","[+]","[+]","[-]","[+]","[+]" ,
"bcd", "[-]", "[+]", "[+]", "[+]", "[-]", "[-]", "[-]", "[-]", "[+]", "[+]", "[+]", "[-]", "[+]", "[-]", "[-]", "[-]", "[-]", "[+]", "[-]", "[+]", "[+]", "[+]", "[-]", "[-]", "[-]","[-]","[+]","[+]","[+]","[-]","[+]" ,
"abcd" , "[+]", "[+]", "[+]", "[+]", "[-]", "[+]", "[+]", "[+]", "[-]", "[+]", "[+]", "[-]", "[+]", "[-]", "[-]", "[+]", "[+]", "[-]", "[+]", "[-]", "[-]", "[+]", "[-]", "[-]", "[-]","[+]","[-]","[-]","[-]","[-]","[-]" ,
"e" , "[-]", "[-]", "[-]", "[-]", "[+]", "[+]", "[+]", "[+]", "[-]", "[+]", "[+]", "[-]", "[+]", "[-]", "[-]", "[-]", "[-]", "[+]", "[-]", "[+]", "[+]", "[-]", "[+]", "[+]", "[+]","[+]","[-]","[-]","[-]","[-]","[+]" ,
"ae", "[+]", "[-]", "[-]", "[-]", "[+]", "[-]", "[-]", "[-]", "[+]", "[+]", "[+]", "[-]", "[+]", "[-]", "[-]", "[+]", "[+]", "[-]", "[+]", "[-]", "[-]", "[-]", "[+]", "[+]", "[+]","[-]","[+]","[+]","[+]","[-]","[-]" ,
"be", "[-]", "[+]", "[-]", "[-]", "[+]", "[-]", "[+]", "[+]", "[-]", "[-]", "[-]", "[+]", "[+]", "[-]", "[-]", "[+]", "[+]", "[-]", "[-]", "[+]", "[+]", "[+]", "[-]", "[-]", "[+]","[-]","[+]","[+]","[-]","[+]","[-]" ,
"ce", "[-]", "[-]", "[+]", "[-]", "[+]", "[+]", "[-]", "[+]", "[-]", "[-]", "[+]", "[-]", "[-]", "[+]", "[-]", "[+]", "[-]", "[+]", "[+]", "[-]", "[+]", "[+]", "[-]", "[+]", "[-]","[-]","[+]","[-]","[+]","[+]","[-]" ,
"de", "[-]", "[-]", "[-]", "[+]", "[+]", "[+]", "[+]", "[-]", "[-]", "[+]", "[-]", "[-]", "[-]", "[-]", "[+]", "[-]", "[+]", "[+]", "[+]", "[+]", "[-]", "[+]", "[+]", "[-]", "[-]","[-]","[-]","[+]","[+]","[+]","[-]" ,
"abe", "[+]", "[+]", "[-]", "[-]", "[+]", "[+]", "[-]", "[-]", "[+]", "[-]", "[-]", "[+]", "[+]", "[-]", "[-]", "[-]", "[-]", "[+]", "[+]", "[-]", "[-]", "[+]", "[-]", "[-]", "[+]","[+]","[-]","[-]","[+]","[+]","[+]" ,
"ace", "[+]", "[-]", "[+]", "[-]", "[+]", "[-]", "[+]", "[-]", "[+]", "[-]", "[+]", "[-]", "[-]", "[+]", "[-]", "[-]", "[+]", "[-]", "[-]", "[+]", "[-]", "[+]", "[-]", "[+]", "[-]","[+]","[-]","[+]","[-]","[+]","[+]" ,
"ade", "[+]", "[-]", "[-]", "[+]", "[+]", "[-]", "[-]", "[+]", "[+]", "[+]", "[-]", "[-]", "[-]", "[-]", "[+]", "[+]", "[-]", "[-]", "[-]", "[-]", "[+]", "[+]", "[+]", "[-]", "[-]","[+]","[+]","[-]","[-]","[+]","[+]" ,
"bce", "[-]", "[+]", "[+]", "[-]", "[+]", "[-]", "[-]", "[+]", "[-]", "[+]", "[-]", "[+]", "[-]", "[+]", "[-]", "[-]", "[+]", "[-]", "[+]", "[-]", "[+]", "[-]", "[+]", "[-]", "[-]","[+]","[-]","[+]","[+]","[-]","[+]" ,
"bde", "[-]", "[+]", "[-]", "[+]", "[+]", "[-]", "[+]", "[-]", "[-]", "[-]", "[+]", "[+]", "[-]", "[-]", "[+]", "[+]", "[-]", "[-]", "[+]", "[+]", "[-]", "[-]", "[-]", "[+]", "[-]","[+]","[+]","[-]","[+]","[-]","[+]" ,
"cde", "[-]", "[-]", "[+]", "[+]", "[+]", "[+]", "[-]", "[-]", "[-]", "[-]", "[-]", "[-]", "[+]", "[+]", "[+]", "[+]", "[+]", "[+]", "[-]", "[-]", "[-]", "[-]", "[-]", "[-]", "[+]","[+]","[+]","[+]","[-]","[-]","[+]" ,
"abce" , "[+]", "[+]", "[+]", "[-]", "[+]", "[+]", "[+]", "[-]", "[+]", "[+]", "[-]", "[+]", "[-]", "[+]", "[-]", "[+]", "[-]", "[+]", "[-]", "[+]", "[-]", "[-]", "[+]", "[-]", "[-]","[-]","[+]","[-]","[-]","[-]","[-]" ,
"abde" , "[+]", "[+]", "[-]", "[+]", "[+]", "[+]", "[-]", "[+]", "[+]", "[-]", "[+]", "[+]", "[-]", "[-]", "[+]", "[-]", "[+]", "[+]", "[-]", "[-]", "[+]", "[-]", "[-]", "[+]", "[-]","[-]","[-]","[+]","[-]","[-]","[-]" ,
"acde" , "[+]", "[-]", "[+]", "[+]", "[+]", "[-]", "[+]", "[+]", "[+]", "[-]", "[-]", "[-]", "[+]", "[+]", "[+]", "[-]", "[-]", "[-]", "[+]", "[+]", "[+]", "[-]", "[-]", "[-]", "[+]","[-]","[-]","[-]","[+]","[-]","[-]" ,
"bcde" , "[-]", "[+]", "[+]", "[+]", "[+]", "[-]", "[-]", "[-]", "[-]", "[+]", "[+]", "[+]", "[+]", "[+]", "[+]", "[-]", "[-]", "[-]", "[-]", "[-]", "[-]", "[+]", "[+]", "[+]", "[+]","[-]","[-]","[-]","[-]","[+]","[-]" ,
"abcde", "[+]", "[+]", "[+]", "[+]", "[+]", "[+]", "[+]", "[+]", "[+]", "[+]", "[+]", "[+]", "[+]", "[+]", "[+]", "[+]", "[+]", "[+]", "[+]", "[+]", "[+]", "[+]", "[+]", "[+]", "[+]","[+]","[+]","[+]","[+]","[+]","[+]"
)

five_factor%>%mutate(ABCD=cell_spec(ABCD,background=ifelse(ABCD=="[-]","#EAFAF1","#f9ece9")))%>%kable(,"html",escape = F) %>%kable_styling("condensed",font_size = 9, full_width = F)

```

(ii) Write a contrast statement for the four factor interaction effect using standard order of letters as we did in class. Assume n replications for each treatment combination.

$$ABCD=\bar y_{(1)}-\bar y_a-\bar y_b+\bar y_{ab}-\bar y_c+\bar y_{ac}+\bar y_{bc}-\bar y_{abc}-\bar y_d+\bar y_{ad}+\bar y_{bd}+\bar y_{cd}-\bar y_{abd}-\bar y_{acd}-\bar y_{bcd}+\bar y_{abcd}$$
$$+\bar y_e-\bar y_{ae}-\bar y_{be}-\bar y_{ce}-\bar y_{de}+\bar y_{abe}+\bar y_{ace}+\bar y_{ade}+\bar y_{bce}+\bar y_{bde}+\bar y_{cde}-\bar y_{abce}-\bar y_{abde}-\bar y_{acde}-\bar y_{bcde}+\bar y_{abcde}$$

### Problem 3:

(a). contruct a normal probability plot of these effects. Then explain and name the effects that are neglegible.

```{r echo=FALSE,message=FALSE, results='hold', out.width='30%'}
# hw1_3 <- tribble(~factors, ~effect, "A",76.95, "B", -67.52,"C",-7.84, "D",-18.73, "AB", -51.32, "AC", 11.69, "AD",9.78, "BC", 20.78, "BD", 14.74, "CD", 1.27, "ABC", -2.82, "ABD", -6.5, "ACD",10.2, "BCD",-7.98, "ABCD",-6.25)

factors <-c("A","B","C","D","AB","AC","AD","BC","BD","CD","ABC","ABD", "ACD","BCD","ABCD")
effect <- c(76.95,-67.52,-7.84,-18.73,-51.32,11.69,9.78,20.78,14.74,1.27,-2.82,-6.5,10.2,-7.98,-6.25)
hw1_3 <- data.frame(factors,effect)

library(ggplot2)
# hw1_3_qq<-ggplot(hw1_3, aes(sample = effect),label=factors) + stat_qq()
# hw1_3_new<-ggplot_build(hw1_3_qq)$data[[1]]
#### hw1_3_new$name<-hw1_3_new$name[order(hw1_3$effect)]
# hw1_3_new%>%mutate(name=hw1_3$factors[order(hw1_3$effect)])%>%
# ggplot(aes(theoretical,sample,label=name))+geom_text()+stat_qq_line(aes(sample = y), line.p = c(0.25, 0.75),linetype=4, col = "red")+ theme_light()
library(gghalfnorm)
gghalfnorm(x = hw1_3$effect,labs = as.character(hw1_3$factors), nlab = 3)+ theme_light()
# library(daewr) 
# halfnorm(hw1_3$effect, names(as.character(hw1_3$factors)), alpha = 0.05)
```

Based on the normal QQ plot, only the effects of $A, B, AB$ are non-negligible. The reset factors gather around the quantile-quantile line. The effects or interaction effects of rest factors are negligible.

(b).Identify a tentative model, based on the plot of the effects in part (a) write the theoretical model fit for these data without neglegible terms.

The single replicate $2^4$ factorial model with fixed effects and replications, and without neglegible terms is:

$$y_{ij}=\mu+\tau_i+\beta_j+(\tau\beta)_{ij}+\varepsilon_{ij}$$
for $i=1,2$ ; $j=1,2$, where 

$\tau_i$ is fixed main effect of $i^{th}$ level of  Factor A;

$\beta_j$ is fixed main effect of $j^{th}$ level of Factor B;

$(\tau\beta)_{ij}$ is fixed interaction effect of $i^{th}$ level of Factor A and $j^{th}$ level of Factor B;

$\varepsilon_{ij}$ is random error for the single replicate Exerimental Unit when $i^{th}$ level of Factor A and $j^{th}$ level of Factor B are applied.

$y_{ij}$ is response for the single replicate EU when  $i^{th}$ level of Factor A and  $j^{th}$ level of Factor B are applied.

Assumptions:

$\varepsilon_{ij}\sim iid N(0,\sigma^2)$; $\sum_i^2\tau_i=0$; $\sum_j^2\beta_j=0$; $\sum_i^2(\tau\beta)_{ij}=0$; $\sum_j^2(\tau\beta)_{ij}=0$