---
title: 'USP537 Final Exam'
author: "Shen Qu"
date: "Winter 2019"
output: html_document
---

#  {.tabset .tabset-fade .tabset-pills}

### Q1

Bid-rent theory says different land users will compete with one another for land close to the city center. Because retailer or producers wish to maximize their profitability. The more accessible an area, the less transportation cost and the more profitable. They are willing to pay more for land close to the CBD and less for land further away from this area.  

A congestion charge will increase the transportation cost. At the same distance from city center, higher transportation cost means less profits. The location-sensitive producer will move closer to city center and the rent of this place will decline. The land use will change to the type of less location sensitive. The urban land use patterns will shrink at each circle and it will results in smaller city size.


### Q2

Loss of ridership is true. 
By law of demand, the quantity demanded of a good falls when the price of the good rises, other things equal. As for how dramatic it is, depend on the price elasticity of demand. If most riders are sensitive for the price, higher elasticity will result larger amount of loss of ridership.

Lower revenues is not true. 
Less ridership means less ticket income and less operation cost. The setting of transit fare often make the marginal revenue less than the marginal cost. Even less ticket income, a higher fare will narrow the gap and increase the profit and producer surplus.  

Moreover, this argument neglect the benefit to society as a result of the transit service. When raising the price, the subsidy decreased while the consumer surplus decreased by a larger amount. There is an efficiency loss (DWL) associated with this change. Additional negative externalities or equity should be considered.

### Q3

From consumer utility maximization theory, consumer will change their choices to maximize their utility.

The substitution effect reflects the effect of the changed relative prices of the two goods, which alters the slope of the budget constraint but leaves the consumer on the same indifference curve. When TriMet fares increase,  Consumer will choose more auto trips to maximize their utility.

The income effect is the phenomenon observed through changes in purchasing power. It reveals the change in quantity demanded brought by a change in real income. Higher TriMet fares means less income and lower purchasing power.  Consumer will choose less transit trips to maximize their utility.

### Q4

The spatial mismatch hypothesis
This hypothesis describes the spatial mismatch between employment opportunities and workers. Kain (1968) found "the distribution of black employment within the metropolitan areas was affected by the pattern of black residences; the intense residential segregation in these two cities affected the spatial distribution of employment for black workers". "In addition to affecting the location of employment, the spatial mismatch hypothesis purports to affect labor force participation, and therefore the level of employment". "constraints on residential patterns increasingly disadvantaged black households in the labor market".

Bring jobs to people
Provide more employment opportunities nearby the low-income residential areas. This policy will improve workers' accessibility for jobs. It makes workers finding jobs easier and reduce their commuting distance.

Bring people to jobs
Provide more transit services between the residential areas and working areas. This policy will improve the workers' mobility for jobs. It makes workers reach their working places easier.

Connecting people to jobs.
Provide some information tools such as LinkedIn to help workers to find jobs. This policy makes worker and employers find each other easier.

Example:
The Industrial Park and airport areas in Portland have more jobs opportunities. A policy should provide convenient and cheap transit services between main residential areas and these working areas. such as more bus routes cover these areas, or a discount monthly pass of Max Red Line to airport for workers. It is the policy direction of "bring people to jobs".


### Q6

$$V_{auto}=-0.05-0.1TT_{auto}-0.12Cost_{parking}+0.00002Income_{HH}$$
$$V_{auto}=-0.05-0.1\times30-0.12\times5.00+0.00002\times60000= $$

$$V_{bus}=-0.06TT_{bus}-0.15WT_{bus}-0.6Fare_{bus}$$
$$V_{bus}=-0.06\times50-0.15\times\frac{20}2-0.6\times2=$$
 
$$P_{auto}=\frac{\mathrm{e}^{V_{auto}}}{\mathrm{e}^{V_{auto}}+\mathrm{e}^{V_{bus}}}=\frac{\mathrm{e}^{-3.29}}{\mathrm{e}^{-3.29}+\mathrm{e}^{-7.1}}=0.9818363=98.2\%$$

96.3%; 3.7%

```{r eval=FALSE, include=FALSE}
-0.05-0.1*30-0.12*5.00+0.00002*60000
-0.06*50-0.15*20/2-0.6*2.00
exp(-2.45)/(exp(-2.45)+exp(-5.7))
exp(-5.7)/(exp(-2.45)+exp(-5.7))

-0.06*30-0.15*20/2-0.6*2.00
exp(-4.5)/(exp(-2.45)+exp(-4.5))

-0.05-0.1*30-0.12*8.00+0.00002*60000
exp(-2.81)/(exp(-2.81)+exp(-5.7))

```

$$VOT =\frac{\frac{\Delta V_{bus}}{\Delta TT_{bus}}}{\frac{\Delta V_{bus}}{\Delta Fare_{bus}}}=\frac{-0.06}{-0.6}\times60=6.00\$/hour$$

### Q7

A bus rapid transit route is more effective.

A bus rapid transit route will decrease bus TT by 20 minutes.  The new bus TT is 30 minutes. The new V(bus) is -4.5. The new probability of bus is 11.4%, which increased 7.7%.

A parking charge will increase \$3 on parking cost. The new parking cost is \$8
The new V(bus) is -2.81. The new probability of auto is 94.7%, which decreased 1.6%.

The results show that the bus rapid transit route can increase more percentage of transit ridership. While higher parking charge only decrease a small percentage of auto trips.

### Q8

BR=b0-50Fare+4Income+15Price+30Summer+5Clousure+ e

300=b0-50*1+4*45 + 15*2+30*Summer+5*Clousure+ e

Price elasticity of demand= $B1*\Delta fare/ \Delta rr=-50*1/300$

income elasticity of demand=$B2*\Delta Income/ \Delta rr=4*45/300$

```{r}
-50*1/300
4*45/300
```

### Q9

Discount rate  is based on interest rates and put all present and future costs and benefits in a common metric, their present value.  When conduct sensitivity analysis, discount rate can help to determine whether and when CBA conclusions are vulnerable to changes in major assumptions.

A high discount rate means a preference for present values, a will to pay less today for the future cash. A low discount rate means a preference for future values and less uncertainty.

### Q10

NSB=Benefit-Cost. If Benefit is greater than Cost, the NSB show a positive evaluation result. Projects with positive net present value should be considered.

BCR=Benefit/Cost. If Benefit is larger than Cost, the BCR is larger than 1 as result. The higher the ratio, the greater the benefits relative to the costs. 

Advantage of BCR: It provides a simple tool to know if a project will success or not.

Disadvantage of BCR: It is insensitive to the magnitude of net benefits. It favors projects with small costs and benefits over those with higher net benefits.

### Q11

Revealed Preference: is a valuation based on actual observable choices.
Stated Preference: is a contingent valuation based on a survey where the respondents state their value.

Revealed preference is more preferable when performing a cost-benefit analysis. 

Because revealed preference has more method to evaluate: Direct methods include Market Price, Benefits Transfer, and Meta-analysis. Indirect methods include Travel Cost Method, Hedonic Analysis, Value of Statistical Life. It has more feasible data sources. It is more reliable because it comes from actual observable choices. 
Stated preference has some shortcoming such as strategic bias, information bias, starting-point bias, hypothetical bias, etc.
