---
title: 'USP537 Problem Sets'
author: ""
date: "Winter 2019"
output: html_document
---

#  {.tabset .tabset-fade .tabset-pills}

## Problem Set 1

 1. Portland’s transit agency TriMet wants to increase transit usage by 15%. They expect no changes in the demand for transit and will lower the fare to attract the additional riders. The estimated price elasticity of demand is 0.3(actual elasticity used by TriMet).

 a. Calculate the percentage change necessary in fares to achieve the increased transit usage and the change in fare box revenue (use simple percentages rather than the midpoint elasticities). 

When no changes in the demand for transit, the value of usage moves along the MD curve, then 

$$\frac{\Delta Q/Q}{\Delta P/P}=E\implies\Delta P/P=\frac{\Delta Q/Q}{E}=\frac{0.15}{0.3}=50\%$$
$$TR=50\%P(1+15\%)Q=57.5\%PQ$$

50% decrease in fares could achive the increase transit usage by 15%. The new revenue will be 57.5% of previous revenue.

 ---
 
 b. Would you expect the long-run price elasticity of demand to be larger or smaller? Explain why?

I expect the long-run price elasticity of demand to be larger. Demand tends to be more elastic in the long rung rather than in the short run, because when prices change consumers often need more time to respond and change their shopping habits such as househould or firms' location, vehicle ownership, further change travel behaviors. (Mankiw, 2004).

 * Reference

_Mankiw (2004) “Principals of Macroeconomics”,South-WesternCollegePub,UK_

 ---
 
 c. Do you expect that costs will rise? Consider total cost, fixed cost, and variable cost. Explain why it would or would not.

If the extra riderships don't exceed current transit capacity, fixed cost would not change. The variable cost will rise for the addtional operation cost such as energy consumption and labor cost. Thus, the total cost would increase. If current transit is underload, the change would be tiny.

 ---
 
 2. In “The Urban Transportation Problem: A Reexamination and Update”, by John F. Kain([GTW] Chap11), the author concludes the paper with a recommendation: “There remains a strong, if not compelling, case to begin charging road users the full social cost of their road use, for reversing the public ownership and operation of transit services, and for stopping wasteful expenditures on ineffective rail systems.” Please write a short analysis (approximately1 page, single-spaced, typed) highlighting your arguments either in support of or against this recommendation.

There are three points in Kain’s statement: charging road users the full social cost, reversing public ownership and operation of transit services, and stopping ineffective rail systems (Kain, 2001 p.396). The three concepts are related. In general, I agree with him and try to make a further discussion.

 - Charging road users the full social cost
 
The principle of road pricing is to balance the cost and externality. Because the road using costs do not reflect true costs to society and environment, users are not paying the full costs of their purchases and even unaware of it. They will tend to use too much that is subsidized by society. The markets are distorted and inefficient.

Road user charges have two functions. The first one is used for revenue generation, which was used to road infrastructure financing. The second function is as a transportation demand management tool, which can reduce peak hour travel, release traffic congestion, and encourage transit development. Kain (2001) believes the later: “In the short run, motorists would make fewer trips, change their time of travel, and alter their destinations…In the long run, higher road user charges would eventually affect the dispersed patterns of workplaces and residences that were in part induced by congestion” (P.393).

The problem is how to get “the right price”? Prices should reflect all the costs and externalities including both negative and positive. Many economists try to measure the lost such as congestion, air pollution, GHG emissions, visual intrusion, noise, and road accidents.  But “economists disagree over how to set tolls, how to cover common costs, what to do with any excess revenues, whether and how "losers" from tolling previously free roads should be compensated, and whether to privatize highways.  It is hard to measure the benefit of road using such as economic growth, the communication of information and technology, the recreation activities, and the emotional need for social lives. Kain’s main idea is charging low-occupancy vehicles and subsidizing transit (p.393). It is also a huge challenge at the operating level, which leads to the second topic: privatization.

The application of congestion charges is currently limited to a small number of cities and urban roads.  The technology of electronic road pricing through ITS is making detailed billing based on actual use feasible and affordable.

 - Reversing public ownership and operation of transit services
 
Transit is a quasi-public good, Off-peak travelers’ consumption does not reduce any other's consumption of that good.  In some time or location, transit is scarce goods. Both excludability and rivalry are possible. Hart (1995) proposes the party with the better investment technology should be the owner , while Besley and Ghatak (2001) argue that the party who has a larger valuation for the public good should be the owner.  Klein, Moore and Reja (1997) point out public transit authorities have tended to over-provide their service. (P.12). It seems that privatization can help to improve efficiency and achieve benefit maximum inside the transportation system, promote the resources redistribution by the hand of market. But Market cannot solute the problem of externalities by itself. The free travel and dispersed patterns might become the wealthy’s privilege. For transit, the scarce time and segments will turn to “club goods,” which means high excludability and low rivalry in consumption. In social justice or equality, the public department still needs to provide alternative goods or subsidies. Especially, mass transit, what is commonly known as natural monopolies, need huge initial investment and have essentially zero marginal costs. It beyond the issues of ownership or operation.

 - The ineffective rail systems
 
Kain (2001) believes busway can close match between demand and capacity with the lower operating costs, (P.379). “Research on mode choice shows that travelers value the following characteristics: short trip times, avoidance of transfers and waiting time, door-to-door service, reliability, comfort, seat availability, storage space, security, and flexibility” (Klein, Moore and Reja, 1997. P.9). In this way, rail system is the most expensive and least flexible, should be stopped.

Other reasons Kain gave are disappointing ridership than projected (p.362), the dispersed employment patterns and low population densities (P.366). It is true that the low density in the U.S. make transit operators difficult to provide competitive services. Meaning while, in some cities with mid-high income and high density such as New York and Hong Kong, a powerful rail system support the cities operation and play an irreplaceable role (Pushkarev and Zupan,1997; Cervero,1998). 

Review the history, “the steady rise in auto ownership and declines in transit use since 1895 were largely efficient responses to changes in technology and rising incomes” (Kain, 2001. P.359). “The transit market is contracting because of natural forces” (Klein, Moore and Reja, 1997. P.16). Therefore, we should be careful when planning a rail in a med-low density area and have to pay the heavy cost of rail system in some cities. Beside advocating BRT and TOD in middle cities or growing area, next generation of technology such as automated driving system could turn the trend in future and lead to an unpredictable land use pattern.

 * Reference

  [1] _Kain, John F. (2001). Essays in transportation economics and policy : a handbook in honor of john r. meyer, Chapter 11. Brookings Institution Press._

  [2] Paul Johnson, Andrew Leicester & George Stoye. (2012). Fuel for Thought – The what, why and how of motoring taxation. Institute for Fiscal Studies and Royal Automobile Club Foundation for Motoring. Executive Summary, pp. v.
  
  [3] Lindsey, Robin (2006). Do Economists Reach a Conclusion on Road Pricing? The Intellectual History of an Idea. Econ Journal Watch. 3 (2): 292–379.
  
  [4] Small, Kenneth A.; Verhoef, Erik T. (2007). The Economics of Urban Transportation. Routledge, New York. pp. 148–153. ISBN 978-0-415-28515-5.
  
  [5] Demsetz, Harold (1970). Full Access The Private Production of Public Goods. Journal of Law and Economics. 13 (2): 293–306. doi:10.1086/466695. JSTOR 229060.
  
  [6] Hart, Oliver (1995). Firms, Contracts, and Financial Structure. Oxford University Press.
  
  [7] Besley, Timothy; Ghatak, Maitreesh (2001). "Government Versus Private Ownership of Public Goods". The Quarterly Journal of Economics. 116 (4): 1343–72. CiteSeerX 10.1.1.584.6739. doi:10.1162/003355301753265598. ISSN 0033-5533.

  [8] Klein, D. B., Moore, A. T., & Reja, B. (1997). Curb rights: a foundation for free enterprise in urban transit. Washington, D.C: Brookings Institution Press.


 ---
 
 3. Rail City operates a train service on weekdays only. The fixed cost is $120,000 per day. The operating cost is $100 per train-hour, with service adjusted to provide 200 rides per train-hour. This yields a marginal cost of $0.50 per rider. Use this where information on marginal cost is needed. The following information is available regarding cost and demand on a daily basis.

|        |Cost|Revenue|Demand
|---     | ---|--- | ---
|Fixed   |$FC=\$120000$| |
|Marginal|$MC=0.5$|$MR=10-0.0002Q$|$P=\$10-0.0001Q$|
|Total   |$TC=FC+VC=\$120000+0.5Q$|$TR=P\times Q=10Q-0.0001Q^2$|

 a. What number of riders will use the system if the price is set equal to marginal cost(This is essentially equivalent to assuming that a benevolent social planner sets the price so that maximum efficiency is reached)? How much service will be provided (Q*)? What are the subsidies or profits at this level of cost and price? What is the consumer surplus?

Let $P^*=MC$, then $10-0.0001Q^*=0.5\implies Q^*=95000$ riders. The service is $95000/200=475$ train-hour.

$TR-TC=(P^*Q^*)–(120000+P^*Q^*)=-\$120000$

$CS^*= \frac12(10-P^*)Q^*=\frac12\times(10-0.5)\times95000=\$451250$

When the price equals marginal cost, there will be 95000 riders and 475 train-hour service will be provided.It needs $120000 subsidies for operation. The consumer surplus reach the maximum \$451250.

 ---
 
 b. If the price is raised to $2.00, how many people will use the train, and what happens to total revenue? What subsidy is required? Calculate any efficiency loss or gain associated with this change. Compare the change in subsidy with the change in consumer surplus.

When $P_2=\$2.00$, there will be $Q_2=\frac{10-2}{0.0001}=80000$ riders.

$TR=P_2Q_2=2\times80000=\$160000$.

$TR-TC=160000–(120000+0.5\times8000)=\$0$, $\Delta(TR-TC)=0-(-120000)=\$120000$

$CS=\frac12(10-P_2)Q_2=\frac12(10-2)80000=320000$, $\Delta{CS}=320000-451250=\$131250$

$PS=(P_2-P^*)Q_2=(2-0.5)80000=120000$

$TS=CS+PS=320000+120000=440000$

$DWL=\frac12\Delta P\Delta Q=\frac12(2.0-0.5)(95000-80000)=\$11250$

When the price equals to \$2, there are 80000 riders. The total revenue is \$160000 and don't need subsidies. But the consumer surplus loss \$131250, which is greater than the decrease in subsidies \$120000. It means we should not raise price for reducing subsidies.

 ---
 
 c. Suppose that a monopolist is allowed to run the system and set price. What quantity ($Q_m$) and price ($P_m$) will be set and what profits or losses will be generated if the monopolist wants to maximize profits? Calculate any efficiency loss associated with this level of service.

When $MR=MC$, the monopolist maximizes profits. $10–0.0002Q_m=0.5\implies Q_m=47500$ riders,  $P_m=10–0.0001Q=10–4.75=\$5.25$

$TR-TC=(P_m-P^*)Q_m-120000=(5.25-0.5)\times47500–120000=\$105625$

$CS=\frac12(10-P_m)Q_m=\frac12(10-5.25)47500=112812.5$

$PS=(P_m-P^*)Q_m=(5.25-0.5)47500=225625$

$TS=CS+PS=112812.5+225625=338437.5$

$DWL=\frac12\Delta P\Delta Q=\frac12(5.25-0.5)(95000-47500)=\$112812.5$

Whne the marginal revenue equals marginal cost, the price is \$5.25 and riders are 47500. The monopolist get the maximum profit of \$105625. However, the total efficientcy lost are \$112812.5. We should not sacrific the publis interest for maximizing the monopolist's profit.

 ---
 
 d. Could the monopolist make a profit if fixed costs are $250,000 per day?  Does this mean that it would be a mistake to build the train based on economic benefits and costs for this level of fixed cost? Explain why you think it would or would not be from an economic efficiency perspective and include appropriate calculations.

Let $FC=\$250000$, the monopolist's maximum profit is $TR-VC-FC=(P_m-P^*)Q_m-250000=-\$24375$.

The monopolist won't make a profit. To examine the efficiency, we should compare subsidies and cunsumer surplus. 

If the price is set to $P^*=0.5$, the total benefit is $CS+TR-TC=451250+(0.5-0.5)95000-250000=\$201250$

If the price is set to $P_m=5.25$, the total benefit is $CS+TR-TC=112812.5-24375=\$88437.5$

Let $TS+TR-TC=\frac12(10+P-2P^*)Q-250000=-0.00005Q^2+9.5Q–250000=0$, $Q_{1,2}=\frac{-b\pm \sqrt{b^2-4ac}}{2a}=\frac{9.5\pm\sqrt{40.25}}{0.0001}$, $Q_1=31558,Q_2=158442$. For $Q\le95000$, so $Q=3155711,P=6.844289$.

Then $CS+TR-TC=49792.56-49792.57=\$0$. Once the price is lower than \$6.844289, the riders are more than 31558, the cunsumer surplus will be larger than subsidies.

The reults prove that the train is worth to be bulit because social benefit are positive rather than considering the profit of monopolist. Only if the riders are less than 31558, building this train is a mistake for a negative social benefit.


|P(\$)|Rider|Fixed Cost|Variable Cost|Total Cost||Revenue|subsidies/profit||Consumer Surplus|Producer Surplus|Total Surplus        | DWL                     |
|---  | --- | ---      |---          | ---      |---| ---| ---       | --- |---             | ---            | ---                 |---                      |
|$P$  |$Q$  |$FC$      |$P^*Q$       |$FC+VC$   ||$PQ$   |$(P-P^*)Q-FC$   ||$\frac12(10-P)Q$|$(P-P^*)Q$      |$\frac12(10+P-2P^*)Q$|$\frac12\Delta P\Delta Q$|
|0.50 |95000|120000    |47500        |167500    ||47500  | -120000        ||451250          | 0              |451250               |0                        |
|2.00 |80000|120000    |40000        |160000    ||160000 | 0              ||320000          | 120000         |440000               |-11250                   |
|5.25 |47500|120000    |23750        |143750    ||249375 | 105625         ||112812.5        | 225625         |338437.5             |-112812.5                |
|0.50 |95000|250000    |47500        |297500    ||47500  | -250000        ||451250          | 0              |451250               |0                        |
|5.25 |47500|250000    |23750        |273750    ||249375 | -24375         ||112812.5        | 225625         |338437.5             |-112812.5                |
|6.84 |31558|250000    |15779        |265779    ||215986 | -49792.57      ||49792.56        | 200207.4       |250000               |-201250                  |

```{r echo=FALSE}
curve(10-0.0001*x, 0, 100000,col = "red",xlab ="Q", ylab ="P");text(40000,7, "MD")
curve(10-0.0002*x, add = TRUE, col = "violet");text(40000,3, "MR")
abline(h = 0.5, lty = 2);abline(h = 2, lty = 4)
abline(v =95000, lty=2);text(95000,0.8, "(95000,0.5)")
abline(v =47500, lty=2);text(50000,5.5, "(47500,5.25)");text(47500,0.8, "(47500,0.5)")
abline(v =80000, lty=4);text(80000,2.5, "(80000,2)")
polygon(x=c(47500,95000,47500), y=c(0.5,0.5,5.25), density=10, angle=45,lty =3);text(60000,2.5, "DWL")
abline(v =31558, lty=4);text(35000,8, "(31558,6.8442)")
polygon(x=c(0,31558,0), y=c(6.8442,6.8442,10), density=6, angle=90,lty =3);text(15000,8, "CS")
polygon(x=c(0,31558,31558,0), y=c(0.5,0.5,6.8442,6.8442), density=6, angle=0,lty =3);text(15000,4, "PS")
```

```{r eval=FALSE, include=FALSE}
uniroot(function(x)(-0.00005*x^2+9.5*x-250000), lower=0, upper=100000)
```



 