---
title: 'USP654 Lab 6: Working with categorical variables'
author: "Liming Wang"
date: "10/30/2017"
output: 
  html_document:
    toc: true
---

```{r setup, include=F}
knitr::opts_chunk$set(message=FALSE, warning=F, echo=TRUE)
options(width = 2000)
options(repos="https://cran.rstudio.com")
```

## Testing Differences in Means Among the Included Groups

[Download california.csv](californiatod.csv)

1. Redefine reference category, with `relevel` in R:
```{r}
require(tidyverse)
californiatod <- read_csv("californiatod.csv")

# redefine the region variables with a new reference category (4 for SD)
catod2 <- californiatod %>% 
  mutate(region = relevel(as.factor(region), ref = 4))
lm(houseval ~ region, data=catod2)  %>% 
  summary
```

**Task**: Can you redefine LA as the reference category?

2. Partial F test:

```{r}
# redefine the region variables where LA_SD becomes a combine region

catod3 <- californiatod %>% 
  mutate(region = ifelse(region =="LA" | region == "SD", "LA_SD", region))
lm(houseval ~ region, data=catod3)  %>% 
  summary

# And then do a partial F test against the model before redefining region

anova(
  lm(houseval ~ region, data=catod3),
  lm(houseval ~ region, data=californiatod)
)
```

3. Hypothesis testing of linear combination of coefficients

```{r}
car::lht(lm(houseval ~ region, data=californiatod), "regionLA = regionSD")

# OR

car::lht(lm(houseval ~ region, data=californiatod), "regionLA - regionSD = 0")
```

## Ordinal Categorical Variables

```{r, results="asis"}
fit_numeric <- lm(mpg ~ cyl, data=mtcars)
fit_factor <- lm(mpg ~ factor(cyl), data=mtcars)
# sdif - sequential difference
fit_ordered <- lm(mpg ~ ordered(cyl), data=mtcars, 
                  contrasts=list(`ordered(cyl)`=MASS::contr.sdif)) 

require(huxtable)
huxreg(fit_numeric, fit_factor, fit_ordered)
```

## Interaction Terms

In R, you can easily create interaction terms by multiplying (*) two variables. In the results, R uses : to indicate the coefficients for interaction terms.

```{r}
huxtable::huxreg(
  lm(houseval ~ region + railtype, data=californiatod),
  lm(houseval ~ region * railtype, data=californiatod),
  lm(houseval ~ transit * region, data=californiatod),
  lm(houseval ~ transit * railtype, data=californiatod),
  lm(houseval ~ transit * density, data=californiatod)
  )
```

In other statistical software, you likely need to create your own interaction terms by multiplying two variables and assigning the value to a new variable manually.

## Tasks

1. ~~Identify the categorical variables in your own dataset~~
2. Select a few of the categorical variables
    - How is each of them code (dummy, numeric, or character)?
    - Recode the categorical variables that are not properly coded currently;
    - Consider combine/recode some of the categories for one of your categorical variables;
    - Consider how you would create interaction terms if your software cannot create them for you?
    - Run regression models including your categorical variables as independent variables; think about how you interpret the results of your regression;
    - Test interaction effects: among your categorical variables, between categorical and numeric variables, and among numeric variables.

