---
title: 'USP654 Lab 10: Generalized Linear Models'
date: "11/27/2017"
output: 
  html_document:
    toc: true
---

```{r setup, include=F}
knitr::opts_chunk$set(message=FALSE, warning=F, echo=TRUE)
options(width = 2000)
options(repos="https://cran.rstudio.com")
```

## Discrete Outcome: Count/Poisson Regression

`mtcars` example: Number of carburetors ~ f(x1, x2, ...)

- lm: lm(carb ~ mpg + qsec, data=mtcars)
- glm: glm(carb ~ mpg + qsec, data=mtcars, family="poisson")

```{r, echo=TRUE}
require(MASS)
require(huxtable)
fit_lm <- lm(carb ~ mpg + qsec, data=mtcars)
fit_glm <- glm(carb ~ mpg + qsec, data=mtcars, family="poisson")
huxreg(OLS=fit_lm, Poisson=fit_glm)
```

## Dichotomous (Binomial) Outcome: Logistic Regression

- `mtcars` example

am ~ f(x):

- lm: lm(am ~ qsec + hp + cyl, data=mtcars)
- glm: glm(am ~ qsec + hp + cyl, data=mtcars, family=binomial("logit"))

```{r, echo=TRUE}
require(MASS)
require(huxtable)
require(ggplot2)
fit_lm <- lm(am ~ qsec + hp, data=mtcars)
fit_glm <- glm(am ~ qsec + hp, data=mtcars, family=binomial("logit"))
huxreg(OLS=fit_lm, logit=fit_glm)
```

- log Likelihood

```{r}
logLik(fit_glm)

fit_glm0 <- update(fit_glm, .~1)
logLik(fit_glm0)


# pseudo R2
1 - logLik(fit_glm)/logLik(fit_glm0)
```

- Interpretation of coefficients

```{r, echo=TRUE}
# odds ratio
(odds <- exp(coef(fit_glm)))

#prob
odds/(1 + odds)
```


## Tasks

1. Create a model for a count outcome in your dataset;
1. Create a model for a Dichotomous outcome in your dataset.
