---
title: "Statistical Literature and Problems"
subtitle: "STAT 501"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document: default
  pdf_document: default
header-includes:
    - \usepackage{amsmath}
---


```{r setup, include=F}
knitr::opts_chunk$set(message=FALSE, warning=F, echo=TRUE)
options(width = 2000)
options(repos="https://cran.rstudio.com")
```

#  {.tabset .tabset-fade .tabset-pills}

## (Albert &  Chib, 1993)

James H. Albert & Siddhartha Chib (1993) Bayesian Analysis of Binary and Polychotomous Response Data, Journal of the American Statistical Association, 88:422, 669-679, DOI: 10.1080/01621459.1993.10476321


### 5.1 Finney Data

```{r,echo=F,out.width='50%'}
# library(likelihoodAsy)
library(tidyverse)
library(robcbi)
data(Finney)
```

```{r,echo=F,out.width='50%'}
Finney

Vol <- Finney$Vol; Rate <- Finney$Rate; Resp <- Finney$Resp
lVol <-log(Vol); lRate <- log(Rate)
```

```{r,echo=F,out.width='50%'}
z.glm <- glm(Resp~lVol+lRate,family=binomial)
summary(z.glm)
covar(z.glm)
Rank(z.glm)
rscale(z.glm)
weights(z.glm)


z.cub <- glm(Resp~lVol+lRate,family=binomial,method="cubinf", ufact=3.2)
summary(z.cub)
# plot(z.cub, smooth=TRUE, ask=TRUE)
```

```{r,echo=F,out.width='50%'}
## Plots for comparing fits
comp <- fits.compare(z.glm,z.cub)
comp
# plot(comp)
```

```{r,echo=F,out.width='50%'}
## Adds a QQ-line for the values in x in the current plot.
x <- residuals(z.cub, type="deviance")
qqnorm(x, ylab = "Deviance Residuals")
QQline(x, lty = 2)
# Predictions provided by a model fit when method is "cubinf".
rVol <- runif(20,0.4,3.7); rRate <- runif(20,0.3,3.75)
newdat <- data.frame(lVol=log(rVol),lRate=log(rRate))
predict(z.cub, newdat, type="response")

```


### 5.2 Election Data


### 5.3 A Trivariate Probit Example



## (Polson etal, 2013)

Nicholas G. Polson, James G. Scott & Jesse Windle (2013) Bayesian Inference for Logistic Models Using Pólya–Gamma Latent Variables, Journal of the American Statistical Association, 108:504, 1339-1349, DOI: 10.1080/01621459.2013.829001