---
title: 'USP654 Lab 2: Bivariate Regression'
author: "Liming Wang"
date: "9/25/2017"
output: 
  html_document:
    toc: true
---

```{r setup, include=F}
knitr::opts_chunk$set(message=FALSE, warning=F, echo=TRUE)
options(width = 2000)
options(repos="https://cran.rstudio.com")
```


## RStudio Project

I recommend you organize your data analysis project with RStudio projects. One data analysis project per Studio project. (Stick to a similar practice even if you're using SPSS.) 

Launch RStudio, go to **File/New Project...** menu, select **New Directory** (if you already created your directory, Existing Directory) and **New Project**, type in a directory name and then click then **Create Project** button. You can use "usp654_project" as your directory name or choose a more expressive name.

Some tips of managing your RStudio project:
1. Create a *data* folder inside your project directory for raw data files (never change data files manually).
2. Create a *code* folder inside your project directory for R script. It is better to work with a script file than using the R console for substantial work (for example, your homework or class project). R console is mostly for testing to make sure your scripts work.
3. Create an *output* folder inside your project directory for any outputs from your R scripts you would like to save for later use, which include data, figures, tables, etc.
4. (Optional) Create an *docs* folder inside your project directory for any documents.

More details of the best practice of [Project Management With RStudio](http://swcarpentry.github.io/r-novice-gapminder/02-project-intro/).

## R Package

To install R packages on to the computers in the lab, you will need to down the [.Rprofile](.Rprofile) file (Save link as...) and save it to your Documents folder.

One of the advantages of R is the thousands packages that extend the functionailities of base R. A major part of learning R is to learn to work with R packages. Besides the base R, we will use a suite of packages named [`tidyverse`](https://www.tidyverse.org/) that covers the whole process of data analysis. To install the `tidyverse` packages, you can use the **Packages** pane or run R function in RStudio's **Console** pane: 

```{r, eval=FALSE}
install.packages("tidyverse")
```

To use R packages you have installed, include this line in your script or in the Console:

```{r, echo=TRUE}
library("tidyverse")
# Note here the quotation marks are optional and
# it is the same as
library(tidyverse)
```

## Import Data

1. Download the data file for this lab from [here](mtcars.csv) and save it to data directory of your RStudio project.
2. What format is the data file in?
3. Import the data in the file into R as a data frame:

```{r, echo=TRUE}
my_df <- read_csv("data/mtcars.csv")
my_df
```

Of course, the step of import data into your statistic software vary a little by software and by the data format. 

More information as of how to import data into R can be found at [Import Data]()

## Descriptive Statistics

### Visualization

1. Visualize a single variable
```{r, echo=TRUE}
ggplot(my_df, aes(x=mpg)) + geom_histogram()
```

```{r, echo=TRUE}
ggplot(my_df, aes(x=wt)) + geom_histogram()
```

2. Visualize a pair of numeric variables
```{r, echo=TRUE}
ggplot(my_df, aes(x=wt, y=mpg)) + geom_point()
```

## Correlation

```{r, echo=TRUE}
cor(my_df$mpg, my_df$wt)
```

## Regression
In R, run linear regressions with `lm` (short for **l**inear **m**odel):

```{r, echo=TRUE}
lm(mpg ~ wt, data=my_df)
```

## More detailed regression results

1. Pass the results from `lm()` to `summary()` for more detailed information:
```{r, echo=TRUE}
lm(mpg ~ wt, data=my_df) %>%
  summary
```

For better formatting of the results (pretty print), we can use the `texreg` package:
```{r, echo=TRUE}
## Install and load texreg package
install.packages("texreg")
library(texreg)

# Pretty print regression results on screen
lm(mpg ~ wt, data=my_df) %>%
  screenreg

# Save regression results to a html file
lm(mpg ~ wt, data=my_df) %>%
  htmlreg(file="output/lm_mpg-wt.html")
```

## Visualize regression results

```{r, echo=TRUE}
ggplot(my_df, aes(x=wt, y=mpg)) + geom_point() + geom_smooth(method = "lm", se = FALSE)
```

## Diagnostic plots of regression

We will use the `ggfortify` package to generate the diagnostic plots for regression
```{r, echo=TRUE}
## Install and load ggfortify package
install.packages("ggfortify")
library(ggfortify)

lm(mpg ~ wt, data=my_df) %>%
  autoplot()

# Since we need the regression results in many places, 
# it would be easier to save it into a R variable
mpg_lm <- lm(mpg ~ wt, data=my_df)

# Save the diagnostic plots as a png file
ggsave("output/mpg_lm_diag.png")

```

## Tasks

1. Create a new RStudio project and create directory structure following the recommendation in [Project Management With RStudio](http://swcarpentry.github.io/r-novice-gapminder/02-project-intro/);
1. Select a dataset from the [Tools and Datasets] section of the course website or use your own dataset, save the dataset in the data folder of your project;
1. Famarlize yourself with your dataset with the data document;
1. Create a script that imports your dataset into R and explores it with descriptive statistics and/or visulization;
1. Choose a dependent variable and a couple of independent variables and run regression models;
1. Create regression plot (scatter plot superimposed with a regression line) & diagnostic plots.

## Resources
- [Introduction to R on Data Camp](https://campus.datacamp.com/courses/free-introduction-to-r/chapter-1-intro-to-basics-1?ex=1): A self-instruction course covering R basics.
- [R coding basics on Software Carpentry](http://swcarpentry.github.io/r-novice-gapminder/), hands-on tutorial on R basics, lesson 4-11.
