---
title: 'STAT564 Lab 0: Introduction to R for Simple Linear Regression Analysis'
author: "Shen Qu"
date: "9/21/2018"
output: 
  html_document:
    toc: true
---

```{r setup, include=F}
knitr::opts_chunk$set(message=FALSE, warning=F, echo=TRUE)
options(width = 2000)
options(repos="https://cran.rstudio.com")
```


## EXAMPLE

These data appeared in the Wall Street Journal. The advertisement were selected by an annual survey conducted by Video Board Tests, Inc., a New York ad-testing company, based on interviews with 20,000 adults who were asked to name the most outstanding TV commercial they had seen, noticed, and liked. The retained impressions were based on a survey of 4,000 adults, in which regular product users were asked to cite a commercial they had seen for that product category in the past week. There are 21 observations on three variables. 

**Variables:**
Firm The name of the firm
Spend TV advertising budget, 1983 ($ millions)
MilImp Millions of retained impressions per week

These data is saved in text file (.txt) called **TVads** which will be imported to R as shown in th efollowing code. The path to the place the text file is saved must be changed accordingly. 
```{r, echo=TRUE}
TVads <- read.table("TVads.txt", header= TRUE) 
head(TVads) # This will print only the first 6 rows of data file 
TVads # This will print the complete data file in the Console Window 

# The following plot function creates the scatterplot shown in Figure 1 below
plot(TVads$Spend, TVads$MilImp, pch=16, xlab = "TV advertising budget, 1983 ($ millions) ", ylab = "Millions of retained impressions per week", main="Figure 1: Scatterplot of Millions of retained impressions per week and TV advertising budget")
```



Simple linear regression analysis in R 
```{r, echo=TRUE}
model <- lm(MilImp ~ Spend, data = TVads) # The lm function is used to fit simple linear regression model
model# This displays only the values of estimated regression coefficients
```

```{r, echo=TRUE}
summary(model) # The summary function is used to get more information about the fitted regression model #
```
The output shows estimated regression coefficients, their standard errors (deviations), test statistics & p-value for testing significance. Also, there is R squared, degrees of freedom and F test statistic for testing significance of fitted model.
You can use the abline function as shown below to plot the fitted regression model on the scatterplot.
```{r, echo=TRUE}
plot(TVads$Spend, TVads$MilImp, pch=16, xlab = "TV advertising budget, 1983 ($ millions) ", ylab = "Millions of retained impressions per week", main="Figure 2: Fitted rgeression model oevrlayed on the scatterplot of Millions of retained impressions per week and TV advertising budget") # Create the scatterplot
abline(model, col = "red", lwd= 3) # Draw the fitted line #
```


The 95% CI for parameter estimates can be obtained using the following function: 
```{r, echo=TRUE}
confint(model, level = 0.95)
```

The 95% confidence interval for slope of the regression line is (0.1599, 0.5664).
If you want to get the predicted value of response for each value of predictor, then use the **predict** function as shown below.
```{r, echo=TRUE}
predict(model, TVads)
```

This same function can be used to get prediction interval for the each value of predictor as shown below. 

```{r, echo=TRUE}
predict(model, TVads, interval = "prediction", level = 0.95)
```

You may want to compute prediction values of response for a given set of new values of predictor. In that case, first, create a new data file with the new value of predictor. Then use the predictor function to predict response as shown below. Here, our new values of the Spend variable are 50 and 100.
```{r, echo=TRUE}
NewData <- data.frame(Spend = c(50, 100)) # Creates a new data frame with only new values of predictor variable
predict(model, NewData, interval = "prediction", level = 0.95) # Use the new data frame in the Predictor function instead of old data set
```

The above output shows that there is $42.32 million of average retained impressions per week when TV advertisement budget is $ 50 million.
To check the model diagnostics using residual plots, use the code below.
```{r, echo=TRUE}
plot(model) # This plots the diagnostic graphs for the fitted regression model. Fit enter key 4 times to see the plots shown below #
```

When you want to do multiple linear regression in R, you can add other predictors to the lm function each separated by a plus sign.

## The Solution by tidyverse
```{r, eval=FALSE}
install.packages("tidyverse")
```

To use R packages you have installed, include this line in your script or in the Console:

```{r, echo=TRUE}
library("tidyverse")
# Note here the quotation marks are optional and
# it is the same as
library(tidyverse)
```

### Import Data

1. Download the data file [here](TVads.txt) .
2. What format is the data file in?
3. Import the data in the file into R as a data frame:

```{r, echo=TRUE}
my_df <- read_table2("TVads.txt")
my_df
```

Of course, the step of import data into your statistic software vary a little by software and by the data format. 

More information as of how to import data into R can be found at [Import Data]()

### Descriptive Statistics

#### Visualization

1. Visualize a single variable
```{r, echo=TRUE}
ggplot(my_df, aes(x=MilImp)) + geom_histogram()
```

```{r, echo=TRUE}
ggplot(my_df, aes(x=Spend)) + geom_histogram()
```

2. Visualize a pair of numeric variables
```{r, echo=TRUE}
ggplot(my_df, aes(x=Spend, y=MilImp)) + geom_point()
```

### Correlation

```{r, echo=TRUE}
cor(my_df$MilImp, my_df$Spend)
```

### Regression
In R, run linear regressions with `lm` (short for **l**inear **m**odel):

```{r, echo=TRUE}
lm(MilImp ~ Spend, data=my_df)
```

### More detailed regression results

1. Pass the results from `lm()` to `summary()` for more detailed information:
```{r, echo=TRUE}
lm(MilImp ~ Spend, data=my_df) %>%
  summary
```

For better formatting of the results (pretty print), we can use the `texreg` package:
```{r, echo=TRUE}
## Install and load texreg package
install.packages("texreg")
library(texreg)

# Pretty print regression results on screen
lm(MilImp ~ Spend, data=my_df) %>%
  screenreg

# Save regression results to a html file
lm(MilImp ~ Spend, data=my_df) %>%
  htmlreg(file="lm_MilImp-Spend.html")
```

### Visualize regression results

```{r, echo=TRUE}
ggplot(my_df, aes(x=Spend, y=MilImp)) + geom_point() + geom_smooth(method = "lm", se = FALSE)
```

### Diagnostic plots of regression

We will use the `ggfortify` package to generate the diagnostic plots for regression
```{r, echo=TRUE}
## Install and load ggfortify package
install.packages("ggfortify")
library(ggfortify)

lm(MilImp ~ Spend, data=my_df) %>%
  autoplot()

# Since we need the regression results in many places, 
# it would be easier to save it into a R variable
MilImp_lm <- lm(MilImp ~ Spend, data=my_df)

# Save the diagnostic plots as a png file
ggsave("MilImp_lm_diag.png")

```


## Resources
- [Introduction to R on Data Camp](https://campus.datacamp.com/courses/free-introduction-to-r/chapter-1-intro-to-basics-1?ex=1): A self-instruction course covering R basics.
- [R coding basics on Software Carpentry](http://swcarpentry.github.io/r-novice-gapminder/), hands-on tutorial on R basics, lesson 4-11.
