---
title: "STAT566 Homework"
author: ''
date: "Spring 2019"
output: html_document
---

#  {.tabset .tabset-fade .tabset-pills}

## HW1

### Problem 1: In a home economic experiment, two temperatures (factor A) and two lengths of washing time (factor B) gave the following treatment means:

|     |(1) |a	  |b	 |ab
| --- | -- | -- | -- | --
|Mean	|6.2 |14.1|5.4 |7.8

The experiment was conducted as a completely randomized design with n=5 for teach treatment combination. The variable of interest was the amount of impurities (in grams) remaining in the material washed. Assume fixed effects for both factors.

(i) Compute the numerical values of the following effects and interpret in terms of question context:

	 (a) main effect of A

$$A=\frac12(\bar y_a+\bar y_{ab}-\bar y_{(1)}-\bar y_b)=\frac12(14.1+7.8-6.2-5.4)=5.15$$
	
	 (b) main effect of B

$$B=\frac12(\bar y_b+\bar y_{ab}-\bar y_{(1)}-\bar y_a)=\frac12(5.4+7.8-6.2-14.1)=-3.55$$

	 (c) interaction effect of A and B
	
$$AB=\frac12(\bar y_{ab}+\bar y_{(1)}-\bar y_a-\bar y_b)=\frac12(7.8+6.2-14.1-5.4)=-2.75$$
	
(ii) Explain which effect is highest from the effects in part (i).

The main effect of A is highest and main effect of B is smaller than that while it is higher than interaction effect.

(iii) Compute the sum of square for each effect in part (i).

$$SS_A=\frac{n}4C_A^2=\frac{n}4(\bar y_a+\bar y_{ab}-\bar y_{(1)}-\bar y_b)^2=\frac54(14.1+7.8-6.2-5.4)^2=132.6125$$
$$SS_B=\frac{n}4C_B^2=\frac{n}4(\bar y_b+\bar y_{ab}-\bar y_{(1)}-\bar y_a)^2=\frac54(5.4+7.8-6.2-14.1)^2=63.0125$$
$$SS_{AB}=\frac{n}4C_{AB}^2=\frac{n}4(\bar y_{ab}+\bar y_{(1)}-\bar y_a-\bar y_b)^2=\frac54(7.8+6.2-14.1-5.4)^2=37.8125$$

```{r eval=FALSE, include=FALSE}
(14.1+7.8-6.2-5.4)^2*5/4
(5.4+7.8-6.2-14.1)^2*5/4
(7.8+6.2-14.1-5.4)^2*5/4
29.8*16+132.6125+63.0125+37.8125

132.6125/29.8
63.0125/29.8
37.8125/29.8

df(132.6125/29.8, 1, 16,log= FALSE)
df(63.0125/29.8, 1, 16,log= FALSE)
df(37.8125/29.8, 1, 16,log= FALSE)

pf(132.6125/29.8, 1, 16, lower.tail =F, log.p = FALSE)
pf(63.0125/29.8, 1, 16, lower.tail =F, log.p = FALSE)
pf(37.8125/29.8, 1, 16, lower.tail =F, log.p = FALSE)

1-pf(132.6125/29.8, 1, 16, lower.tail = TRUE, log.p = FALSE)
1-pf(63.0125/29.8, 1, 16, lower.tail = TRUE, log.p = FALSE)
1-pf(37.8125/29.8, 1, 16, lower.tail = TRUE, log.p = FALSE)
```

(iv) If the MSE = 29.8, then compute all the terms of the ANOVA table for this study.

|Source of Variation|   SS          |      df      | MS     | F      |P-value   |
|         ---       |   ---         |   ---        | ---    | ---    | ---      |
|Factor A           |132.6125       |   $2-1=1$    |132.6125|4.450084|0.0510038 |
|Factor B           | 63.0125       |   $2-1=1$    | 63.0125|2.114513|0.1652436 |
|Interaction AB     | 37.8125       |$(2-1)(2-1)=1$| 37.8125|1.268876|0.2765897 |
|Error              |$29.8*16=476.8$|$4*(5-1)=16$  | 29.8   |
|Total|$132.6125+63.0125+37.8125+476.8=710.2375$|$2^2*5-1=19$  |

(v) Based on your ANOVA table, explain which of effects in part (i) are statistically significant at 5% significance level.

Based on the ANOVA table, the P-value of A is less than 0.05. The effects of factor A is significant at 5% significance level.

### Problem 2: 

(i) Create a table of plus and minus for a $2^5$ factorial design. Make sure to use standard order of letter assignments for treatment combinations.

|     | A	| B	| C | D | E | AB | AC | AD | AE | BC | BD | BE | CD | CE | DE | ABC | ABD | ABE | ACD | ACE | ADE | BCD | BCE | BDE | CDE | ABCD | ABCE | ABDE | ACDE | BCDE | ABCDE |
| --- | --| --| --| --| --| -- | -- | -- | -- | -- | -- | -- | -- | -- | -- | --  | --  | --  | --  | --  | --  | --  | --  | --  | --  |  --  |  --  |  --  |  --  |  --  |  --   |
|(1)  | - | - | - | - | - | +  | +  | +  | +  | +  | +  | +  | +  | +  | +  | -   | -   | -   | -   | -   | -   | -   | -   | -   | -   |  +   |  +   |  +   |  +   |  +   |  -    |
|a    | + | - | - | - | - | -  | -  | -  | -  | +  | +  | +  | +  | +  | +  | +   | +   | +   | +   | +   | +   | -   | -   | -   | -   |  -   |  -   |  -   |  -   |  +   |  +    |
|b    | - | + | - | - | - | -  | +  | +  | +  | -  | -  | -  | +  | +  | +  | +   | +   | +   | -   | -   | -   | +   | +   | +   | -   |  -   |  -   |  -   |  +   |  -   |  +    |
|ab   | + | + | - | - | - | +  | -  | -  | -  | -  | -  | -  | +  | +  | +  | -   | -   | -   | +   | +   | +   | +   | +   | +   | -   |  +   |  +   |  +   |  -   |  -   |  -    |
|c    | - | - | + | - | - | +  | -  | +  | +  | -  | +  | +  | -  | -  | +  | +   | -   | -   | +   | +   | -   | +   | +   | -   | +   |  -   |  -   |  +   |  -   |  -   |  +    |
|ac   | + | - | + | - | - | -- | +  | -  | -  | -  | +  | +  | -  | -  | +  | -   | +   | +   | -   | -   | +   | +   | +   | -   | +   |  +   |  +   |  -   |  +   |  -   |  -    |
|bc   | - | + | + | - | - | -  | -  | +  | +  | +  | -  | -  | -  | -  | +  | -   | --  | --  | --  | --  | --  | -   | -   | --  | --  |  --  |  --  |  --  |  --  |  --  |  --   |
|abc  | --| --| --| --| --| -- | -- | -- | -- | -- | -- | -- | -- | -- | -- | --  | --  | --  | --  | --  | --  | --  | --  | --  | --  |  --  |  --  |  --  |  --  |  --  |  --   |
|d    | - | - | - | + | - | +  | +  | -  | +  | +  | -  | +  | -  | +  | -  | -   | +   | -   | +   | -   | +   | +   | -   | +   | +   |  -   |  +   |  -   |  -   |  -   |  +    |
|ad   | --| --| --| --| --| -- | -- | -- | -- | -- | -- | -- | -- | -- | -- | --  | --  | --  | --  | --  | --  | --  | --  | --  | --  |  --  |  --  |  --  |  --  |  --  |  --   |
|bd   | --| --| --| --| --| -- | -- | -- | -- | -- | -- | -- | -- | -- | -- | --  | --  | --  | --  | --  | --  | --  | --  | --  | --  |  --  |  --  |  --  |  --  |  --  |  --   |
|cd   | --| --| --| --| --| -- | -- | -- | -- | -- | -- | -- | -- | -- | -- | --  | --  | --  | --  | --  | --  | --  | --  | --  | --  |  --  |  --  |  --  |  --  |  --  |  --   |
|abd  | --| --| --| --| --| -- | -- | -- | -- | -- | -- | -- | -- | -- | -- | --  | --  | --  | --  | --  | --  | --  | --  | --  | --  |  --  |  --  |  --  |  --  |  --  |  --   |
|acd  | --| --| --| --| --| -- | -- | -- | -- | -- | -- | -- | -- | -- | -- | --  | --  | --  | --  | --  | --  | --  | --  | --  | --  |  --  |  --  |  --  |  --  |  --  |  --   |
|bcd  | --| --| --| --| --| -- | -- | -- | -- | -- | -- | -- | -- | -- | -- | --  | --  | --  | --  | --  | --  | --  | --  | --  | --  |  --  |  --  |  --  |  --  |  --  |  --   |
|abcd | --| --| --| --| --| -- | -- | -- | -- | -- | -- | -- | -- | -- | -- | --  | --  | --  | --  | --  | --  | --  | --  | --  | --  |  --  |  --  |  --  |  --  |  --  |  --   |
|e    | - | - | - | - | + | +  | +  | +  | -  | +  | +  | -  | +  | -  | -  | -   | -   | +   | -   | +   | +   | -   | +   | +   | +   |  +   |  -   |  -   |  -   |  -   |  +    |
|ae   | --| --| --| --| --| -- | -- | -- | -- | -- | -- | -- | -- | -- | -- | --  | --  | --  | --  | --  | --  | --  | --  | --  | --  |  --  |  --  |  --  |  --  |  --  |  --   |
|be   | --| --| --| --| --| -- | -- | -- | -- | -- | -- | -- | -- | -- | -- | --  | --  | --  | --  | --  | --  | --  | --  | --  | --  |  --  |  --  |  --  |  --  |  --  |  --   |
|ce   | --| --| --| --| --| -- | -- | -- | -- | -- | -- | -- | -- | -- | -- | --  | --  | --  | --  | --  | --  | --  | --  | --  | --  |  --  |  --  |  --  |  --  |  --  |  --   |
|de   | --| --| --| --| --| -- | -- | -- | -- | -- | -- | -- | -- | -- | -- | --  | --  | --  | --  | --  | --  | --  | --  | --  | --  |  --  |  --  |  --  |  --  |  --  |  --   |
|abe  | --| --| --| --| --| -- | -- | -- | -- | -- | -- | -- | -- | -- | -- | --  | --  | --  | --  | --  | --  | --  | --  | --  | --  |  --  |  --  |  --  |  --  |  --  |  --   |
|ace  | --| --| --| --| --| -- | -- | -- | -- | -- | -- | -- | -- | -- | -- | --  | --  | --  | --  | --  | --  | --  | --  | --  | --  |  --  |  --  |  --  |  --  |  --  |  --   |
|ade  | --| --| --| --| --| -- | -- | -- | -- | -- | -- | -- | -- | -- | -- | --  | --  | --  | --  | --  | --  | --  | --  | --  | --  |  --  |  --  |  --  |  --  |  --  |  --   |
|bce  | --| --| --| --| --| -- | -- | -- | -- | -- | -- | -- | -- | -- | -- | --  | --  | --  | --  | --  | --  | --  | --  | --  | --  |  --  |  --  |  --  |  --  |  --  |  --   |
|bde  | --| --| --| --| --| -- | -- | -- | -- | -- | -- | -- | -- | -- | -- | --  | --  | --  | --  | --  | --  | --  | --  | --  | --  |  --  |  --  |  --  |  --  |  --  |  --   |
|cde  | --| --| --| --| --| -- | -- | -- | -- | -- | -- | -- | -- | -- | -- | --  | --  | --  | --  | --  | --  | --  | --  | --  | --  |  --  |  --  |  --  |  --  |  --  |  --   |
|cde  | --| --| --| --| --| -- | -- | -- | -- | -- | -- | -- | -- | -- | -- | --  | --  | --  | --  | --  | --  | --  | --  | --  | --  |  --  |  --  |  --  |  --  |  --  |  --   |
|abce | --| --| --| --| --| -- | -- | -- | -- | -- | -- | -- | -- | -- | -- | --  | --  | --  | --  | --  | --  | --  | --  | --  | --  |  --  |  --  |  --  |  --  |  --  |  --   |
|abde | --| --| --| --| --| -- | -- | -- | -- | -- | -- | -- | -- | -- | -- | --  | --  | --  | --  | --  | --  | --  | --  | --  | --  |  --  |  --  |  --  |  --  |  --  |  --   |
|acde | --| --| --| --| --| -- | -- | -- | -- | -- | -- | -- | -- | -- | -- | --  | --  | --  | --  | --  | --  | --  | --  | --  | --  |  --  |  --  |  --  |  --  |  --  |  --   |
|bcde | --| --| --| --| --| -- | -- | -- | -- | -- | -- | -- | -- | -- | -- | --  | --  | --  | --  | --  | --  | --  | --  | --  | --  |  --  |  --  |  --  |  --  |  --  |  --   |
|abcde| + | + | + | + | + | +  | +  | +  | +  | +  | +  | +  | +  | +  | +  | +   | +   | +   | +   | +   | +   | +   | +   | +   | +   |  +   |  +   |  +   |  +   |  +   |  +    |


```{r eval=FALSE, include=FALSE}
install.packages("FrF2")
library(FrF2)
FrF2(32,5, factor.names = c('RaceA', 'RaceB','CrimHistA', 'CrimHistB', 'Veteran','Children'), randomize = FALSE)
design.info(FrF2(32,5,MaxC2=TRUE))

 install.packages("AlgDesign")
library(AlgDesign)
gen.factorial(levels = 2, nVars = 5, varNames=c("A","B","C","D","E"))

```

(ii) Write a contrast statement for the four factor interaction effect using standard order of letters as we did in class. Assume n replications for each treatment combination.



### Problem 3: Use any software to do this problem


In part (a), show the normal probabiltiy plot. Then explain and name the effects that are neglegible.

In Part (b), write the theoretical model fit for these data without neglegible terms.


